### 2012-09-05 manuscript.tex
Add Fernando's paper template

%
% ICRA 2013 Paper Submission
% created on Aug 31, 2012
%

%!TEX TS-program = latex
%!TEX encoding = UTF-8 Unicode

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts % for \thanks
\overrideIEEEmargins % for \addtolength

% Packages
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[pdftex]{hyperref}

% hyperref setup
\definecolor{linkCol}{gray}{0.25}
\hypersetup{
    pdftitle={},
    pdfauthor={},
    pdfsubject={},
    pdfkeywords={},
    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
}

\title{}
\author{}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{}

%----------------------------------------------------------------------------%
\subsection{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addtolength{\textheight}{-0cm}
                        % This command serves to balance the column lengths
                        % on the last page of the document manually. It shortens
                        % the textheight of the last page by a suitable amount.
                        % This command does not take effect until the next page
                        % so it should come on the page before the last. Make
                        % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS AND FUTURE WORK}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ACKNOWLEDGMENTS}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{IEEEtranS}
%\bibliography{}

\end{document}

### 2012-09-06 manuscript.tex
Manuscript template tested
Add preliminary title and outline

@@
 % ICRA 2013 Paper Submission
-% created on Aug 31, 2012
+% created on Sep 5, 2012
 %
@@
 \hypersetup{
-    pdftitle={},
-    pdfauthor={},
+    pdftitle={Multi-Camera Visual Servoing of a Flapping Wing Micro Air Vehicle},
+    pdfauthor={Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing},
     pdfsubject={},
@@
 
-\title{}
-\author{}
+\title{\LARGE \bf Multi-Camera Visual Servoing of a Flapping Wing Micro Air Vehicle}
+
+\author{
+    Ryan C. Julian$^{*}$,
+    Cameron J. Rose$^{*}$,
+    Humphrey Hu$^{\dagger}$, and
+    Ronald S. Fearing$^{*}$
+    \thanks{$^{*}$
+    	R.C. Julian, C.J. Rose and R.S. Fearing are
+        with the Department of Electrical Engineering and Computer Sciences,
+        University of California, Berkeley, CA 94720 USA.
+        \href{mailto:ryanjulian@berkeley.edu}{\tt ryanjulian@berkeley.edu},
+        \href{mailto:c\_rose@eecs.berkeley.edu}{\tt c\_rose@eecs.berkeley.edu},
+        \href{mailto:ronf@eecs.berkeley.edu}{\tt ronf@eecs.berkeley.edu}
+    }
+    \thanks{$^{\dagger}$
+        H. Hu is with the Robotics Institute,
+        Carnegie Mellon University, Pittsburgh, PA 15213 USA.
+        \href{mailto:humphreh@andrew.cmu.edu.edu}{\tt humphreh@andrew.cmu.edu}
+    }
+%    \thanks{
+%        This material is based upon work supported by the National Science
+%        Foundation under Grants No. CNS-0931463 and IIS-0931463.
+%    }
+}
 
@@
 \begin{abstract}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{INTRODUCTION}
+\section{Introduction}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird.jpg}
+\caption{H2Bird Ornithopter Platform}
+\label{fig:H2Bird}
+\end{figure}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{}
+\section{Robotic and Vision Platforms}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
 %----------------------------------------------------------------------------%
-\subsection{}
+\subsection{Flapping Wing Micro Air Vehicle}
+%TODO(humphrey)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Computer Vision Platform}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{System Concept and Algorithms}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{MAV Internal Control}
+%TODO(humphrey)
+Due to the high pitch angles experienced by the IBird, a quaternion representation of vehicle pose was chosen to avoid Euler angle singularities.
+A quaternion-based controller as used in was adapted to control the iBird thrust, yaw, and pitch actuators.
+Given a reference attitude q, the error qe is dened as the rotation required to reach the reference from the current pose.
+The components of the error quaternion are then linearized to produce body axis angular errors.
+Control of the pitch and yaw axes is handled by three PID loops commanding the wing thrust, elevator, and tail prop.
+This independent PID loop architecture was chosen for simplicity in tuning and implementation, as well as acceptable performance in the absence of a system model.
+The iBird currently does not have a roll actuator, but our current control scheme has sufficed for the reference tracking in this work.
+
+%----------------------------------------------------------------------------%
+\subsection{Pose Estimation}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Visual Servoing}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Window Seeking}
+%TODO(cameron)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Orbiting}
+%TODO(cameron)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Performance}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{MAV Internal Control}
+%TODO(cameron)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Pose Estimation}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Visual Servoing}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{CONCLUSIONS AND FUTURE WORK}
+\section{Conclusions and Future Work}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{ACKNOWLEDGMENTS}
+\section{Acknowledgements}
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 \bibliographystyle{IEEEtranS}
-%\bibliography{}
+\bibliography{}
 

### 2012-09-11 manuscript.tex
Updated Flapping Wing MAV section.

@@
     }
-%    \thanks{
-%        This material is based upon work supported by the National Science
-%        Foundation under Grants No. CNS-0931463 and IIS-0931463.
-%    }
+    \thanks{
+        This material is based upon work supported by the National Science
+        Foundation under Grants No. CNS-0931463, and by the Army Research
+        Laboratory under the Micro Autonomous System and Technology
+        Collaborative Technology Alliance.
+    }
 }
@@
 %TODO(humphrey)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
+The flying platform used in this experiment is the H2Bird flapping wing MAV.
+It features a X-wing clap-fling flapping design, with a brushed DC motor powering the wings for propulsion and a tail propeller and servo-controlled tail elevator for control.
+Onboard electronics include a 40 MIPS microprocessor, 6 DoF IMU, 802.15.4 radio, VGA camera, and motor drivers powered by a 90 mAh lithium polymer battery.
+In routine flight, the H2Bird achieves [insert speed] m/s ground speed and can operate for approximately [10] minutes.
 

### 2012-09-12 manuscript.tex
Updates to MAV platform and control sections. - Added guesses to platform parameters - Added control equations and explanations

@@
 The flying platform used in this experiment is the H2Bird flapping wing MAV.
-It features a X-wing clap-fling flapping design, with a brushed DC motor powering the wings for propulsion and a tail propeller and servo-controlled tail elevator for control.
-Onboard electronics include a 40 MIPS microprocessor, 6 DoF IMU, 802.15.4 radio, VGA camera, and motor drivers powered by a 90 mAh lithium polymer battery.
-In routine flight, the H2Bird achieves [insert speed] m/s ground speed and can operate for approximately [10] minutes.
+Built around the Silverlit I-Bird RC flyer powertrain and clap-fling wings, the H2Bird has a wingspan of [40?] cm and weighs [14?] grams fully loaded.
+A tail propeller and servo-controlled tail elevator provide control in the yaw and pitch axes.
+An onboard electronics package combines a 40 MIPS microprocessor, 6 DoF IMU, 802.15.4 radio, VGA camera, and motor drivers with a 90 mAh lithium polymer battery for autonomous flight.
+In routine flight, the H2Bird averages [insert speed] m/s ground speed and can operate for approximately [10] minutes.
 
@@
 %----------------------------------------------------------------------------%
-\subsection{MAV Internal Control}
+\subsection{MAV Onboard Control}
 %TODO(humphrey)
-Due to the high pitch angles experienced by the IBird, a quaternion representation of vehicle pose was chosen to avoid Euler angle singularities.
-A quaternion-based controller as used in was adapted to control the iBird thrust, yaw, and pitch actuators.
-Given a reference attitude q, the error qe is dened as the rotation required to reach the reference from the current pose.
+Rudimentary attitude control for the H2Bird runs onboard, estimating pose by naively integrating the IMU and depending on PID for control.
+The simplistic nature of our state estimation and control is due to the stringent resource constraints applied by the platform.
+
+A quaternion representation of vehicle pose was chosen to avoid Euler angle singularities.
+This choice of pose representation also lends itself well to remote guidance, as relative body and world coordinate angular displacements can be represented compactly with quaternion multiplication.
+
+Given a reference attitude $q$, the error $q_e$ is defined as the rotation required to reach the reference from the current pose.
 The components of the error quaternion are then linearized to produce body axis angular errors.
 Control of the pitch and yaw axes is handled by three PID loops commanding the wing thrust, elevator, and tail prop.
-This independent PID loop architecture was chosen for simplicity in tuning and implementation, as well as acceptable performance in the absence of a system model.
-The iBird currently does not have a roll actuator, but our current control scheme has sufficed for the reference tracking in this work.
+
+\begin{equation}
+\label{quat_error}
+q_e = q'\otimes q_r
+\end{equation}
+
+\begin{equation}
+\label{quat_angle}
+\alpha = 2acos(q_{e,w})
+\end{equation}
+
+\begin{equation}
+\label{quat_linearize}
+Q_e = \alpha*q_e/sin(\alpha /2)
+\end{equation}
 

### 2012-09-14 manuscript.tex
Begin figures mock-up
- Added start of abstract
- New title
- Acknowledgements

--- 2012-09-06 manuscript.tex
@@
 %
+\def \papertitle{Cooperative Control for Narrow Passage Traversal with a Flapping-Wing Micro Air Vehicle and Lightweight Ground Station}
+
+% NOTE: WE WRITE ALL TEXT IN FIRST PERSON PLURAL ("we")
+% Conventions: flapping-wing, small-scale
 
@@
 \hypersetup{
-    pdftitle={Multi-Camera Visual Servoing of a Flapping Wing Micro Air Vehicle},
+    pdftitle={\papertitle},
     pdfauthor={Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing},
@@
 
-\title{\LARGE \bf Multi-Camera Visual Servoing of a Flapping Wing Micro Air Vehicle}
+% Title
+\title{\LARGE \bf \papertitle}
 
+% Authors
 \author{
@@
         Carnegie Mellon University, Pittsburgh, PA 15213 USA.
-        \href{mailto:humphreh@andrew.cmu.edu.edu}{\tt humphreh@andrew.cmu.edu}
+        \href{mailto:humphreh@andrew.cmu.edu}{\tt humphreh@andrew.cmu.edu}
     }
-%    \thanks{
-%        This material is based upon work supported by the National Science
-%        Foundation under Grants No. CNS-0931463 and IIS-0931463.
-%    }
+    \thanks{
+        This material is based upon work supported by the National Science
+        Foundation under Grants No. CNS-0931463, and by the Army Research
+        Laboratory under the Micro Autonomous System and Technology
+        Collaborative Technology Alliance.
+    }
 }
@@
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
+The power, size, and weight constraints of micro air vehicle (MAV) platforms limit their on-board sensing and computational resources.
+In this work, we demonstrate cooperative control of a 14 gram ornithopter MAV by a lightweight ground station using computer vision.
+We also introduce a new ornithopter MAV platform, the h$^2$Bird, which...
 \end{abstract}
@@
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/h2bird.jpg}
-\caption{H2Bird Ornithopter Platform}
-\label{fig:H2Bird}
-\end{figure}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
+The power, size, and weight constraints of micro air vehicle (MAV) platforms significantly limit on-board sensing and computational resources, and thus the complexity of the control and state estimation algorithms which can be executed on-board.
+These constraints are particularly severe for flapping wing MAVs offer the promise of superior safety, efficiency, and noise profile to rotor craft MAVs, but at the expense of payload weight.
 
@@
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
 
 %----------------------------------------------------------------------------%
-\subsection{Flapping Wing Micro Air Vehicle}
+\subsection{h$^2$Bird Ornithoper MAV}
 %TODO(humphrey)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird.jpg}
+\caption{h$^2$Bird Ornithopter MAV}
+\label{fig:H2Bird}
+\end{figure}
 
 %----------------------------------------------------------------------------%
-\subsection{Computer Vision Platform}
+\subsection{Ground Station Computer Vision Platform}
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
 
 %----------------------------------------------------------------------------%
-\subsection{MAV Internal Control}
+\subsection{Ornithoper Flight Control}
 %TODO(humphrey)
-Due to the high pitch angles experienced by the IBird, a quaternion representation of vehicle pose was chosen to avoid Euler angle singularities.
-A quaternion-based controller as used in was adapted to control the iBird thrust, yaw, and pitch actuators.
-Given a reference attitude q, the error qe is dened as the rotation required to reach the reference from the current pose.
-The components of the error quaternion are then linearized to produce body axis angular errors.
-Control of the pitch and yaw axes is handled by three PID loops commanding the wing thrust, elevator, and tail prop.
-This independent PID loop architecture was chosen for simplicity in tuning and implementation, as well as acceptable performance in the absence of a system model.
-The iBird currently does not have a roll actuator, but our current control scheme has sufficed for the reference tracking in this work.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird_axes.jpg}
+\caption{h$^2$Bird Ornithopter with flight control axes}
+\label{fig:H2Bird}
+\end{figure}
+
+We use simple controllers to actuate the h$^2$Bird's wing beat frequency, tail rotor, and tail elevator, which we designed to actuate thrust, yaw, and pitch respectively.
+Due to the high pitch angles experienced by the h$^2$Bird, we use a quaternion representation of vehicle attitude to avoid the numerical instabilities of Euler angle representations.
+Given a reference attitude $q$, we define the error $q_e$ as the rotation required to reach the reference attitude from the measured attitude.
+We then linearize the error quaternion to produce body axis angular errors.
+Three independent proportional-integral-derivative controllers commanding the wing thrust, elevator, and tail propeller.
+We use this independent PID controller architecture for simplicity in tuning and implementation, and because it demonstrates acceptable performance in the absence of a system model.
 
@@
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
-
-%----------------------------------------------------------------------------%
-\subsection{Window Seeking}
-%TODO(cameron)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
-
-%----------------------------------------------------------------------------%
-\subsection{Orbiting}
-%TODO(cameron)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(cameron)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 \section{Acknowledgements}
-%TODO(ryan)
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed tincidunt erat et dolor commodo varius.
-Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
-Aliquam sit amet felis nibh, in bibendum est.
-Ut augue tortor, elementum vel bibendum ac, faucibus at est.
-Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
-Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
-Donec ac nulla tellus.
-Aenean viverra rhoncus odio a sagittis.
-Quisque gravida imperdiet lorem ut condimentum.
-Pellentesque eget quam dolor.
-Vestibulum a eros nisl, quis malesuada ipsum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
-Etiam consequat molestie massa, a pellentesque magna tincidunt non.
-Nullam aliquet augue et risus hendrerit facilisis.
-
+The authors would like to thank Fernando L.
+Garcia Bermudez for his assistance with the Vicon motion capture system, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-09-14 manuscript.tex
Add labeled h2Bird portrait.

@@
 \section{Acknowledgements}
-The authors would like to thank Fernando L.
-Garcia Bermudez for his assistance with the Vicon motion capture system, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
+The authors would like to thank Fernando Garcia Bermudez for his assistance with the Vicon motion capture system, Andrew Pullin for his help with robot photography, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-09-20 manuscript.tex
Pre-reinstall commit

@@
 %
-\def \papertitle{Cooperative Control for Narrow Passage Traversal with a Flapping-Wing Micro Air Vehicle and Lightweight Ground Station}
+\def \papertitle{Cooperative Control for Window Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station}
 
@@
 %----------------------------------------------------------------------------%
-\subsection{h$^2$Bird Ornithoper MAV}
+\subsection{H$^2$Bird Ornithoper MAV}
 %TODO(humphrey)
+
 \begin{figure}[tb]
@@
 \includegraphics[width=\linewidth]{figures/h2bird.jpg}
-\caption{h$^2$Bird Ornithopter MAV}
-\label{fig:H2Bird}
+\caption{H$^2$Bird Ornithopter MAV}
+\label{fig:h2bird}
 \end{figure}
+
+The flying platform used in this experiment is the H$^2$Bird flapping wing MAV.
+Built around the Silverlit I-Bird RC flyer powertrain and clap-fling wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams fully loaded.
+A tail propeller and servo-controlled tail elevator provide control in the yaw and pitch axes.
+An onboard electronics package combines a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor drivers with a 90 mAh lithium polymer battery for autonomous flight.
+In routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and can operate for approximately [10] minutes.
 
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
+\caption{Overall block diagram of cooperative control}
+\label{fig:block_diagram}
+\end{figure}
+
 %----------------------------------------------------------------------------%
@@
 \includegraphics[width=\linewidth]{figures/h2bird_axes.jpg}
-\caption{h$^2$Bird Ornithopter with flight control axes}
-\label{fig:H2Bird}
+\caption{H$^2$Bird Ornithopter with flight control axes}
+\label{fig:h2Bird_axes}
 \end{figure}
 
-We use simple controllers to actuate the h$^2$Bird's wing beat frequency, tail rotor, and tail elevator, which we designed to actuate thrust, yaw, and pitch respectively.
-Due to the high pitch angles experienced by the h$^2$Bird, we use a quaternion representation of vehicle attitude to avoid the numerical instabilities of Euler angle representations.
-Given a reference attitude $q$, we define the error $q_e$ as the rotation required to reach the reference attitude from the measured attitude.
-We then linearize the error quaternion to produce body axis angular errors.
+Rudimentary attitude control for the H$^2$Bird runs on-board.
+It estimates pose by naively integrating the IMU and depending on PID for control.
+The simplistic nature of our state estimation and control is due to the stringent resource constraints applied by the platform.
+
+We use simple controllers to actuate the H$^2$Bird's wing beat frequency, tail rotor, and tail elevator, which we designed to actuate thrust, yaw, and pitch respectively.
+
+Due to the high pitch angles experienced by the H$^2$Bird, we use a quaternion representation of vehicle pose to avoid the numerical instabilities of Euler angle representations.
+This choice of pose representation also lends itself well to remote guidance, as relative body and world coordinate angular displacements can be represented compactly with quaternion multiplication.
+
+Given a reference pose $q$, we define the error $q_e$ as the rotation required to reach the reference pose from the measured pose.
+We then linearize the error quaternion to produce body axis angular errors as as inputs to the attitude controllers.
+
 Three independent proportional-integral-derivative controllers commanding the wing thrust, elevator, and tail propeller.
 We use this independent PID controller architecture for simplicity in tuning and implementation, and because it demonstrates acceptable performance in the absence of a system model.
+
+\begin{equation}
+\label{quat_error}
+q_e = q'\otimes q_r
+\end{equation}
+
+\begin{equation}
+\label{quat_angle}
+\alpha = 2acos(q_{e,w})
+\end{equation}
+
+\begin{equation}
+\label{quat_linearize}
+Q_e = \alpha*q_e/sin(\alpha /2)
+\end{equation}
 
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird.jpg}
+\caption{Frame sequence from particle filter video, showing particles and state estimate}
+\label{fig:pf_screencap}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %TODO(ryan)
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/experiment_cartoon.png}
+\caption{Conceptual sketch of experimental environment.}
+\label{fig:experiment_sketch}
+\end{figure}
 
@@
 %----------------------------------------------------------------------------%
-\subsection{MAV Internal Control}
+\subsection{Ornithopter Flight Control}
 %TODO(cameron)
-
-%----------------------------------------------------------------------------%
-\subsection{Pose Estimation}
-%TODO(ryan)
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
+\caption{H$^2$Bird step response}
+\label{fig:step_response}
+\end{figure}
 
@@
 %TODO(ryan)
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/window_trials.jpg}
+\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius cone, with experimental trials overlayed.}
+\label{fig:experiment_cone}
+\end{figure}
 

### 2012-10-09 manuscript.tex
added some text

--- 2012-09-14 manuscript.tex
@@
 %
-\def \papertitle{Cooperative Control for Narrow Passage Traversal with a Flapping-Wing Micro Air Vehicle and Lightweight Ground Station}
-
-% NOTE: WE WRITE ALL TEXT IN FIRST PERSON PLURAL ("we")
-% Conventions: flapping-wing, small-scale
 
@@
 \hypersetup{
-    pdftitle={\papertitle},
+    pdftitle={Multi-Camera Visual Servoing of a Flapping Wing Micro Air Vehicle},
     pdfauthor={Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing},
@@
 
-% Title
-\title{\LARGE \bf \papertitle}
+\title{\LARGE \bf Multi-Camera Visual Servoing of a Flapping Wing Micro Air Vehicle}
 
-% Authors
 \author{
@@
         Carnegie Mellon University, Pittsburgh, PA 15213 USA.
-        \href{mailto:humphreh@andrew.cmu.edu}{\tt humphreh@andrew.cmu.edu}
+        \href{mailto:humphreh@andrew.cmu.edu.edu}{\tt humphreh@andrew.cmu.edu}
     }
-    \thanks{
-        This material is based upon work supported by the National Science
-        Foundation under Grants No. CNS-0931463, and by the Army Research
-        Laboratory under the Micro Autonomous System and Technology
-        Collaborative Technology Alliance.
-    }
+%    \thanks{
+%        This material is based upon work supported by the National Science
+%        Foundation under Grants No. CNS-0931463 and IIS-0931463.
+%    }
 }
@@
 %TODO(ryan)
-The power, size, and weight constraints of micro air vehicle (MAV) platforms limit their on-board sensing and computational resources.
-In this work, we demonstrate cooperative control of a 14 gram ornithopter MAV by a lightweight ground station using computer vision.
-We also introduce a new ornithopter MAV platform, the h$^2$Bird, which...
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 \end{abstract}
@@
 %TODO(ryan)
-The power, size, and weight constraints of micro air vehicle (MAV) platforms significantly limit on-board sensing and computational resources, and thus the complexity of the control and state estimation algorithms which can be executed on-board.
-These constraints are particularly severe for flapping wing MAVs offer the promise of superior safety, efficiency, and noise profile to rotor craft MAVs, but at the expense of payload weight.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird.jpg}
+\caption{H2Bird Ornithopter Platform}
+\label{fig:H2Bird}
+\end{figure}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
 %----------------------------------------------------------------------------%
-\subsection{h$^2$Bird Ornithoper MAV}
+\subsection{Flapping Wing Micro Air Vehicle}
 %TODO(humphrey)
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/h2bird.jpg}
-\caption{h$^2$Bird Ornithopter MAV}
-\label{fig:H2Bird}
-\end{figure}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
 %----------------------------------------------------------------------------%
-\subsection{Ground Station Computer Vision Platform}
+\subsection{Computer Vision Platform}
 %TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
 %----------------------------------------------------------------------------%
-\subsection{Ornithoper Flight Control}
+\subsection{MAV Internal Control}
 %TODO(humphrey)
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/h2bird_axes.jpg}
-\caption{h$^2$Bird Ornithopter with flight control axes}
-\label{fig:H2Bird}
-\end{figure}
-
-We use simple controllers to actuate the h$^2$Bird's wing beat frequency, tail rotor, and tail elevator, which we designed to actuate thrust, yaw, and pitch respectively.
-Due to the high pitch angles experienced by the h$^2$Bird, we use a quaternion representation of vehicle attitude to avoid the numerical instabilities of Euler angle representations.
-Given a reference attitude $q$, we define the error $q_e$ as the rotation required to reach the reference attitude from the measured attitude.
-We then linearize the error quaternion to produce body axis angular errors.
-Three independent proportional-integral-derivative controllers commanding the wing thrust, elevator, and tail propeller.
-We use this independent PID controller architecture for simplicity in tuning and implementation, and because it demonstrates acceptable performance in the absence of a system model.
+Due to the high pitch angles experienced by the IBird, a quaternion representation of vehicle pose was chosen to avoid Euler angle singularities.
+A quaternion-based controller as used in was adapted to control the iBird thrust, yaw, and pitch actuators.
+Given a reference attitude q, the error qe is dened as the rotation required to reach the reference from the current pose.
+The components of the error quaternion are then linearized to produce body axis angular errors.
+Control of the pitch and yaw axes is handled by three PID loops commanding the wing thrust, elevator, and tail prop.
+This independent PID loop architecture was chosen for simplicity in tuning and implementation, as well as acceptable performance in the absence of a system model.
+The iBird currently does not have a roll actuator, but our current control scheme has sufficed for the reference tracking in this work.
 
@@
 %TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
+%----------------------------------------------------------------------------%
+\subsection{Window Seeking}
+%TODO(cameron)
+The window seeking experiment was designed to demonstrate a method of cooperative control between outside observers and the IBird.
+Figure ~\ref{fig:WindowSetup} depicts the setup of the experiment.
+To steer the IBird to the center of the window, the
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis
+
+%----------------------------------------------------------------------------%
+\subsection{Orbiting}
+%TODO(cameron)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 %TODO(cameron)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
 
@@
 \section{Acknowledgements}
-The authors would like to thank Fernando L.
-Garcia Bermudez for his assistance with the Vicon motion capture system, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
+%TODO(ryan)
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed tincidunt erat et dolor commodo varius.
+Integer arcu metus, egestas vel tincidunt a, tincidunt quis massa.
+Aliquam sit amet felis nibh, in bibendum est.
+Ut augue tortor, elementum vel bibendum ac, faucibus at est.
+Etiam lacus libero, aliquet id eleifend vitae, sollicitudin et nunc.
+Sed condimentum, lacus nec ullamcorper vestibulum, elit quam eleifend augue, id iaculis est elit at tortor.
+Donec ac nulla tellus.
+Aenean viverra rhoncus odio a sagittis.
+Quisque gravida imperdiet lorem ut condimentum.
+Pellentesque eget quam dolor.
+Vestibulum a eros nisl, quis malesuada ipsum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Vestibulum diam augue, ultrices et consectetur ut, fringilla egestas magna.
+Etiam consequat molestie massa, a pellentesque magna tincidunt non.
+Nullam aliquet augue et risus hendrerit facilisis.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{IEEEtranS}
-\bibliography{}
+%\bibliography{}
 

### 2012-10-09 manuscript.tex

--- 2012-09-20 manuscript.tex
@@
 \begin{abstract}
-%TODO(ryan)
 The power, size, and weight constraints of micro air vehicle (MAV) platforms limit their on-board sensing and computational resources.
-In this work, we demonstrate cooperative control of a 14 gram ornithopter MAV by a lightweight ground station using computer vision.
-We also introduce a new ornithopter MAV platform, the h$^2$Bird, which...
+Ground vehicles have less mobility than MAVs, but relaxed size constrains and typically more computing power, creating a opportunity for cooperation between the two platforms.
+In this work, we demonstrate cooperative target-seeking by a 14 gram ornithopter MAV with a lightweight ground station using computer vision.
+We also introduce a new ornithopter MAV platform, the H$^2$Bird, which features clap and fling flight mechanics and improves upon a previous design with yaw and pitch control and improved payload.
+To augment the built-in navigation and sensing capabilities of the ornithopter, the ground station to provides heading estimates to the ornithoper by running a real-time motion tracking algorithm over the a live video stream.
+This allows the ornithopter to negotiate indoor navigation through a window.
+We discuss the performance of both the location estimation and navigation algorithms.
+We analytically determine the backwards-reachable region for successfully navigating through the window, and use it to determine the operational limits of the system.
+Finally, we verify this region through data collected in experimental trials.
+We achieve a success rate of X\% within the feasible region.
 \end{abstract}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/h2bird.jpg}
+\includegraphics[width=\linewidth]{figures/h2bird.png}
 \caption{H$^2$Bird Ornithopter MAV}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
-\caption{Overall block diagram of cooperative control}
-\label{fig:block_diagram}
+\includegraphics[width=\linewidth]{figures/process_flow.pdf}
+\caption{Process flow diagram for the cooperative control system.}
+\label{fig:process_flow}
 \end{figure}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/h2bird.jpg}
-\caption{Frame sequence from particle filter video, showing particles and state estimate}
+\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
+\caption{Overall block diagram of cooperative control}
+\label{fig:block_diagram}
+\end{figure}
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
+\caption{Frame sequence from video feed used for tracking, showing particles and state estimate}
 \label{fig:pf_screencap}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/window_trials.jpg}
+\includegraphics[width=\linewidth]{figures/flight_paths.pdf}
 \caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius cone, with experimental trials overlayed.}
-\label{fig:experiment_cone}
+\label{fig:flight_paths}
 \end{figure}

### 2012-10-10 manuscript.tex

--- 2012-09-20 manuscript.tex
@@
+% This is "aamas2013 .tex" September 2012 
+% This file should be compiled with "aamas2013 .cls" 
+% This example file demonstrates the use of the 'aamas2013 .cls'
+% LaTeX2e document class file. It is for those submitting
+% articles to AAMAS 2013  conference. This file is based on
+% the sig-alternate.tex example file.
+% The 'sig-alternate.cls' file of ACM will produce a similar-looking,
+% albeit, 'tighter' paper resulting in, invariably, fewer pages.
+% than the original style ACM style.
 %
-% ICRA 2013 Paper Submission
-% created on Sep 5, 2012
+% ----------------------------------------------------------------------------------------------------------------
+% This .tex file (and associated .cls ) produces:
+%       1) The Permission Statement
+%       2) The Conference (location) Info information
+%       3) The Copyright Line with AAMAS data
+%       4) NO page numbers
 %
-\def \papertitle{Cooperative Control for Window Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station}
+% as against the acm_proc_article-sp.cls file which
+% DOES NOT produce 1) thru' 3) above.
+%
+% Using 'aamas2013 .cls' you don't have control
+% from within the source .tex file, over both the CopyrightYear
+% (defaulted to 200X) and the IFAAMAS Copyright Data
+% (defaulted to X-XXXXX-XX-X/XX/XX).
+% These information will be overwritten by fixed AAMAS 2013  information
+% in the style files - it is NOT as you are used with ACM style files.
+%
+% ---------------------------------------------------------------------------------------------------------------
+% This .tex source is an example which *does* use
+% the .bib file (from which the .bbl file % is produced).
+% REMEMBER HOWEVER: After having produced the .bbl file,
+% and prior to final submission, you *NEED* to 'insert'
+% your .bbl file into your source .tex file so as to provide
+% ONE 'self-contained' source file.
+%
 
-% NOTE: WE WRITE ALL TEXT IN FIRST PERSON PLURAL ("we")
-% Conventions: flapping-wing, small-scale
+% This is the document class for full camera ready papers and extended abstracts repsectively
 
-%!TEX TS-program = latex
-%!TEX encoding = UTF-8 Unicode
+\documentclass{aamas2013}
 
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
-\IEEEoverridecommandlockouts % for \thanks
-\overrideIEEEmargins % for \addtolength
+% if you are using PDF LaTex and you cannot find a way for producing
+% letter, the following explicit settings may help
 
-% Packages
-\usepackage{color}
-\usepackage[pdftex]{graphicx}
-\usepackage{subfig}
-\usepackage{multirow}
-\usepackage{rotating}
-\usepackage[pdftex]{hyperref}
-
-% hyperref setup
-\definecolor{linkCol}{gray}{0.25}
-\hypersetup{
-    pdftitle={\papertitle},
-    pdfauthor={Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing},
-    pdfsubject={},
-    pdfkeywords={},
-    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
-    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
-}
-
-% Title
-\title{\LARGE \bf \papertitle}
-
-% Authors
-\author{
-    Ryan C. Julian$^{*}$,
-    Cameron J. Rose$^{*}$,
-    Humphrey Hu$^{\dagger}$, and
-    Ronald S. Fearing$^{*}$
-    \thanks{$^{*}$
-    	R.C. Julian, C.J. Rose and R.S. Fearing are
-        with the Department of Electrical Engineering and Computer Sciences,
-        University of California, Berkeley, CA 94720 USA.
-        \href{mailto:ryanjulian@berkeley.edu}{\tt ryanjulian@berkeley.edu},
-        \href{mailto:c\_rose@eecs.berkeley.edu}{\tt c\_rose@eecs.berkeley.edu},
-        \href{mailto:ronf@eecs.berkeley.edu}{\tt ronf@eecs.berkeley.edu}
-    }
-    \thanks{$^{\dagger}$
-        H. Hu is with the Robotics Institute,
-        Carnegie Mellon University, Pittsburgh, PA 15213 USA.
-        \href{mailto:humphreh@andrew.cmu.edu}{\tt humphreh@andrew.cmu.edu}
-    }
-    \thanks{
-        This material is based upon work supported by the National Science
-        Foundation under Grants No. CNS-0931463, and by the Army Research
-        Laboratory under the Micro Autonomous System and Technology
-        Collaborative Technology Alliance.
-    }
-}
+\pdfpagewidth=8.5truein
+\pdfpageheight=11truein
 
@@
 
+% In the original styles from ACM, you would have needed to
+% add meta-info here. This is not necessary for AAMAS 2013  as
+% the complete copyright information is generated by the cls-files.
+
+\title{Cooperative Control for Narrow Passage Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station}
+
+% AUTHORS
+
+% For initial submission, do not give author names, but the
+% tracking number, instead, as the review process is blind.
+
+% You need the command \numberofauthors to handle the 'placement
+% and alignment' of the authors beneath the title.
+%
+% For aesthetic reasons, we recommend 'three authors at a time'
+% i.e. three 'name/affiliation blocks' be placed beneath the title.
+%
+% NOTE: You are NOT restricted in how many 'rows' of
+% "name/affiliations" may appear. We just ask that you restrict
+% the number of 'columns' to three.
+%
+% Because of the available 'opening page real-estate'
+% we ask you to refrain from putting more than six authors
+% (two rows with three columns) beneath the article title.
+% More than six makes the first-page appear very cluttered indeed.
+%
+% Use the \alignauthor commands to handle the names
+% and affiliations for an 'aesthetic maximum' of six authors.
+% Add names, affiliations, addresses for
+% the seventh etc. author(s) as the argument for the
+% \additionalauthors command.
+% These 'additional authors' will be output/set for you
+% without further effort on your part as the last section in
+% the body of your article BEFORE References or any Appendices.
+
+%\numberofauthors{8} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+
+\numberofauthors{4}
+
+\author{
+% You can go ahead and credit any number of authors here,
+% e.g. one 'row of three' or two rows (consisting of one row of three
+% and a second row of one, two or three).
+%
+% The command \alignauthor (no curly braces needed) should
+% precede each author name, affiliation/snail-mail address and
+% e-mail address. Additionally, tag each line of
+% affiliation/address with \affaddr, and tag the
+% e-mail address with \email.
+% Paper number
+%\alignauthor Paper  XXX
+% 1st. author
+\alignauthor Ryan C. Julian\\
+       \affaddr{Department of EECS}\\
+       \affaddr{Univ. of California, Berkeley}\\
+       \affaddr{Berkeley, CA 94720}\\
+       \email{ryanjulian@berkeley.edu}
+% 2nd. author
+\alignauthor Cameron J. Rose\\
+       \affaddr{Department of EECS}\\
+       \affaddr{Univ. of California, Berkeley}\\
+       \affaddr{Berkeley, CA 94720}\\
+       \email{c\_rose@eecs.berkeley.edu}
+% 3rd. author
+\alignauthor Humphrey Hu\\
+       \affaddr{Robotics Institute}\\
+       \affaddr{Carnegie Mellon University}\\
+       \affaddr{Pittsburgh, PA 15213}\\
+       \email{humhu@cmu.edu}
+\and  % use '\and' if you need 'another row' of author names
+% 4th. author
+\alignauthor Ronald S. Fearing\\
+       \affaddr{Department of EECS}\\
+       \affaddr{Univ. of California, Berkeley}\\
+       \affaddr{Berkeley, CA 94720}\\
+       \email{ronf@eecs.berkeley.edu}
+}
+
+%% There's nothing stopping you putting the seventh, eighth, etc.
+%% author on the opening page (as the 'third row') but we ask,
+%% for aesthetic reasons that you place these 'additional authors'
+%% in the \additional authors block, viz.
+%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
+%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
+%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
+%\date{30 July 1999}
+%% Just remember to make sure that the TOTAL number of authors
+%% is the number that will appear on the first page PLUS the
+%% number that will appear in the \additionalauthors section.
+
 \maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
-%TODO(ryan)
-The power, size, and weight constraints of micro air vehicle (MAV) platforms limit their on-board sensing and computational resources.
-In this work, we demonstrate cooperative control of a 14 gram ornithopter MAV by a lightweight ground station using computer vision.
-We also introduce a new ornithopter MAV platform, the h$^2$Bird, which...
+The power, size, and weight constraints of micro air vehicles (MAVs) limit their on-board sensing and computational resources.
+Ground vehicles have less mobility than MAVs, but relaxed size constraints and typically more computing power, creating opportunities for cooperation.
+In this work, we demonstrate cooperative target-seeking between a 13 gram ornithopter MAV and a lightweight ground station using computer vision.
+
+We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird.
+It features clap and fling wings, improves upon a previous design by utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a payload of 2.8 grams.
+We augment the ornithopter's built-in gyroscope-based control with a ground station, which has power and weight requirements appropriate for deployment on a ground vehicle.
+The ground station provides heading estimates to the ornithoper by running a real-time motion tracking algorithm on a live video stream.
+This allows the ornithopter to reliably negotiate flight through narrow openings.
+
+We demonstrate the performance of the location estimation and navigation algorithms.
+We model the backwards-reachable region for successfully navigating through narrow openings.
+Finally, we verify this model through with collected in experiments with window traversal.
 \end{abstract}
+
+% Note that the category section should be completed after reference to the ACM Computing Classification Scheme available at
+% http://www.acm.org/about/class/1998/.
+
+\category{I.2.9}{Robotics}{Autonomous Vehicles}
+\category{I.2.9}{Robotics}{Distributed Artificial Intelligence}[Multiagent systems] 
+\category{I.2.9}{Robotics}{Vision and Scene Understanding}[Motion]
+
+%A category including the fourth, optional field follows...
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+
+%General terms should be selected from the following 16 terms: Algorithms, Management, Measurement, Documentation, Performance, Design, Economics, Reliability, Experimentation, Security, Human Factors, Standardization, Languages, Theory, Legal Aspects, Verification.
+
+\terms{Algorithms, Performance, Design, Experimentation}
+
+%Keywords are your own choice of terms you would like the paper to be indexed by.
+
+\keywords{AAMAS proceedings, multiagent systems, cooperative control, particle filter, micro air vehicle, ground station, visual servoing, pose estimation, flapping wing, ornithopter, biomimetic}
 
@@
 These constraints are particularly severe for flapping wing MAVs offer the promise of superior safety, efficiency, and noise profile to rotor craft MAVs, but at the expense of payload weight.
+\begin{figure}[!tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird.png}
+\caption{H$^2$Bird Ornithopter MAV}
+\label{fig:h2bird}
+\end{figure}
 
@@
 %TODO(humphrey)
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/h2bird.jpg}
-\caption{H$^2$Bird Ornithopter MAV}
-\label{fig:h2bird}
-\end{figure}
-
 The flying platform used in this experiment is the H$^2$Bird flapping wing MAV.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/h2bird_axes.jpg}
+%\includegraphics[width=\linewidth]{figures/h2bird_axes.png}
+\includegraphics[height=150pt]{figures/h2bird_axes.png}
 \caption{H$^2$Bird Ornithopter with flight control axes}
@@
 \end{equation}
-
 \begin{equation}
@@
 \section{Performance}
-%TODO(ryan)
+%TODO(cameron)
 
@@
 \subsection{Visual Servoing}
-%TODO(ryan)
+%TODO(cameron)
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/window_trials.jpg}
+\includegraphics[width=\linewidth]{figures/flight_paths.pdf}
 \caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius cone, with experimental trials overlayed.}
-\label{fig:experiment_cone}
+\label{fig:flight_paths}
 \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addtolength{\textheight}{-0cm}
-                        % This command serves to balance the column lengths
-                        % on the last page of the document manually. It shortens
-                        % the textheight of the last page by a suitable amount.
-                        % This command does not take effect until the next page
-                        % so it should come on the page before the last. Make
-                        % sure that you do not shorten the textheight too much.
 
@@
 The authors would like to thank Fernando Garcia Bermudez for his assistance with the Vicon motion capture system, Andrew Pullin for his help with robot photography, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{IEEEtranS}
-\bibliography{}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% The following two commands are all you need in the
+%% initial runs of your .tex file to
+%% produce the bibliography for the citations in your paper.
+%\bibliographystyle{abbrv}
+%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+%% You must have a proper ".bib" file
+%%  and remember to run:
+%% latex bibtex latex latex
+%% to resolve all references
+%%
+%% ACM needs 'a single self-contained file'!
+%%
+%%APPENDICES are optional
+%%\balancecolumns
+%\appendix
+%%Appendix A
+%\section{Headings in Appendices}
+%The rules about hierarchical headings discussed above for
+%the body of the article are different in the appendices.
+%In the \textbf{appendix} environment, the command
+%\textbf{section} is used to
+%indicate the start of each Appendix, with alphabetic order
+%designation (i.e. the first is A, the second B, etc.) and
+%a title (if you include one).  So, if you need
+%hierarchical structure
+%\textit{within} an Appendix, start with \textbf{subsection} as the
+%highest level. Here is an outline of the body of this
+%document in Appendix-appropriate form:
+%\subsection{Introduction}
+%\subsection{The Body of the Paper}
+%\subsubsection{Type Changes and  Special Characters}
+%\subsubsection{Math Equations}
+%\paragraph{Inline (In-text) Equations}
+%\paragraph{Display Equations}
+%\subsubsection{Citations}
+%\subsubsection{Tables}
+%\subsubsection{Figures}
+%\subsubsection{Theorem-like Constructs}
+%\subsubsection*{A Caveat for the \TeX\ Expert}
+%\subsection{Conclusions}
+%\subsection{Acknowledgments}
+%\subsection{Additional Authors}
+%This section is inserted by \LaTeX; you do not insert it.
+%You just add the names and information in the
+%\texttt{{\char'134}additionalauthors} command at the start
+%of the document.
+%\subsection{References}
+%Generated by bibtex from your ~.bib file.  Run latex,
+%then bibtex, then latex twice (to resolve references)
+%to create the ~.bbl file.  Insert that ~.bbl file into
+%the .tex source file and comment out
+%the command \texttt{{\char'134}thebibliography}.
+%% This next section command marks the start of
+%% Appendix B, and does not continue the present hierarchy
+%\section{More Help for the Hardy}
+%The aamas2013 .cls file is based on the
+%sig-alternate.cls file that is itself chock-full of succinct and
+%helpful comments.  If you consider yourself a moderately experienced
+%to expert user of \LaTeX, you may find reading it useful but please
+%remember not to change it.
+%%\balancecolumns % GM June 2007
+%% That's all folks!
 \end{document}

### 2012-10-10 manuscript.tex
New figures and some text additions

@@
 \subsection{Ornithopter Flight Control}
+\label{sec:flight_control}
 %TODO(cameron)
@@
 
+We determined the turning speed and turning radius of the H$^2$Bird experimentally using the response of the system to the input of a 90 degree turn to the right.
+The results of this experiment are depicted in Figure \ref{fig:step_response}.
+The data for the experiment was gathered from the gyro data integrated on-board the H$^2$Bird.
+We used Matlab to fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the turning system to the step response with the physical heading as the state and the heading setpoint as the input.
+We use this model in Section X to determine the backwards reachable set for navigating through the window in simulation.
+
+The step response exhibits a rise time of approximately 1.4 seconds. The H$^2$Bird flies at an average forward velocity of 1.2 m/s, so the estimated minimum turning radius of the robot is 1.07 meters.
+
 %----------------------------------------------------------------------------%
@@
 %TODO(cameron)
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/feasible_set.pdf}
+\caption{Plot of backwards reachable set for successful window traversal.}
+\label{fig:feasible_set}
+\end{figure}
 
@@
 \end{figure}
+
+We determined the backwards reachable set of initial states for sussessful window traversal both geometrically and in simulation.
+The results are shown in Figure X.
+The geometrically infeasible region is indicated in blue and was computed using the minimum turning radius and simple geometery.
+Both the regions in red and blue indicate the infeasible region determined in simulation.
+We executed the simulation on the fitted turning model described in Section \ref{sec:flight_control}, with a PID controller steering the system to the center of the window as the input.
+The infeasible region calculated in simulation is a superset of the geometrically infeasible region, and is closer to the true infeasible region of the real system.
+
+The
 

### 2012-10-10 manuscript.tex
Update H2Bird drawing, abstract

@@
 %
-% ----------------------------------------------------------------------------------------------------------------
+% ----------------------------------------------------------------------
 % This .tex file (and associated .cls ) produces:
@@
 %
-% ---------------------------------------------------------------------------------------------------------------
+% ----------------------------------------------------------------------
 % This .tex source is an example which *does* use
@@
 
-% This is the document class for full camera ready papers and extended abstracts repsectively
+% This is the document class for full camera ready papers and extended 
+% abstracts repsectively
 
@@
 
-\title{Cooperative Control for Narrow Passage Traversal with an Ornithopter Micro Air Vehicle and Lightweight Ground Station}
+\title{Cooperative Control for Narrow Passage Traversal with an 
+Ornithopter Micro Air Vehicle and Lightweight Ground Station}
 
 % AUTHORS
-
 % For initial submission, do not give author names, but the
@@
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
+% reasons) and the remaining two appear in the \additionalauthors 
+% section.
 %
@@
 \begin{abstract}
-The power, size, and weight constraints of micro air vehicles (MAVs) limit their on-board sensing and computational resources.
-Ground vehicles have less mobility than MAVs, but relaxed size constraints and typically more computing power, creating opportunities for cooperation.
-In this work, we demonstrate cooperative target-seeking between a 13 gram ornithopter MAV and a lightweight ground station using computer vision.
+The power, size, and weight constraints of micro air vehicles (MAVs) limit 
+their on-board sensing and computational resources. Ground vehicles have 
+less mobility than MAVs, but relaxed size constraints and typically more 
+computing power, creating opportunities for cooperation. In this work, we 
+demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
+and a lightweight ground station using computer vision.
 
-We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird.
-It features clap and fling wings, improves upon a previous design by utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a payload of 2.8 grams.
-We augment the ornithopter's built-in gyroscope-based control with a ground station, which has power and weight requirements appropriate for deployment on a ground vehicle.
-The ground station provides heading estimates to the ornithoper by running a real-time motion tracking algorithm on a live video stream.
-This allows the ornithopter to reliably negotiate flight through narrow openings.
+We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
+features clap and fling wings, improves upon a previous design by 
+utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
+payload of 2.8 grams. We augment the ornithopter's built-in 
+gyroscope-based control with a ground station, which has power and weight 
+requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
+station provides heading estimates to the ornithoper by running a 
+real-time motion tracking algorithm on a live video stream. This allows 
+the ornithopter to reliably negotiate flight through narrow openings.
 
-We demonstrate the performance of the location estimation and navigation algorithms.
-We model the backwards-reachable region for successfully navigating through narrow openings.
-Finally, we verify this model through with collected in experiments with window traversal.
+We demonstrate the performance of the location estimation and navigation 
+algorithms. We model the backwards-reachable region for successfully 
+navigating through narrow openings. Finally, we verify this model through 
+with collected in experiments with window traversal.
 \end{abstract}
 
-% Note that the category section should be completed after reference to the ACM Computing Classification Scheme available at
+% Note that the category section should be completed after reference to 
+% the ACM Computing Classification Scheme available at
 % http://www.acm.org/about/class/1998/.
@@
 
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+% A category including the fourth, optional field follows...
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, 
+% performance measures]
 
-%General terms should be selected from the following 16 terms: Algorithms, Management, Measurement, Documentation, Performance, Design, Economics, Reliability, Experimentation, Security, Human Factors, Standardization, Languages, Theory, Legal Aspects, Verification.
+% General terms should be selected from the following 16 terms: 
+% Algorithms, Management, Measurement, Documentation, Performance, 
+% Design, Economics, Reliability, Experimentation, Security, 
+% Human Factors, Standardization, Languages, Theory, Legal Aspects, 
+% Verification.
 
@@
 
-%Keywords are your own choice of terms you would like the paper to be indexed by.
+% Keywords are your own choice of terms you would like the paper to be indexed 
+% by.
 
-\keywords{AAMAS proceedings, multiagent systems, cooperative control, particle filter, micro air vehicle, ground station, visual servoing, pose estimation, flapping wing, ornithopter, biomimetic}
+\keywords{AAMAS proceedings, multiagent systems, cooperative control, 
+particle filter, micro air vehicle, ground station, visual servoing, pose 
+estimation, flapping wing, ornithopter, biomimetic}
 
@@
 %TODO(ryan)
-The power, size, and weight constraints of micro air vehicle (MAV) platforms significantly limit on-board sensing and computational resources, and thus the complexity of the control and state estimation algorithms which can be executed on-board.
-These constraints are particularly severe for flapping wing MAVs offer the promise of superior safety, efficiency, and noise profile to rotor craft MAVs, but at the expense of payload weight.
+The power, size, and weight constraints of micro air vehicle (MAV) 
+platforms significantly limit on-board sensing and computational 
+resources, and thus the complexity of the control and state estimation 
+algorithms which can be executed on-board. These constraints are 
+particularly severe for flapping wing MAVs offer the promise of superior 
+safety, efficiency, and noise profile to rotor craft MAVs, but at the 
+expense of payload weight.
+
 \begin{figure}[!tb]
@@
 %TODO(humphrey)
-The flying platform used in this experiment is the H$^2$Bird flapping wing MAV.
-Built around the Silverlit I-Bird RC flyer powertrain and clap-fling wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams fully loaded.
-A tail propeller and servo-controlled tail elevator provide control in the yaw and pitch axes.
-An onboard electronics package combines a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor drivers with a 90 mAh lithium polymer battery for autonomous flight.
-In routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and can operate for approximately [10] minutes.
+The flying platform used in this experiment is the H$^2$Bird flapping wing 
+MAV. Built around the Silverlit I-Bird RC flyer powertrain and clap-fling 
+wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams 
+fully loaded. A tail propeller and servo-controlled tail elevator provide 
+control in the yaw and pitch axes. An onboard electronics package combines 
+a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor 
+drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
+routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
+can operate for approximately [10] minutes.
 
@@
 
-Rudimentary attitude control for the H$^2$Bird runs on-board.
-It estimates pose by naively integrating the IMU and depending on PID for control.
-The simplistic nature of our state estimation and control is due to the stringent resource constraints applied by the platform.
+Rudimentary attitude control for the H$^2$Bird runs on-board. It estimates 
+pose by naively integrating the IMU and depending on PID for control. The 
+simplistic nature of our state estimation and control is due to the 
+stringent resource constraints applied by the platform.
 
-We use simple controllers to actuate the H$^2$Bird's wing beat frequency, tail rotor, and tail elevator, which we designed to actuate thrust, yaw, and pitch respectively.
+We use simple controllers to actuate the H$^2$Bird's wing beat frequency, 
+tail rotor, and tail elevator, which we designed to actuate thrust, yaw, 
+and pitch respectively.
 
-Due to the high pitch angles experienced by the H$^2$Bird, we use a quaternion representation of vehicle pose to avoid the numerical instabilities of Euler angle representations.
-This choice of pose representation also lends itself well to remote guidance, as relative body and world coordinate angular displacements can be represented compactly with quaternion multiplication.
+Due to the high pitch angles experienced by the H$^2$Bird, we use a 
+quaternion representation of vehicle pose to avoid the numerical 
+instabilities of Euler angle representations. This choice of pose 
+representation also lends itself well to remote guidance, as relative body 
+and world coordinate angular displacements can be represented compactly 
+with quaternion multiplication.
 
-Given a reference pose $q$, we define the error $q_e$ as the rotation required to reach the reference pose from the measured pose.
-We then linearize the error quaternion to produce body axis angular errors as as inputs to the attitude controllers.
+Given a reference pose $q$, we define the error $q_e$ as the rotation 
+required to reach the reference pose from the measured pose. We then 
+linearize the error quaternion to produce body axis angular errors as as 
+inputs to the attitude controllers.
 
-Three independent proportional-integral-derivative controllers commanding the wing thrust, elevator, and tail propeller.
-We use this independent PID controller architecture for simplicity in tuning and implementation, and because it demonstrates acceptable performance in the absence of a system model.
+Three independent proportional-integral-derivative controllers commanding 
+the wing thrust, elevator, and tail propeller. We use this independent PID 
+controller architecture for simplicity in tuning and implementation, and 
+because it demonstrates acceptable performance in the absence of a system 
+model.
 
@@
 \includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
-\caption{Frame sequence from video feed used for tracking, showing particles and state estimate}
+\caption{Frame sequence from video feed used for tracking, showing 
+particles and state estimate}
 \label{fig:pf_screencap}
@@
 
-We determined the turning speed and turning radius of the H$^2$Bird experimentally using the response of the system to the input of a 90 degree turn to the right.
-The results of this experiment are depicted in Figure \ref{fig:step_response}.
-The data for the experiment was gathered from the gyro data integrated on-board the H$^2$Bird.
-We used Matlab to fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the turning system to the step response with the physical heading as the state and the heading setpoint as the input.
-We use this model in Section X to determine the backwards reachable set for navigating through the window in simulation.
+We determined the turning speed and turning radius of the H$^2$Bird 
+experimentally using the response of the system to the input of a 90 
+degree turn to the right. The results of this experiment are depicted in 
+Figure \ref{fig:step_response}. The data for the experiment was gathered 
+from the gyro data integrated on-board the H$^2$Bird. We used Matlab to 
+fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the 
+turning system to the step response with the physical heading as the state 
+and the heading setpoint as the input. We use this model in Section X to 
+determine the backwards reachable set for navigating through the window in 
+simulation.
 
-The step response exhibits a rise time of approximately 1.4 seconds. The H$^2$Bird flies at an average forward velocity of 1.2 m/s, so the estimated minimum turning radius of the robot is 1.07 meters.
+The step response exhibits a rise time of approximately 1.4 seconds. The H
+$^2$Bird flies at an average forward velocity of 1.2 m/s, so the estimated 
+minimum turning radius of the robot is 1.07 meters.
 
@@
 \includegraphics[width=\linewidth]{figures/flight_paths.pdf}
-\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius cone, with experimental trials overlayed.}
+\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius 
+cone, with experimental trials overlayed.}
 \label{fig:flight_paths}
@@
 
-We determined the backwards reachable set of initial states for sussessful window traversal both geometrically and in simulation.
-The results are shown in Figure X.
-The geometrically infeasible region is indicated in blue and was computed using the minimum turning radius and simple geometery.
-Both the regions in red and blue indicate the infeasible region determined in simulation.
-We executed the simulation on the fitted turning model described in Section \ref{sec:flight_control}, with a PID controller steering the system to the center of the window as the input.
-The infeasible region calculated in simulation is a superset of the geometrically infeasible region, and is closer to the true infeasible region of the real system.
+We determined the backwards reachable set of initial states for sussessful 
+window traversal both geometrically and in simulation. The results are 
+shown in Figure X. The geometrically infeasible region is indicated in 
+blue and was computed using the minimum turning radius and simple 
+geometery. Both the regions in red and blue indicate the infeasible region 
+determined in simulation. We executed the simulation on the fitted turning 
+model described in Section \ref{sec:flight_control}, with a PID controller 
+steering the system to the center of the window as the input. The 
+infeasible region calculated in simulation is a superset of the 
+geometrically infeasible region, and is closer to the true infeasible 
+region of the real system.
 
@@
 \section{Acknowledgements}
-The authors would like to thank Fernando Garcia Bermudez for his assistance with the Vicon motion capture system, Andrew Pullin for his help with robot photography, and the members of the Biomimetic Millisystems Laboratory and the EECS community at the University of California, Berkeley for their advice and support.
+The authors would like to thank Fernando Garcia Bermudez for his 
+assistance with the Vicon motion capture system, Andrew Pullin for his 
+help with robot photography, and the members of the Biomimetic 
+Millisystems Laboratory and the EECS community at the University of 
+California, Berkeley for their advice and support.
 
@@
 %\bibliographystyle{abbrv}
-%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this 
+% case
 %% You must have a proper ".bib" file

### 2012-10-10 manuscript.tex
Added some text

@@
 gyroscope-based control with a ground station, which has power and weight 
-requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
+requirements appropriate for deployment on a ground vehicle. The ground 
 station provides heading estimates to the ornithoper by running a 
@@
 \category{I.2.9}{Robotics}{Autonomous Vehicles}
-\category{I.2.9}{Robotics}{Distributed Artificial Intelligence}[Multiagent systems] 
+\category{I.2.9}{Robotics}{Distributed Artificial Intelligence}[Multiagent Systems]
 \category{I.2.9}{Robotics}{Vision and Scene Understanding}[Motion]
@@
 \end{equation}
+
 \begin{equation}
@@
 We determined the turning speed and turning radius of the H$^2$Bird
-experimentally using the response of the system to the input of a 90 
-degree turn to the right. The results of this experiment are depicted in 
-Figure \ref{fig:step_response}. The data for the experiment was gathered 
-from the gyro data integrated on-board the H$^2$Bird. We used Matlab to 
+experimentally using the response of the system to the step input of a
+clockwise 90 degree turn. The results of this experiment are depicted in
+Figure~\ref{fig:step_response}. The data for the experiment was gathered from
+the integration of the H$^2$Bird's built in gyroscope data. We used Matlab to
 fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
-turning system to the step response with the physical heading as the state 
-and the heading setpoint as the input. We use this model in Section X to 
-determine the backwards reachable set for navigating through the window in 
-simulation.
+turning behavior to the step response with the physical heading as the state
+and the heading set point as the input. This model serves as a simplification
+of both the turning dynamics of the ornithopter and the on-board PID
+controller regulating the actuator inputs to reach the desired orientation. We
+use this model in Section~\ref{sec:visual_servoing} to determine the backwards
+reachable region for navigating through the window in simulation.
 
-The step response exhibits a rise time of approximately 1.4 seconds. The H
-$^2$Bird flies at an average forward velocity of 1.2 m/s, so the estimated 
-minimum turning radius of the robot is 1.07 meters.
+The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
+flies at an average forward velocity of 1.2 m/s, so the estimated minimum
+turning radius of the robot is 1.07 meters. This flight speed is slower than
+the maximum speed of the ornithopter to facilitate more reliable tracking.
 
@@
 \subsection{Visual Servoing}
+\label{sec:visual_servoing}
 %TODO(cameron)
 
-\begin{figure}[tb]
+\begin{figure*}[tb]
+\begin{minipage}[b]{0.45\linewidth}
 \centering
-\includegraphics[width=\linewidth]{figures/feasible_set.pdf}
+\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
 \caption{Plot of backwards reachable set for successful window traversal.}
 \label{fig:feasible_set}
-\end{figure}
-
-\begin{figure}[tb]
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[b]{0.45\linewidth}
 \centering
-\includegraphics[width=\linewidth]{figures/flight_paths.pdf}
+\includegraphics[width=\textwidth]{figures/flight_paths.pdf}
 \caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius 
@@
 \label{fig:flight_paths}
-\end{figure}
+\end{minipage}
+\end{figure*}
 
-We determined the backwards reachable set of initial states for sussessful 
-window traversal both geometrically and in simulation. The results are 
-shown in Figure X. The geometrically infeasible region is indicated in 
-blue and was computed using the minimum turning radius and simple 
-geometery. Both the regions in red and blue indicate the infeasible region 
-determined in simulation. We executed the simulation on the fitted turning 
-model described in Section \ref{sec:flight_control}, with a PID controller 
-steering the system to the center of the window as the input. The 
-infeasible region calculated in simulation is a superset of the 
-geometrically infeasible region, and is closer to the true infeasible 
-region of the real system.
+We determined the backwards reachable set of initial states for successful
+window traversal both geometrically and in simulation. The results are shown
+in Figure~\ref{fig:feasible_set}. The initial conditions that we chose are
+spatially gridded in the camera viewing region at 0.05 meter increments and an
+initial angular position perpendicular to the window plane. The geometrically
+infeasible region is indicated in blue and was computed using the minimum
+turning radius and simple geometry. Both the regions in red and blue indicate
+the infeasible region determined in simulation. We simulated the turning model
+described in Section~\ref{sec:flight_control}, with a PID controller steering
+the system to the center of the window as the input. The infeasible region
+calculated in simulation is a superset of the geometrically infeasible region.
+The simulation is more realistic as the rotational damping during the turn and
+the on-board PID controller cause the real system to react to inputs at a
+non-constant angular velocity.
 
-The
+To verify the feasible region experimentally, we conducted trials at initial
+distances in X meter increments from the window along the edges of the image
+frame of the camera. We started each trial with an initial heading
+perpendicular to the window plane. The results of the experiments are shown in
+Figure~\ref{fig:flight_paths}.
 

### 2012-10-10 manuscript.tex

@@
 gyroscope-based control with a ground station, which has power and weight 
-requirements appropriate for deployment on a ground vehicle. The ground 
+requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
 station provides heading estimates to the ornithoper by running a 

### 2012-10-10 manuscript.tex
Minor edits, update TODOs

@@
 % the complete copyright information is generated by the cls-files.
-
 \title{Cooperative Control for Narrow Passage Traversal with an 
@@
 The flying platform used in this experiment is the H$^2$Bird flapping wing 
-MAV. Built around the Silverlit I-Bird RC flyer powertrain and clap-fling 
-wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams 
-fully loaded. A tail propeller and servo-controlled tail elevator provide 
-control in the yaw and pitch axes. An onboard electronics package combines 
-a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor 
+MAV. Built around the Silverlit I-Bird RC flyer power train and clap-fling 
+wings, the H$^2$Bird has a wingspan of [40?] cm and has a flight weight of 
+13 grams. A tail propeller and servo-controlled tail elevator provide 
+control in the yaw and pitch axes. An on-board electronics package comprises 
+a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
 drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
 routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
-can operate for approximately [10] minutes.
+can operate for approximately 10 minutes.
+
+%TODO footnote for Silverlit I-Bird
+%TODO measure wingspan
+%TODO ground speed
 
@@
 \subsection{Ground Station Computer Vision Platform}
-%TODO(ryan)
+%TODO ryan
 
@@
 \section{System Concept and Algorithms}
-%TODO(ryan)
+%TODO ryan
 
@@
 \subsection{Ornithoper Flight Control}
-%TODO(humphrey)
 
@@
 \centering
-%\includegraphics[width=\linewidth]{figures/h2bird_axes.png}
 \includegraphics[height=150pt]{figures/h2bird_axes.png}
@@
 required to reach the reference pose from the measured pose. We then 
-linearize the error quaternion to produce body axis angular errors as as 
+linearise the error quaternion to produce body axis angular errors as as 
 inputs to the attitude controllers.
@@
 \subsection{Pose Estimation}
-%TODO(ryan)
+%TODO ryan
 
@@
 \subsection{Visual Servoing}
-%TODO(ryan)
+%TODO ryan
 
@@
 \section{Experiments}
-%TODO(ryan)
+%TODO ryan
 
@@
 \section{Performance}
-%TODO(cameron)
+%TODO cameron
 
@@
 \label{sec:flight_control}
-%TODO(cameron)
+%TODO cameron
 \begin{figure}[tb]
@@
 \label{sec:visual_servoing}
-%TODO(cameron)
+%TODO cameron
 
@@
 \section{Conclusions and Future Work}
-%TODO(ryan)
+%TODO ryan
 

### 2012-10-10 manuscript.tex
Add bibliography

@@
 and world coordinate angular displacements can be represented compactly 
-with quaternion multiplication.
+with quaternion multiplication.~\cite{bowman:reasoning}
 
@@
 %% produce the bibliography for the citations in your paper.
-%\bibliographystyle{abbrv}
-%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this 
-% case
+\bibliographystyle{abbrv}
+\bibliography{manuscript}
+
 %% You must have a proper ".bib" file
@@
 %%
-%%APPENDICES are optional
-%%\balancecolumns
-%\appendix
-%%Appendix A
-%\section{Headings in Appendices}
-%The rules about hierarchical headings discussed above for
-%the body of the article are different in the appendices.
-%In the \textbf{appendix} environment, the command
-%\textbf{section} is used to
-%indicate the start of each Appendix, with alphabetic order
-%designation (i.e. the first is A, the second B, etc.) and
-%a title (if you include one).  So, if you need
-%hierarchical structure
-%\textit{within} an Appendix, start with \textbf{subsection} as the
-%highest level. Here is an outline of the body of this
-%document in Appendix-appropriate form:
-%\subsection{Introduction}
-%\subsection{The Body of the Paper}
-%\subsubsection{Type Changes and  Special Characters}
-%\subsubsection{Math Equations}
-%\paragraph{Inline (In-text) Equations}
-%\paragraph{Display Equations}
-%\subsubsection{Citations}
-%\subsubsection{Tables}
-%\subsubsection{Figures}
-%\subsubsection{Theorem-like Constructs}
-%\subsubsection*{A Caveat for the \TeX\ Expert}
-%\subsection{Conclusions}
-%\subsection{Acknowledgments}
-%\subsection{Additional Authors}
-%This section is inserted by \LaTeX; you do not insert it.
-%You just add the names and information in the
-%\texttt{{\char'134}additionalauthors} command at the start
-%of the document.
 %\subsection{References}
@@
 %the command \texttt{{\char'134}thebibliography}.
-%% This next section command marks the start of
-%% Appendix B, and does not continue the present hierarchy
-%\section{More Help for the Hardy}
 %The aamas2013 .cls file is based on the
@@
 %%\balancecolumns % GM June 2007
-%% That's all folks!
 \end{document}

### 2012-10-10 manuscript.tex
Add vector cartoons
Add fancy PDF links

@@
 \documentclass{aamas2013}
+
+% Packages
+\usepackage{color}
+\usepackage[pdftex]{hyperref}
+
+% PDF bookmarks
+\definecolor{linkCol}{gray}{0.25}
+\hypersetup{
+    pdftitle={Cooperative Control for Narrow Passage Traversal with an 
+Ornithopter Micro Air Vehicle and Lightweight Ground Station},
+    pdfauthor={R.C. Julian, C.J. Rose, H. Hu, and R.S. Fearing},
+    pdfsubject={},
+    pdfkeywords={AAMAS proceedings, multiagent systems, cooperative control, 
+particle filter, micro air vehicle, ground station, visual servoing, pose 
+estimation, flapping wing, ornithopter, biomimetic},
+    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
+    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
+}
 
@@
 \centering
-\includegraphics[height=150pt]{figures/h2bird_axes.png}
+\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
 \caption{H$^2$Bird Ornithopter with flight control axes}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/experiment_cartoon.png}
+\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
 \caption{Conceptual sketch of experimental environment.}
-\label{fig:experiment_sketch}
+\label{fig:experiment_cartoon}
 \end{figure}
@@
 reachable region for navigating through the window in simulation.
-
+\\
 The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
@@
 non-constant angular velocity.
-
+\\
 To verify the feasible region experimentally, we conducted trials at initial
@@
 Figure~\ref{fig:flight_paths}.
-
+\\
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-10-11 manuscript.tex
Update TODO
Clean up template and comments

@@
-% This is "aamas2013 .tex" September 2012 
-% This file should be compiled with "aamas2013 .cls" 
-% This example file demonstrates the use of the 'aamas2013 .cls'
-% LaTeX2e document class file. It is for those submitting
-% articles to AAMAS 2013  conference. This file is based on
-% the sig-alternate.tex example file.
-% The 'sig-alternate.cls' file of ACM will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-% than the original style ACM style.
+% AAMAS 2013 Manuscript
 %
+% This file should be compiled with "aamas2013.cls" 
 % ----------------------------------------------------------------------
-% This .tex file (and associated .cls ) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with AAMAS data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'aamas2013 .cls' you don't have control
-% from within the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the IFAAMAS Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% These information will be overwritten by fixed AAMAS 2013  information
-% in the style files - it is NOT as you are used with ACM style files.
-%
-% ----------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
+% REMEMBER: After having produced the .bbl file,
 % and prior to final submission, you *NEED* to 'insert'
@@
 % ONE 'self-contained' source file.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% STYLE NOTES
 %
+% TENSE AND VOICE: WE WRITE ALL TEXT IN PRESENT TENSE, FIRST PERSON PLURAL ("we")
+% CONVENTIONS: flapping-wing, small-scale, lightweight, MAV
 
-% This is the document class for full camera ready papers and extended 
-% abstracts repsectively
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% SETUP
 \documentclass{aamas2013}
 
-% Packages
+%----------------------------------------------------------------------------%
+% PACKAGES
 \usepackage{color}
 \usepackage[pdftex]{hyperref}
+
+%----------------------------------------------------------------------------%
+% METADATA
+\def \papertitle{Cooperative Control for Narrow Passage Traversal with an 
+Ornithopter Micro Air Vehicle and Lightweight Ground Station}
+\def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
+particle filter, micro air vehicle, ground station, visual servoing, pose 
+estimation, flapping wing, ornithopter, biomimetic}
+\def \paperterms{Algorithms, Performance, Design, Experimentation}
+\def \paperauthors{Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing}
+
+% Set title
+\title{\papertitle}
 
@@
 \hypersetup{
-    pdftitle={Cooperative Control for Narrow Passage Traversal with an 
-Ornithopter Micro Air Vehicle and Lightweight Ground Station},
-    pdfauthor={R.C. Julian, C.J. Rose, H. Hu, and R.S. Fearing},
-    pdfsubject={},
-    pdfkeywords={AAMAS proceedings, multiagent systems, cooperative control, 
-particle filter, micro air vehicle, ground station, visual servoing, pose 
-estimation, flapping wing, ornithopter, biomimetic},
+    pdftitle={\papertitle},
+    pdfauthor={\paperauthors},
+    pdfsubject={\paperterms},
+    pdfkeywords={\paperkeywords},
     pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
@@
 
-% if you are using PDF LaTex and you cannot find a way for producing
-% letter, the following explicit settings may help
-
+%----------------------------------------------------------------------------%
+% pdfLatex clean-up 
 \pdfpagewidth=8.5truein
@@
 
-% In the original styles from ACM, you would have needed to
-% add meta-info here. This is not necessary for AAMAS 2013  as
-% the complete copyright information is generated by the cls-files.
-\title{Cooperative Control for Narrow Passage Traversal with an 
-Ornithopter Micro Air Vehicle and Lightweight Ground Station}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AUTHORS
 
-% AUTHORS
-% For initial submission, do not give author names, but the
-% tracking number, instead, as the review process is blind.
-
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-%\numberofauthors{8} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors 
-% section.
-%
+%% Paper number instead of authors
+%\numberofauthors{1}
+%\alignauthor Paper  XXX
 
 \numberofauthors{4}
-
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-% Paper number
-%\alignauthor Paper  XXX
-% 1st. author
 \alignauthor Ryan C. Julian\\
@@
        \email{ryanjulian@berkeley.edu}
-% 2nd. author
 \alignauthor Cameron J. Rose\\
@@
        \email{c\_rose@eecs.berkeley.edu}
-% 3rd. author
 \alignauthor Humphrey Hu\\
@@
        \email{humhu@cmu.edu}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
+\and
 \alignauthor Ronald S. Fearing\\
@@
 
-%% There's nothing stopping you putting the seventh, eighth, etc.
-%% author on the opening page (as the 'third row') but we ask,
-%% for aesthetic reasons that you place these 'additional authors'
-%% in the \additional authors block, viz.
-%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-%\date{30 July 1999}
-%% Just remember to make sure that the TOTAL number of authors
-%% is the number that will appear on the first page PLUS the
-%% number that will appear in the \additionalauthors section.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TITLE
 \maketitle
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABSTRACT
 \begin{abstract}
@@
 \end{abstract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ACM METADATA
 
@@
 % http://www.acm.org/about/class/1998/.
-
+% A category including the fourth, optional field follows...
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, 
+% performance measures]
 \category{I.2.9}{Robotics}{Autonomous Vehicles}
@@
 \category{I.2.9}{Robotics}{Vision and Scene Understanding}[Motion]
-
-% A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, 
-% performance measures]
 
@@
 % Verification.
-
-\terms{Algorithms, Performance, Design, Experimentation}
+\terms{\paperterms}
 
@@
 % by.
-
-\keywords{AAMAS proceedings, multiagent systems, cooperative control, 
-particle filter, micro air vehicle, ground station, visual servoing, pose 
-estimation, flapping wing, ornithopter, biomimetic}
+\keywords{\paperkeywords}
 
@@
 \section{Robotic and Vision Platforms}
-%TODO(ryan)
+%TODO ryan
 
@@
 
+%TODO trim PDF
 \begin{figure}[tb]
@@
 the maximum speed of the ornithopter to facilitate more reliable tracking.
-
 %----------------------------------------------------------------------------%
@@
 
-\begin{figure*}[tb]
+%TODO make these span the page, with identical scales/style
+\begin{figure*}[h]
+
 \begin{minipage}[b]{0.45\linewidth}
@@
 \end{minipage}
+
 \hspace{0.5cm}
@@
 \end{minipage}
+
 \end{figure*}
@@
 %TODO ryan
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%
-%% The following two commands are all you need in the
-%% initial runs of your .tex file to
-%% produce the bibliography for the citations in your paper.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% BIBLIOGRAPHY
+
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
 \bibliographystyle{abbrv}
@@
 
-%% You must have a proper ".bib" file
-%%  and remember to run:
-%% latex bibtex latex latex
-%% to resolve all references
-%%
 %% ACM needs 'a single self-contained file'!
-%%
 %\subsection{References}
-%Generated by bibtex from your ~.bib file.  Run latex,
-%then bibtex, then latex twice (to resolve references)
-%to create the ~.bbl file.  Insert that ~.bbl file into
-%the .tex source file and comment out
-%the command \texttt{{\char'134}thebibliography}.
-%The aamas2013 .cls file is based on the
-%sig-alternate.cls file that is itself chock-full of succinct and
-%helpful comments.  If you consider yourself a moderately experienced
-%to expert user of \LaTeX, you may find reading it useful but please
-%remember not to change it.
-%%\balancecolumns % GM June 2007
+%% Generated by bibtex from your ~.bib file.  Run latex,
+%% then bibtex, then latex twice (to resolve references)
+%% to create the ~.bbl file.  Insert that ~.bbl file into
+%% the .tex source file and comment out
+%% the command \texttt{{\char'134}thebibliography}.
 \end{document}

### 2012-10-11 manuscript.tex
Fixed intro type and updated ornithopter platform sections.

@@
 % AAMAS 2013 Manuscript
-%
+
 % This file should be compiled with "aamas2013.cls" 
@@
 % ONE 'self-contained' source file.
+% ----------------------------------------------------------------------
 
@@
 %
-% TENSE AND VOICE: WE WRITE ALL TEXT IN PRESENT TENSE, FIRST PERSON PLURAL ("we")
-% CONVENTIONS: flapping-wing, small-scale, lightweight, MAV
-
+% TENSE AND VOICE
+% * Actual experimental events
+% 	We wrote descriptions of actual experimental events using past tense, in 
+%	active voice, as the first person plural.
+%
+% * All other contexts
+%	We write all other text using present tense, in active voices, as the
+%	the first person plural.  
+%
+% CONVENTIONS
+% flapping-wing, small-scale, lightweight, MAV
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 algorithms which can be executed on-board. These constraints are 
-particularly severe for flapping wing MAVs offer the promise of superior 
+particularly severe for flapping wing MAVs, a class of MAVs that have inherently superior 
 safety, efficiency, and noise profile to rotor craft MAVs, but at the 
@@
 13 grams. A tail propeller and servo-controlled tail elevator provide 
-control in the yaw and pitch axes. An on-board electronics package comprises 
+control in the yaw and pitch axes. The on-board electronics package consists of 
 a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
-drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
-routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
-can operate for approximately 10 minutes.
+drivers with a 90 mAh lithium polymer battery. In routine flight, the H$^2$Bird 
+averages [1.2 m/s] and can operate for approximately 10 minutes.
 
@@
 
-Rudimentary attitude control for the H$^2$Bird runs on-board. It estimates 
-pose by naively integrating the IMU and depending on PID for control. The 
-simplistic nature of our state estimation and control is due to the 
-stringent resource constraints applied by the platform.
+Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
+acceptable tracking performance. We estimate the vehicle pose by naively
+integrating the IMU, and use PID for controlling the vehicle. The simplistic 
+nature of our approach arises from the stringent resource constraints 
+of our platform, but nonetheless is not a completely trivial implementation.
 
-We use simple controllers to actuate the H$^2$Bird's wing beat frequency, 
-tail rotor, and tail elevator, which we designed to actuate thrust, yaw, 
-and pitch respectively.
+One characteristic of the H$^2$Bird that separates it from other MAVs is the
+high pitch angles experienced in routine flight. An Euler angle 
+parameterization of the vehicle orientation would suffer from singularities
+around these high pitch angles, so we instead represent the vehicle orientation
+with quaternions. An additional benefit of the quaternion representation is 
+the ability to represent relative body and world coordinate angular displacements 
+compactly with quaternion multiplication.~\cite{bowman:reasoning}
 
-Due to the high pitch angles experienced by the H$^2$Bird, we use a 
-quaternion representation of vehicle pose to avoid the numerical 
-instabilities of Euler angle representations. This choice of pose 
-representation also lends itself well to remote guidance, as relative body 
-and world coordinate angular displacements can be represented compactly 
-with quaternion multiplication.~\cite{bowman:reasoning}
-
-Given a reference pose $q$, we define the error $q_e$ as the rotation 
+Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
 required to reach the reference pose from the measured pose. We then 
-linearise the error quaternion to produce body axis angular errors as as 
-inputs to the attitude controllers.
-
-Three independent proportional-integral-derivative controllers commanding 
-the wing thrust, elevator, and tail propeller. We use this independent PID 
-controller architecture for simplicity in tuning and implementation, and 
-because it demonstrates acceptable performance in the absence of a system 
-model.
+convert this error quaternion to an angle axis representation which is given
+as an input to the PID controller.
 
@@
 
-%% ACM needs 'a single self-contained file'!
+% ACM needs 'a single self-contained file'!
 %\subsection{References}
-%% Generated by bibtex from your ~.bib file.  Run latex,
-%% then bibtex, then latex twice (to resolve references)
-%% to create the ~.bbl file.  Insert that ~.bbl file into
-%% the .tex source file and comment out
-%% the command \texttt{{\char'134}thebibliography}.
+% Generated by bibtex from your ~.bib file.  Run latex,
+% then bibtex, then latex twice (to resolve references)
+% to create the ~.bbl file.  Insert that ~.bbl file into
+% the .tex source file and comment out
+% the command \texttt{{\char'134}thebibliography}.
 \end{document}

### 2012-10-11 manuscript.tex
Adding more path figures (not using them all, just so everyone can see) and added text

--- 2012-10-10 manuscript.tex
@@
 
-% Packages
-\usepackage{color}
-\usepackage[pdftex]{hyperref}
-
-% PDF bookmarks
-\definecolor{linkCol}{gray}{0.25}
-\hypersetup{
-    pdftitle={Cooperative Control for Narrow Passage Traversal with an 
-Ornithopter Micro Air Vehicle and Lightweight Ground Station},
-    pdfauthor={R.C. Julian, C.J. Rose, H. Hu, and R.S. Fearing},
-    pdfsubject={},
-    pdfkeywords={AAMAS proceedings, multiagent systems, cooperative control, 
-particle filter, micro air vehicle, ground station, visual servoing, pose 
-estimation, flapping wing, ornithopter, biomimetic},
-    pdfpagemode=UseOutlines, bookmarksopen, bookmarksnumbered,
-    colorlinks, linkcolor=linkCol, citecolor=linkCol, urlcolor=linkCol
-}
-
 % if you are using PDF LaTex and you cannot find a way for producing
@@
 % the complete copyright information is generated by the cls-files.
+
 \title{Cooperative Control for Narrow Passage Traversal with an 
@@
 The flying platform used in this experiment is the H$^2$Bird flapping wing 
-MAV. Built around the Silverlit I-Bird RC flyer power train and clap-fling 
-wings, the H$^2$Bird has a wingspan of [40?] cm and has a flight weight of 
-13 grams. A tail propeller and servo-controlled tail elevator provide 
-control in the yaw and pitch axes. An on-board electronics package comprises 
-a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
+MAV. Built around the Silverlit I-Bird RC flyer powertrain and clap-fling 
+wings, the H$^2$Bird has a wingspan of [40?] cm and weighs [14?] grams 
+fully loaded. A tail propeller and servo-controlled tail elevator provide 
+control in the yaw and pitch axes. An onboard electronics package combines 
+a 40 MIPS microprocessor, 6 DOF IMU, 802.15.4 radio, VGA camera, and motor 
 drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
 routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
-can operate for approximately 10 minutes.
-
-%TODO footnote for Silverlit I-Bird
-%TODO measure wingspan
-%TODO ground speed
+can operate for approximately [10] minutes.
 
@@
 \subsection{Ground Station Computer Vision Platform}
-%TODO ryan
+%TODO(ryan)
 
@@
 \section{System Concept and Algorithms}
-%TODO ryan
+%TODO(ryan)
 
@@
 \subsection{Ornithoper Flight Control}
+%TODO(humphrey)
 
@@
 \centering
-\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
+%\includegraphics[width=\linewidth]{figures/h2bird_axes.png}
+\includegraphics[height=150pt]{figures/h2bird_axes.png}
 \caption{H$^2$Bird Ornithopter with flight control axes}
@@
 and world coordinate angular displacements can be represented compactly 
-with quaternion multiplication.~\cite{bowman:reasoning}
+with quaternion multiplication.
 
@@
 required to reach the reference pose from the measured pose. We then 
-linearise the error quaternion to produce body axis angular errors as as 
+linearize the error quaternion to produce body axis angular errors as as 
 inputs to the attitude controllers.
@@
 \subsection{Pose Estimation}
-%TODO ryan
+%TODO(ryan)
 
@@
 \subsection{Visual Servoing}
-%TODO ryan
+%TODO(ryan)
 
@@
 \section{Experiments}
-%TODO ryan
+%TODO(ryan)
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
+\includegraphics[width=\linewidth]{figures/experiment_cartoon.png}
 \caption{Conceptual sketch of experimental environment.}
-\label{fig:experiment_cartoon}
+\label{fig:experiment_sketch}
 \end{figure}
@@
 \section{Performance}
-%TODO cameron
+%TODO(cameron)
+
+To judge the performance of our system described in the previous sections, we
+determined the feasible set of initial conditions that resulted in successful
+window traversal. We then conducted experiments to calculate the success rate
+both within the feasible region and on the edges of the feasible region. The
+following sections detail how we calculated the feasible and infeasible
+regions using system identification and present the results of our
+experimental trials.
 
@@
 \label{sec:flight_control}
-%TODO cameron
+%TODO(cameron)
 \begin{figure}[tb]
@@
 reachable region for navigating through the window in simulation.
-\\
+
 The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
@@
 \label{sec:visual_servoing}
-%TODO cameron
+%TODO(cameron)
 
@@
 non-constant angular velocity.
-\\
+
 To verify the feasible region experimentally, we conducted trials at initial
@@
 Figure~\ref{fig:flight_paths}.
-\\
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions and Future Work}
-%TODO ryan
+%TODO(ryan)
 
@@
 %% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{manuscript}
-
+%\bibliographystyle{abbrv}
+%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this 
+% case
 %% You must have a proper ".bib" file
@@
 %%
+%%APPENDICES are optional
+%%\balancecolumns
+%\appendix
+%%Appendix A
+%\section{Headings in Appendices}
+%The rules about hierarchical headings discussed above for
+%the body of the article are different in the appendices.
+%In the \textbf{appendix} environment, the command
+%\textbf{section} is used to
+%indicate the start of each Appendix, with alphabetic order
+%designation (i.e. the first is A, the second B, etc.) and
+%a title (if you include one).  So, if you need
+%hierarchical structure
+%\textit{within} an Appendix, start with \textbf{subsection} as the
+%highest level. Here is an outline of the body of this
+%document in Appendix-appropriate form:
+%\subsection{Introduction}
+%\subsection{The Body of the Paper}
+%\subsubsection{Type Changes and  Special Characters}
+%\subsubsection{Math Equations}
+%\paragraph{Inline (In-text) Equations}
+%\paragraph{Display Equations}
+%\subsubsection{Citations}
+%\subsubsection{Tables}
+%\subsubsection{Figures}
+%\subsubsection{Theorem-like Constructs}
+%\subsubsection*{A Caveat for the \TeX\ Expert}
+%\subsection{Conclusions}
+%\subsection{Acknowledgments}
+%\subsection{Additional Authors}
+%This section is inserted by \LaTeX; you do not insert it.
+%You just add the names and information in the
+%\texttt{{\char'134}additionalauthors} command at the start
+%of the document.
 %\subsection{References}
@@
 %the command \texttt{{\char'134}thebibliography}.
+%% This next section command marks the start of
+%% Appendix B, and does not continue the present hierarchy
+%\section{More Help for the Hardy}
 %The aamas2013 .cls file is based on the
@@
 %%\balancecolumns % GM June 2007
+%% That's all folks!
 \end{document}

### 2012-10-11 manuscript.tex
More text. Added a legend to the feasible set figure

@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
+\caption{Plot of camera field of view and experimental trials within the feasible region.}
+\label{fig:flight_paths_feasible}
+\end{figure}
+
 We determined the backwards reachable set of initial states for successful
@@
 To verify the feasible region experimentally, we conducted trials at initial
-distances in X meter increments from the window along the edges of the image
-frame of the camera. We started each trial with an initial heading
-perpendicular to the window plane. The results of the experiments are shown in
-Figure~\ref{fig:flight_paths}.
+distances in approximately 0.4 meter increments from the window along the
+edges of the image frame of the camera. We started each trial with an initial
+heading perpendicular to the window plane. The results of the experiments are
+shown in Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+
+For initial conditions in the middle of the feasible region and in front of
+the camera, we achieved a success rate of 79\%. This case is depicted in
+Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
+feasible region (Figure~\ref{flight_paths}), the success rate diminishes
+because of several factors. On the edges of the camera frame, the yaw control
+inputs are large in magnitude, although this is not necessary in all cases.
+Since the control algorithm we are using has no notion of depth, it applies
+the same control input whether the ornithopter is close to the camera, where
+small changes in heading are adequate, or far away from the camera, where
+large changes in heading are necessary. Due to this phenomenon, the controller
+often over- or under-compensates in certain regions, depending on the PID
+tuning and the distance from the camera.
 

### 2012-10-11 manuscript.tex

--- 2012-10-11 manuscript.tex
@@
 
+To judge the performance of our system described in the previous sections, we
+determined the feasible set of initial conditions that resulted in successful
+window traversal. We then conducted experiments to calculate the success rate
+both within the feasible region and on the edges of the feasible region. The
+following sections detail how we calculated the feasible and infeasible
+regions using system identification and present the results of our
+experimental trials.
+
 %----------------------------------------------------------------------------%
@@
 To verify the feasible region experimentally, we conducted trials at initial
-distances in X meter increments from the window along the edges of the image
-frame of the camera. We started each trial with an initial heading
-perpendicular to the window plane. The results of the experiments are shown in
-Figure~\ref{fig:flight_paths}.
-\\
+distances in approximately 0.4 meter increments from the window along the
+edges of the image frame of the camera. We started each trial with an initial
+heading perpendicular to the window plane. The results of the experiments are
+shown in Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+
+For initial conditions in the middle of the feasible region and in front of
+the camera, we achieved a success rate of 79\%. This case is depicted in
+Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
+feasible region (Figure~\ref{flight_paths}), the success rate diminishes
+because of several factors. On the edges of the camera frame, the yaw control
+inputs are large in magnitude, although this is not necessary in all cases.
+Since the control algorithm we are using has no notion of depth, it applies
+the same control input whether the ornithopter is close to the camera, where
+small changes in heading are adequate, or far away from the camera, where
+large changes in heading are necessary. Due to this phenomenon, the controller
+often over- or under-compensates in certain regions, depending on the PID
+tuning and the distance from the camera.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-10-11 manuscript.tex
final commit of the night

@@
 the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication.~\cite{bowman:reasoning}
+compactly with quaternion multiplication.%~\cite{bowman:reasoning}
 
@@
 Figure~\ref{fig:step_response}. The data for the experiment was gathered from
-the integration of the H$^2$Bird's built in gyroscope data. We used Matlab to
+the integration of the H$^2$Bird's on-board gyroscope data. We used Matlab to
 fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
@@
 %TODO make these span the page, with identical scales/style
-\begin{figure*}[h]
-
+\begin{figure*}[tb]
 \begin{minipage}[b]{0.45\linewidth}
@@
 \end{minipage}
-
 \hspace{0.5cm}
@@
 \end{minipage}
-
 \end{figure*}
@@
 window traversal both geometrically and in simulation. The results are shown
-in Figure~\ref{fig:feasible_set}. The initial conditions that we chose are
-spatially gridded in the camera viewing region at 0.05 meter increments and an
+in Figure~\ref{fig:feasible_set}. For the initial conditions, we used a
+spatial grid of 0.05 meter increments within the camera viewing region and an
 initial angular position perpendicular to the window plane. The geometrically
@@
 turning radius and simple geometry. Both the regions in red and blue indicate
-the infeasible region determined in simulation. We simulated the turning model
-described in Section~\ref{sec:flight_control}, with a PID controller steering
-the system to the center of the window as the input. The infeasible region
-calculated in simulation is a superset of the geometrically infeasible region.
-The simulation is more realistic as the rotational damping during the turn and
-the on-board PID controller cause the real system to react to inputs at a
-non-constant angular velocity.
+the infeasible region determined in simulation. We simulated the turning
+model, described in Section~\ref{sec:flight_control}, with a PID controller
+steering the system to the center of the window as the input. The infeasible
+region calculated in simulation is a superset of the geometrically infeasible
+region. The simulation is more realistic as the rotational damping during the
+turn and the on-board PID controller cause the real system to react to inputs
+at a non-constant angular velocity.
 \\
@@
 distances in approximately 0.4 meter increments from the window along the
-edges of the image frame of the camera. We started each trial with an initial
-heading perpendicular to the window plane. The results of the experiments are
-shown in Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
-
+edges of the camera image frame. We began each trial with an initial heading
+perpendicular to the window plane. The results of the experiments are shown in
+Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+\\
 For initial conditions in the middle of the feasible region and in front of
@@
 Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region (Figure~\ref{flight_paths}), the success rate diminishes
+feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
 because of several factors. On the edges of the camera frame, the yaw control
@@
 large changes in heading are necessary. Due to this phenomenon, the controller
-often over- or under-compensates in certain regions, depending on the PID
-tuning and the distance from the camera.
+often over- or under-compensates in certain regions, depending upon the PID
+tuning and the distance from the camera. Tracking noise can also cause the
+window navigation to fail.
 

### 2012-10-11 manuscript.tex
Minor edits

--- 2012-10-11 manuscript.tex
@@
 % * All other contexts
-%	We write all other text using present tense, in active voices, as the
-%	the first person plural.  
+%	We write all other text using present tense, in active voice, as the the 
+%   first person plural.  
 %
@@
 % flapping-wing, small-scale, lightweight, MAV
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{abstract}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Introduction}
-%TODO(ryan)
+%TODO ryan
 The power, size, and weight constraints of micro air vehicle (MAV) 
@@
 algorithms which can be executed on-board. These constraints are 
-particularly severe for flapping wing MAVs, a class of MAVs that have inherently superior 
+particularly severe for flapping wing MAVs offer the promise of superior 
 safety, efficiency, and noise profile to rotor craft MAVs, but at the 
@@
 13 grams. A tail propeller and servo-controlled tail elevator provide 
-control in the yaw and pitch axes. The on-board electronics package consists of 
+control in the yaw and pitch axes. An on-board electronics package comprises 
 a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
-drivers with a 90 mAh lithium polymer battery. In routine flight, the H$^2$Bird 
-averages [1.2 m/s] and can operate for approximately 10 minutes.
+drivers with a 90 mAh lithium polymer battery for autonomous flight. In 
+routine flight, the H$^2$Bird averages [insert speed] m/s ground speed and 
+can operate for approximately 10 minutes.
 
@@
 
-Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
-acceptable tracking performance. We estimate the vehicle pose by naively
-integrating the IMU, and use PID for controlling the vehicle. The simplistic 
-nature of our approach arises from the stringent resource constraints 
-of our platform, but nonetheless is not a completely trivial implementation.
+Rudimentary attitude control for the H$^2$Bird runs on-board. It estimates 
+pose by naively integrating the IMU and depending on PID for control. The 
+simplistic nature of our state estimation and control is due to the 
+stringent resource constraints applied by the platform.
 
-One characteristic of the H$^2$Bird that separates it from other MAVs is the
-high pitch angles experienced in routine flight. An Euler angle 
-parameterization of the vehicle orientation would suffer from singularities
-around these high pitch angles, so we instead represent the vehicle orientation
-with quaternions. An additional benefit of the quaternion representation is 
-the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication.~\cite{bowman:reasoning}
+We use simple controllers to actuate the H$^2$Bird's wing beat frequency, 
+tail rotor, and tail elevator, which we designed to actuate thrust, yaw, 
+and pitch respectively.
 
-Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
+Due to the high pitch angles experienced by the H$^2$Bird, we use a 
+quaternion representation of vehicle pose to avoid the numerical 
+instabilities of Euler angle representations. This choice of pose 
+representation also lends itself well to remote guidance, as relative body 
+and world coordinate angular displacements can be represented compactly 
+with quaternion multiplication.~\cite{bowman:reasoning}
+
+Given a reference pose $q$, we define the error $q_e$ as the rotation 
 required to reach the reference pose from the measured pose. We then 
-convert this error quaternion to an angle axis representation which is given
-as an input to the PID controller.
+linearise the error quaternion to produce body axis angular errors as as 
+inputs to the attitude controllers.
+
+Three independent proportional-integral-derivative controllers commanding 
+the wing thrust, elevator, and tail propeller. We use this independent PID 
+controller architecture for simplicity in tuning and implementation, and 
+because it demonstrates acceptable performance in the absence of a system 
+model.
 
@@
 
-To judge the performance of our system described in the previous sections, we
-determined the feasible set of initial conditions that resulted in successful
-window traversal. We then conducted experiments to calculate the success rate
-both within the feasible region and on the edges of the feasible region. The
-following sections detail how we calculated the feasible and infeasible
-regions using system identification and present the results of our
-experimental trials.
-
 %----------------------------------------------------------------------------%
@@
 the maximum speed of the ornithopter to facilitate more reliable tracking.
+\\
 %----------------------------------------------------------------------------%
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
-\caption{Plot of camera field of view and experimental trials within the feasible region.}
-\label{fig:flight_paths_feasible}
-\end{figure}
-
 We determined the backwards reachable set of initial states for successful
@@
 To verify the feasible region experimentally, we conducted trials at initial
-distances in approximately 0.4 meter increments from the window along the
-edges of the image frame of the camera. We started each trial with an initial
-heading perpendicular to the window plane. The results of the experiments are
-shown in Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
-
-For initial conditions in the middle of the feasible region and in front of
-the camera, we achieved a success rate of 79\%. This case is depicted in
-Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region (Figure~\ref{flight_paths}), the success rate diminishes
-because of several factors. On the edges of the camera frame, the yaw control
-inputs are large in magnitude, although this is not necessary in all cases.
-Since the control algorithm we are using has no notion of depth, it applies
-the same control input whether the ornithopter is close to the camera, where
-small changes in heading are adequate, or far away from the camera, where
-large changes in heading are necessary. Due to this phenomenon, the controller
-often over- or under-compensates in certain regions, depending on the PID
-tuning and the distance from the camera.
-
+distances in X meter increments from the window along the edges of the image
+frame of the camera. We started each trial with an initial heading
+perpendicular to the window plane. The results of the experiments are shown in
+Figure~\ref{fig:flight_paths}.
+\\
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %TODO ryan
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 California, Berkeley for their advice and support.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2012-10-11 manuscript.tex
Figure improvements and grouping for path and feasible set

--- 2012-10-11 manuscript.tex
@@
 \begin{figure*}[tb]
+
+\begin{minipage}{\linewidth}
+\centering
+\includegraphics[width=0.45\textwidth]{figures/feasible_set.pdf}
+\caption{Plot of camera field-of-view and backwards reachable set for successful window traversal.}
+\label{fig:feasible_set}
+\end{minipage}
+\vspace{1em}
 \begin{minipage}[b]{0.45\linewidth}
 \centering
-\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
-\caption{Plot of backwards reachable set for successful window traversal.}
-\label{fig:feasible_set}
+\includegraphics[width=\textwidth]{figures/flight_paths_feasible.pdf}
+\caption{Plot of camera field of view and experimental trials within the feasible region.}
+\label{fig:flight_paths_feasible}
 \end{minipage}
-\hspace{0.5cm}
+\hfill
 \begin{minipage}[b]{0.45\linewidth}
@@
 \includegraphics[width=\textwidth]{figures/flight_paths.pdf}
-\caption{Plot of camera field-of-view and H$^2$Bird minimum turning radius 
-cone, with experimental trials overlayed.}
+\caption{Plot of camera field-of-view and perimeter experimental trials overlayed.}
 \label{fig:flight_paths}
@@
 \end{figure*}
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
-\caption{Plot of camera field of view and experimental trials within the feasible region.}
-\label{fig:flight_paths_feasible}
-\end{figure}
 

### 2012-10-11 manuscript.tex
Pre-merge commit

--- 2012-10-11 manuscript.tex
@@
 Millisystems Laboratory and the EECS community at the University of 
-California, Berkeley for their advice and support.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+California, Berkeley for their advice and support.~\cite{bowman:reasoning}
 

### 2012-10-11 manuscript.tex
Minor edits, update TODOs

@@
 \section{Robotic and Vision Platforms}
-%TODO ryan
+% ryan
 
@@
 \subsection{H$^2$Bird Ornithoper MAV}
-%TODO(humphrey)
-The flying platform used in this experiment is the H$^2$Bird flapping wing 
-MAV. Built around the Silverlit I-Bird RC flyer power train and clap-fling 
-wings, the H$^2$Bird has a wingspan of [40?] cm and has a flight weight of 
-13 grams. A tail propeller and servo-controlled tail elevator provide 
-control in the yaw and pitch axes. The on-board electronics package consists of 
-a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
-drivers with a 90 mAh lithium polymer battery. In routine flight, the H$^2$Bird 
-averages [1.2 m/s] and can operate for approximately 10 minutes.
-
+% humphrey
+In order to conduct these experiments on cooperation we designed a new
+flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
+RC flyer power train and clap-fling wings, the H$^2$Bird's wingspan is 26.5 
+cm and it has a flight weight of 13 grams. A tail propeller and 
+servo-controlled tail elevator provide control in the yaw and pitch axes. The 
+on-board electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
+IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
+lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
+ground speed and can operate for approximately 10 minutes.
 %TODO footnote for Silverlit I-Bird
-%TODO measure wingspan
-%TODO ground speed
+%TODO footnote for ImageProc
 
@@
 \end{figure}
+%TODO label ibird drawing
 
@@
 acceptable tracking performance. We estimate the vehicle pose by naively
-integrating the IMU, and use PID for controlling the vehicle. The simplistic 
-nature of our approach arises from the stringent resource constraints 
-of our platform, but nonetheless is not a completely trivial implementation.
+integrating the IMU, and use PID for controlling the vehicle.
 
 One characteristic of the H$^2$Bird that separates it from other MAVs is the
-high pitch angles experienced in routine flight. An Euler angle 
-parameterization of the vehicle orientation would suffer from singularities
+high pitch angles achieved in routine flight. An Euler angle 
+parametrization of the vehicle orientation suffers from singularities
 around these high pitch angles, so we instead represent the vehicle orientation
@@
 the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication.%~\cite{bowman:reasoning}
+compactly with quaternion multiplication
 
@@
 \end{figure}
+%TODO label experiment cartoon
 
@@
 %TODO cameron
-
-To judge the performance of our system described in the previous sections, we
+To bound the performance of our system described in the previous sections, we
 determined the feasible set of initial conditions that resulted in successful
@@
 experimentally using the response of the system to the step input of a
-clockwise 90 degree turn. The results of this experiment are depicted in
-Figure~\ref{fig:step_response}. The data for the experiment was gathered from
-the integration of the H$^2$Bird's on-board gyroscope data. We used Matlab to
+clockwise 90 degree turn. We depict the results of this experiment in
+Figure~\ref{fig:step_response}. We gathered the data for the experiment by 
+integrating the output of H$^2$Bird's on-board gyroscope data. We used MATLAB to
 fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
@@
 reachable region for navigating through the window in simulation.
-\\
-The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
+
+The step response rise time is approximately 1.4 seconds. The H$^2$Bird
 flies at an average forward velocity of 1.2 m/s, so the estimated minimum
-turning radius of the robot is 1.07 meters. This flight speed is slower than
-the maximum speed of the ornithopter to facilitate more reliable tracking.
-\\
+turning radius of the robot is 1.07 meters. In experiments, we choose a flight
+speed lower thanthe maximum speed of the ornithopter, to facilitate more
+robust tracking.
+%TODO footnote for MATLAB
+
 %----------------------------------------------------------------------------%

### 2012-10-11 manuscript.tex
Integrate Ron's comments from Monday
Fixed block diagram. Started on intro.

--- 2012-10-11 manuscript.tex
@@
 % * All other contexts
-%	We write all other text using present tense, in active voice, as the the 
-%   first person plural.  
+%	We write all other text using present tense, in active voices, as the
+%	the first person plural.  
 %
@@
 % flapping-wing, small-scale, lightweight, MAV
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{abstract}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Introduction}
-%TODO ryan
+%TODO(ryan)
 The power, size, and weight constraints of micro air vehicle (MAV) 
@@
 expense of payload weight.
-
+%Intro portion about previous works in ornithopter coorperative control/
+%modeling ( won't necessarily stay in this spot)
+\\
+Autonomous control of flapping winged MAVs is a complex undertaking, because
+of the nonlinear flight dynamics and underactuation of the system due to size
+constraints. Many of the methods for controlling these platforms make
+extensive use of GPS in outdoor environments [CITATIONS]. When operating
+indoors, using GPS to estimate the pose of the MAV is not reliable. The
+periodic motion and payload limits on MAVs make on-board sensing difficult, so
+external sensing and multi-agent cooperative control schemes are apt
+solutions.
+\\
+Many of the control laws that previous researches have developed from the dynamic models of flapping-winged flight are computationally complex and require powerful on-board computing [CITATIONS].
+In our work, we develop a simpler, less computationally intensive method of control for accomplishing navigation through a narrow passage.
 \begin{figure}[!tb]
@@
 the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication.%~\cite{bowman:reasoning}
+compactly with quaternion multiplication.~\cite{bowman:reasoning}
 
@@
 the maximum speed of the ornithopter to facilitate more reliable tracking.
-\\
 %----------------------------------------------------------------------------%
@@
 \end{minipage}
-\hspace{0.5cm}
+\hfill
 \begin{minipage}[b]{0.45\linewidth}
@@
 turning radius and simple geometry. Both the regions in red and blue indicate
-the infeasible region determined in simulation. We simulated the turning
-model, described in Section~\ref{sec:flight_control}, with a PID controller
-steering the system to the center of the window as the input. The infeasible
-region calculated in simulation is a superset of the geometrically infeasible
-region. The simulation is more realistic as the rotational damping during the
-turn and the on-board PID controller cause the real system to react to inputs
-at a non-constant angular velocity.
+the infeasible region determined in simulation. We ran a simulation on a
+motion model similar to Dubin's car in ~\cite{lavalle:planning}, with the
+angular heading as the input rather than the angular velocity. We used the
+turning model, described in Section~\ref{sec:flight_control}, as the input to
+the motion model, with a PID controller steering the entire system to the
+center of the window as the input to the tuning model. The infeasible region
+calculated in simulation is a superset of the geometrically infeasible region.
+The simulation is more realistic as the rotational damping during the turn and
+the on-board PID controller cause the real system to react to inputs at a
+non-constant angular velocity.
 \\
@@
 Millisystems Laboratory and the EECS community at the University of 
-California, Berkeley for their advice and support.~\cite{bowman:reasoning}
+California, Berkeley for their advice and support.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2012-10-11 manuscript.tex

--- 2012-10-11 manuscript.tex
@@
 \section{Robotic and Vision Platforms}
-% ryan
+%TODO ryan
 
@@
 \includegraphics[width=\linewidth]{figures/process_flow.pdf}
-\caption{Process flow diagram for the cooperative control system.}
+\caption{Overview of the cooperative control system.}
 \label{fig:process_flow}
@@
 \end{figure}
+%TODO reduce number of sig figs
 

### 2012-10-11 manuscript.tex
Remove bogus citation

@@
 Millisystems Laboratory and the EECS community at the University of 
-California, Berkeley for their advice and support.~\cite{bowman:reasoning}
+California, Berkeley for their advice and support.
 

### 2012-10-12 manuscript.tex

@@
 % AAMAS 2013 Manuscript
-
+%
 % This file should be compiled with "aamas2013.cls" 
@@
 %\numberofauthors{1}
-%\alignauthor Paper  XXX
+%\alignauthor Paper  580
 
@@
 %modeling ( won't necessarily stay in this spot)
-\\
+
 Autonomous control of flapping winged MAVs is a complex undertaking, because
-of the nonlinear flight dynamics and underactuation of the system due to size
+of the non-linear flight dynamics and underactuation of the system due to size
 constraints. Many of the methods for controlling these platforms make
-extensive use of GPS in outdoor environments [CITATIONS]. When operating
-indoors, using GPS to estimate the pose of the MAV is not reliable. The
-periodic motion and payload limits on MAVs make on-board sensing difficult, so
-external sensing and multi-agent cooperative control schemes are apt
-solutions.
-\\
+extensive use of GPS in outdoor environments [CITATIONS] and extensive motion
+capture systems in indoor environments. While operating indoors, using GPS to 
+estimate the pose of the MAV is not possible, and motion capture requires expensive
+and stationary modification of the environment. The power and payload limits on MAVs
+make on-board sensing difficult, so multi-agent cooperative control schemes are 
+apt solutions.
+
 Many of the control laws that previous researches have developed from the dynamic models of flapping-winged flight are computationally complex and require powerful on-board computing [CITATIONS].
-In our work, we develop a simpler, less computationally intensive method of control for accomplishing navigation through a narrow passage.
+In our work, we develop a simpler, less computationally intensive method of control for negotiating narrow passages.
+
 \begin{figure}[!tb]
@@
 flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
-RC flyer power train and clap-fling wings, the H$^2$Bird's wingspan is 26.5 
+RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan is 26.5 
 cm and it has a flight weight of 13 grams. A tail propeller and 
@@
 IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
-lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
+lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 1.2 m/s 
 ground speed and can operate for approximately 10 minutes.
@@
 %TODO ryan
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
-\caption{Frame sequence from video feed used for tracking, showing 
-particles and state estimate}
+\caption{Frame sequence from video feed used for tracking. Particles 
+are shown in \textcolor{red}{red}, the current state estimate as a 
+\textcolor{blue}{blue} circle, the window bounding box in
+\textcolor{red}{red}, and the target location as a \textcolor{green}{green} 
+circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
+dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
+Particles converge to its location immediately; 4.~Release H$^2$Bird, which 
+begins cooperative autonomous flight toward the target; 5.~Visual servoing 
+controller overshoots the window on the left, and commands H$^2$Bird with a 
+right-turn maneuver to recover; 6.~H$^2$Bird successfully navigates through 
+the window.}
 \label{fig:pf_screencap}
@@
 turning radius of the robot is 1.07 meters. In experiments, we choose a flight
-speed lower thanthe maximum speed of the ornithopter, to facilitate more
+speed lower than the maximum speed of the ornithopter, to facilitate more
 robust tracking.
@@
 Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
-\\
+
 For initial conditions in the middle of the feasible region and in front of

### 2012-10-12 manuscript.tex
Update outline, compress files

@@
 navigating through narrow openings. Finally, we verify this model through 
-with collected in experiments with window traversal.
+with collected in experiments with narrow passage traversal.
 \end{abstract}
@@
 \section{Introduction}
-%TODO(ryan)
+
+\begin{figure}[!tb]
+\centering
+\includegraphics[width=\linewidth]{figures/h2bird.png}
+\caption{H$^2$Bird Ornithopter MAV}
+\label{fig:h2bird}
+\end{figure}
+
+%TODO all
 The power, size, and weight constraints of micro air vehicle (MAV) 
@@
 
-\begin{figure}[!tb]
-\centering
-\includegraphics[width=\linewidth]{figures/h2bird.png}
-\caption{H$^2$Bird Ornithopter MAV}
-\label{fig:h2bird}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
-RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan is 26.5 
-cm and it has a flight weight of 13 grams. A tail propeller and 
+RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan
+is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
 servo-controlled tail elevator provide control in the yaw and pitch axes. The 
@@
 IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
-lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 1.2 m/s 
-ground speed and can operate for approximately 10 minutes.
+lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 
+a ground speed of 1.2 m/s and operates for approximately 10 minutes.
 %TODO footnote for Silverlit I-Bird
@@
 %----------------------------------------------------------------------------%
-\subsection{Ornithoper Flight Control}
+\subsection{Ornithoper Attitude Control}
 
-\begin{figure}[tb]
+\begin{figure}[!tb]
 \centering
 \includegraphics[height=150pt]{figures/h2bird_axes.pdf}
-\caption{H$^2$Bird Ornithopter with flight control axes}
+\caption{H$^2$Bird ornithopter with attitude control axes and labeled control surfaces.}
 \label{fig:h2Bird_axes}
 \end{figure}
-%TODO label ibird drawing
+%TODO label h2bird drawing
 
@@
 with quaternions. An additional benefit of the quaternion representation is 
-the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication
+the ability to represent relative body and world coordinate angular
+displacements compactly with quaternion multiplication
 
@@
 
-%TODO trim PDF
 \begin{figure}[tb]
@@
 \includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
-\caption{Overall block diagram of cooperative control}
+\caption{Overall block diagram of the cooperative control system.}
 \label{fig:block_diagram}
@@
 dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
-Particles converge to its location immediately; 4.~Release H$^2$Bird, which 
+Particles converge on its location; 4.~Release H$^2$Bird, which 
 begins cooperative autonomous flight toward the target; 5.~Visual servoing 
-controller overshoots the window on the left, and commands H$^2$Bird with a 
-right-turn maneuver to recover; 6.~H$^2$Bird successfully navigates through 
-the window.}
+controller overshoots the window on the left, and commands a right-turn 
+maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
 \label{fig:pf_screencap}
@@
 \section{Experiments}
-%TODO ryan
+%TODO humphrey
 
-\begin{figure}[tb]
+\begin{figure}[!tb]
 \centering
@@
 %TODO cameron
-To bound the performance of our system described in the previous sections, we
-determined the feasible set of initial conditions that resulted in successful
-window traversal. We then conducted experiments to calculate the success rate
-both within the feasible region and on the edges of the feasible region. The
-following sections detail how we calculated the feasible and infeasible
+To bound the performance of the cooperative control system, we determine 
+the feasible set of initial conditions that result in successful narrow 
+passage traversal using both analytical and empirical methods. We conducted 
+experiments to validate this model.
+
+The following sections detail how we calculated the feasible and infeasible
 regions using system identification and present the results of our
@@
 %----------------------------------------------------------------------------%
-\subsection{Ornithopter Flight Control}
+\subsection{Ornithopter Attitude Control}
 \label{sec:flight_control}
@@
 clockwise 90 degree turn. We depict the results of this experiment in
-Figure~\ref{fig:step_response}. We gathered the data for the experiment by 
-integrating the output of H$^2$Bird's on-board gyroscope data. We used MATLAB to
-fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
-turning behavior to the step response with the physical heading as the state
-and the heading set point as the input. This model serves as a simplification
-of both the turning dynamics of the ornithopter and the on-board PID
-controller regulating the actuator inputs to reach the desired orientation. We
-use this model in Section~\ref{sec:visual_servoing} to determine the backwards
-reachable region for navigating through the window in simulation.
+Figure~\ref{fig:step_response}.
+
+We gathered the data for the experiment by integrating the output of 
+H$^2$Bird's on-board gyroscope data. We then used MATLAB [FOOTNOTE] to fit a 
+simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the turning 
+behavior to the step response with the physical heading as the state and the 
+heading set point as the input. This model serves as a simplification of 
+both the turning dynamics of the ornithopter and the on-board PID-based 
+attitude controller. We use this model in Section~\ref{sec:visual_servoing} 
+to determine the backwards reachable region for navigating through narrow passages.
+%TODO H(s) in a labeled equation
+%TODO footnote for MATLAB
 
@@
 robust tracking.
+%TODO equation(s) for how you arrive at the minimum turning radius
 %TODO footnote for MATLAB
@@
 \subsection{Visual Servoing}
-\label{sec:visual_servoing}
 %TODO cameron
 
+\subsubsection{Reachability Model}
+\label{sec:visual_servoing}
 %TODO make these span the page, with identical scales/style
-\begin{figure*}[tb]
+\begin{figure*}[tb!]
 
@@
 
-We determined the backwards reachable set of initial states for successful
-window traversal both geometrically and in simulation. The results are shown
-in Figure~\ref{fig:feasible_set}. For the initial conditions, we used a
-spatial grid of 0.05 meter increments within the camera viewing region and an
-initial angular position perpendicular to the window plane. The geometrically
-infeasible region is indicated in blue and was computed using the minimum
-turning radius and simple geometry. Both the regions in red and blue indicate
-the infeasible region determined in simulation. We ran a simulation on a
-motion model similar to Dubin's car in ~\cite{lavalle:planning}, with the
+We determine the backwards reachable set of initial states for successful
+narrow passage traversal both geometrically and in simulation 
+(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform 
+grid of 0.05 meter increments within the camera viewing region, and an 
+initial angular position perpendicular to the narrow passage plane.
+
+The geometrically infeasible region is computed using the minimum
+turning radius and the geometry of the camera's field-of-view. Both the 
+regions in red and blue indicate the infeasible region determined in simulation. 
+We ran a simulation on a motion model similar to Dubin's car in ~\cite{lavalle:planning}, with the
 angular heading as the input rather than the angular velocity. We used the
@@
 the motion model, with a PID controller steering the entire system to the
-center of the window as the input to the tuning model. The infeasible region
+center of the passage as the input to the tuning model. The infeasible region
 calculated in simulation is a superset of the geometrically infeasible region.
@@
 
+\subsubsection{Experimental Validation}
 To verify the feasible region experimentally, we conducted trials at initial
-distances in approximately 0.4 meter increments from the window along the
+distances in approximately 0.4 meter increments from a window along the
 edges of the camera image frame. We began each trial with an initial heading
@@
 often over- or under-compensates in certain regions, depending upon the PID
-tuning and the distance from the camera. Tracking noise can also cause the
-window navigation to fail.
+tuning and the distance from the camera. Tracking noise can also cause passage
+negotiation to fail.
 

### 2012-10-12 manuscript.tex
Added platform footnotes, ornithopter section references,  and experiment section.

--- 2012-10-12 manuscript.tex
@@
 flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
-RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan is 26.5 
-cm and it has a flight weight of 13 grams. A tail propeller and 
+RC flyer power train and clap-fling wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
+the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
 servo-controlled tail elevator provide control in the yaw and pitch axes. The 
-on-board electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
-IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
-lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 1.2 m/s 
-ground speed and can operate for approximately 10 minutes.
+on-board ImageProc 2.4 electronics package consists of a 40 MIPS microprocessor, 
+6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by 
+a 90 mAh lithium polymer battery\footnote{Expand! Accredit it to Stan Baek!}. In routine flight, the H$^2$Bird 
+averages 1.2 m/s ground speed and can operate for approximately 10 minutes.
 %TODO footnote for Silverlit I-Bird
@@
 \end{figure}
-%TODO label ibird drawing
+%TODO check ibird labels
 
@@
 integrating the IMU, and use PID for controlling the vehicle.
-
-One characteristic of the H$^2$Bird that separates it from other MAVs is the
-high pitch angles achieved in routine flight. An Euler angle 
-parametrization of the vehicle orientation suffers from singularities
-around these high pitch angles, so we instead represent the vehicle orientation
+%TODO Cite Stan paper
+One behavior of the H$^2$Bird that separates it from other MAVs is the
+high pitch angles achieved in routine flight. Baek\cite{baek:tracking} 
+demonstrated control on a similar vehicle using PID and an Euler angle
+parameterization of orientation. However, Euler angles suffer from singularities
+around high pitch angles, so we instead represent the vehicle orientation
 with quaternions. An additional benefit of the quaternion representation is 
@@
 
+Our implementation of quaternion-based control is similar to Knoebe's\cite{knoebe:quatcontrol}: 
 Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
@@
 %TODO ryan
+The task we chose to test our cooperative visual servoing system with is a 
+simple target seeking one: the H$^2$Bird must fly through a specified obstacle,
+in this case, a wooden frame meant to simulate a window. As the H$^2$Bird does
+cannot detect the target on its own, it depends on remote guidance from the
+base station. Ground truth data was collected via a motion capture system in
+the test room. A sketch of the experimental setup is shown below in Figure 
+~\ref{fig:experiment_cartoon}.
+
+% TODO clarify or roll into paragraph
+Trials were run as follows:
+\begin{enumerate}
+\item The window is identified by a user. 
+\item The H$^2$Bird is released in view of the camera by hand.
+\item The H$^2$Bird attempts to fly to the window with remote guidance from
+the base station.
+\end{enumerate}
+
+% TODO Fill in test # per battery
+Variations in the H$^2$Bird hand launch cause the system to respond differently 
+for otherwise identical initial initial conditions. In addition, the H$^2$Bird 
+battery was changed approximately every [10] tests, adding more variation to
+the H$^2$Bird performance by affecting the motor power output. The effects of 
+these noisy factors is considered in Section ~\ref{sec:performance}.
 
@@
 \section{Performance}
+\label{sec:performance}
 %TODO cameron

### 2012-10-12 manuscript.tex
Added intro citations, updated platform and experiment sections.

@@
 gyroscope-based control with a ground station, which has power and weight 
-requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
+requirements appropriate for deployment on a 30 gram ground vehicle. The ground 
 station provides heading estimates to the ornithoper by running a 
@@
 expense of payload weight.
-%Intro portion about previous works in ornithopter coorperative control/
+%Intro portion about previous works in ornithopter cooperative control/
 %modeling ( won't necessarily stay in this spot)
+Modeling and control of flapping wing MAVs is a well studied subject. 
+Shigeoka discusses velocity and altitude control in \cite{Shigeoka:ornithopter}. In 
+\cite{delfly:design} deCroon et al.
 
-Autonomous control of flapping winged MAVs is a complex undertaking, because
+Due to restrictive payload weights, much previous work on guidance of flapping wing 
+with cameras depends on powerful offboard processing. Baek demonstrated altitude control 
+with an external camera\cite{baek:altitude}. de Croon et al. demonstrated obstacle 
+avoidance with an onboard camera\cite{delfly:avoid}. [However, these papers have not studied
+the system blah blah blah.]
+
+Autonomous control of flapping winged MAVs is a complex undertaking because
 of the non-linear flight dynamics and underactuation of the system due to size
@@
 % TODO Fill in test # per battery
-Variations in the H$^2$Bird hand launch cause the system to respond differently 
-for otherwise identical initial initial conditions. In addition, the H$^2$Bird 
+In total, [50+] good trials were run over a variety of starting positions. 
+Variations in the H$^2$Bird hand launch caused the system to respond very 
+differently for similar initial conditions. In addition, the H$^2$Bird 
 battery was changed approximately every [10] tests, adding more variation to
@@
 \includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
-\caption{Conceptual sketch of experimental environment.}
+\caption{Conceptual sketch of experimental environment. Dimensions are in meters.}
 \label{fig:experiment_cartoon}

### 2012-10-12 manuscript.tex
Intro update.

@@
 
-Due to restrictive payload weights, much previous work on guidance of flapping wing 
-with cameras depends on powerful offboard processing. Baek demonstrated altitude control 
-with an external camera\cite{baek:altitude}. de Croon et al. demonstrated obstacle 
-avoidance with an onboard camera\cite{delfly:avoid}. [However, these papers have not studied
-the system blah blah blah.]
+In contrast, there is little previous work on guidance of flapping wing MAVs. Baek demonstrated 
+altitude control with an external camera\cite{baek:altitude}. de Croon et al. demonstrated obstacle 
+avoidance with an onboard camera and offboard processing\cite{delfly:avoid}. [However, these papers 
+do not study the system blah blah blah.]
 

### 2012-10-12 manuscript.tex
Filled in section on BeagleBoard, minor edits

--- 2012-10-12 manuscript.tex
@@
 % humphrey
-In order to conduct these experiments on cooperation we designed a new
+In order to conduct these experiments on cooperation, we designed a new
 flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
-RC flyer power train and clap-fling wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
-the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
+RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan
+is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
 servo-controlled tail elevator provide control in the yaw and pitch axes. The 
-on-board ImageProc 2.4 electronics package consists of a 40 MIPS microprocessor, 
-6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by 
-a 90 mAh lithium polymer battery\footnote{Expand! Accredit it to Stan Baek!}. In routine flight, the H$^2$Bird 
-averages 1.2 m/s ground speed and can operate for approximately 10 minutes.
+on-board electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
+IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
+lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 
+a ground speed of 1.2 m/s and operates for approximately 10 minutes.
 %TODO footnote for Silverlit I-Bird
@@
 \subsection{Ground Station Computer Vision Platform}
-%TODO ryan
+% ryan
+Cooperative behavior is a promising strategy for autonomous navigation of 
+unexplored and unstructured environments. We intend that this system will be 
+fully mobile in these environments, not tied to a static ground station or 
+motion capture laboratory. To demonstrate the feasibility of this approach, 
+we designed our ground station to conform to the power and weight 
+constraints of highly mobile millirobotic ground vehicles, such as 
+OctoRoACH, which can traverse rough terrains with speed and agility [CITATION]. 
+%TODO OctoRoACH citation
 
+Rather than power-intensive PC processors, we used the ARM-based 
+BeagleBoard single-board computer as reference computation platform. The
+BeagleBoard consumes approximately 1 watt while running our computer vision 
+and control algorithms. It uses the same processor as popular sub-10 gram 
+processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
+light enough for deployment on millirobotic ground vehicles. We pair the 
+BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
+quality and resolution to the miniature camera modules available at 
+millirobot scale. The BeagleBoard communicates with the H$^2$Bird using a 
+USB radio module which implements the IEEE 802.15.4 wireless standard.
+%TODO footnotes: Gumstix, LogidPD
+
+Our ground station software is based on the Ubuntu Linux distribution, 
+modified with a custom version of the Linux kernel. We use the Python 
+programming language for control, robot communication, and telemetry, and 
+the OpenCV computer vision libraries for image capture and processing.
+%TODO footnotes: Ubuntu, Linux, Python, OpenCV
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{figure}
+%TODO h2bird label in diagram should have superscript
 
 %----------------------------------------------------------------------------%
-\subsection{Ornithoper Flight Control}
+\subsection{Ornithoper Attitude Control}
 
-\begin{figure}[tb]
+\begin{figure}[!tb]
 \centering
 \includegraphics[height=150pt]{figures/h2bird_axes.pdf}
-\caption{H$^2$Bird Ornithopter with flight control axes}
+\caption{H$^2$Bird ornithopter with attitude control axes and labeled control surfaces.}
 \label{fig:h2Bird_axes}
 \end{figure}
-%TODO check ibird labels
+%TODO label h2bird drawing
 
 Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
-acceptable tracking performance. We estimate the vehicle pose by naively
+maintain high-rate attitude control. We estimate the vehicle pose by naively
 integrating the IMU, and use PID for controlling the vehicle.
-%TODO Cite Stan paper
-One behavior of the H$^2$Bird that separates it from other MAVs is the
-high pitch angles achieved in routine flight. Baek\cite{baek:tracking} 
-demonstrated control on a similar vehicle using PID and an Euler angle
-parameterization of orientation. However, Euler angles suffer from singularities
-around high pitch angles, so we instead represent the vehicle orientation
-with quaternions. An additional benefit of the quaternion representation is 
-the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication
 
-Our implementation of quaternion-based control is similar to Knoebe's\cite{knoebe:quatcontrol}: 
+The high pitch angles H$^2$Bird achieves in routine flight distinguishes it 
+from other MAVs. Euler angle parametrizations of the vehicle orientation 
+suffer from singularities around these high pitch angles, so we instead 
+represent the vehicle orientation with quaternions. An additional benefit of 
+the quaternion representation its ability to represent relative body and 
+world coordinate angular displacements compactly with quaternion 
+multiplication.
+
 Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
-required to reach the reference pose from the measured pose. We then 
-convert this error quaternion to an angle axis representation which is given
-as an input to the PID controller.
-
+required to reach the reference pose from the measured pose (Eq.~\ref{eq:quat_error}). 
 \begin{equation}
-\label{quat_error}
+\label{eq:quat_error}
 q_e = q'\otimes q_r
 \end{equation}
+We then convert this error quaternion to an angle axis, and use this 
+representation as input to the PID controller 
+(Eqs.~\ref{eq:quat_angle} and~\ref{eq:quat_linearize}). 
 \begin{equation}
-\label{quat_angle}
-\alpha = 2acos(q_{e,w})
+\label{eq:quat_angle}
+\alpha = 2\arccos(q_{e,w})
 \end{equation}
 \begin{equation}
-\label{quat_linearize}
-Q_e = \alpha*q_e/sin(\alpha /2)
+\label{eq:quat_linearize}
+Q_e = \alpha*q_e/\sin(\alpha /2)
 \end{equation}
@@
 \section{Experiments}
-%TODO ryan
-The task we chose to test our cooperative visual servoing system with is a 
-simple target seeking one: the H$^2$Bird must fly through a specified obstacle,
-in this case, a wooden frame meant to simulate a window. As the H$^2$Bird does
-cannot detect the target on its own, it depends on remote guidance from the
-base station. Ground truth data was collected via a motion capture system in
-the test room. A sketch of the experimental setup is shown below in Figure 
-~\ref{fig:experiment_cartoon}.
+%TODO humphrey
 
-% TODO clarify or roll into paragraph
-Trials were run as follows:
-\begin{enumerate}
-\item The window is identified by a user. 
-\item The H$^2$Bird is released in view of the camera by hand.
-\item The H$^2$Bird attempts to fly to the window with remote guidance from
-the base station.
-\end{enumerate}
-
-% TODO Fill in test # per battery
-Variations in the H$^2$Bird hand launch cause the system to respond differently 
-for otherwise identical initial initial conditions. In addition, the H$^2$Bird 
-battery was changed approximately every [10] tests, adding more variation to
-the H$^2$Bird performance by affecting the motor power output. The effects of 
-these noisy factors is considered in Section ~\ref{sec:performance}.
-
-\begin{figure}[tb]
+\begin{figure}[!tb]
 \centering
@@
 \section{Performance}
-\label{sec:performance}
 %TODO cameron
-To bound the performance of our system described in the previous sections, we
-determined the feasible set of initial conditions that resulted in successful
-window traversal. We then conducted experiments to calculate the success rate
-both within the feasible region and on the edges of the feasible region. The
-following sections detail how we calculated the feasible and infeasible
+To bound the performance of the cooperative control system, we determine 
+the feasible set of initial conditions that result in successful narrow 
+passage traversal using both analytical and empirical methods. We conducted 
+experiments to validate this model.
+
+The following sections detail how we calculated the feasible and infeasible
 regions using system identification and present the results of our

### 2012-10-12 manuscript.tex
better figures

--- 2012-10-12 manuscript.tex
@@
 %\numberofauthors{1}
-%\alignauthor Paper  580
+%\alignauthor Paper  XXX
 
@@
 gyroscope-based control with a ground station, which has power and weight 
-requirements appropriate for deployment on a 30 gram ground vehicle. The ground 
+requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
 station provides heading estimates to the ornithoper by running a 
@@
 expense of payload weight.
-%Intro portion about previous works in ornithopter cooperative control/
+%Intro portion about previous works in ornithopter coorperative control/
 %modeling ( won't necessarily stay in this spot)
-Modeling and control of flapping wing MAVs is a well studied subject. 
-Shigeoka discusses velocity and altitude control in \cite{Shigeoka:ornithopter}. In 
-\cite{delfly:design} deCroon et al.
-
-In contrast, there is little previous work on guidance of flapping wing MAVs. Baek demonstrated 
-altitude control with an external camera\cite{baek:altitude}. de Croon et al. demonstrated obstacle 
-avoidance with an onboard camera and offboard processing\cite{delfly:avoid}. [However, these papers 
-do not study the system blah blah blah.]
-
-Autonomous control of flapping winged MAVs is a complex undertaking because
-of the non-linear flight dynamics and underactuation of the system due to size
+\\
+Autonomous control of flapping winged MAVs is a complex undertaking, because
+of the nonlinear flight dynamics and underactuation of the system due to size
 constraints. Many of the methods for controlling these platforms make
-extensive use of GPS in outdoor environments [CITATIONS] and extensive motion
-capture systems in indoor environments. While operating indoors, using GPS to 
-estimate the pose of the MAV is not possible, and motion capture requires expensive
-and stationary modification of the environment. The power and payload limits on MAVs
-make on-board sensing difficult, so multi-agent cooperative control schemes are 
-apt solutions.
-
+extensive use of GPS in outdoor environments [CITATIONS]. When operating
+indoors, using GPS to estimate the pose of the MAV is not reliable. The
+periodic motion and payload limits on MAVs make on-board sensing difficult, so
+external sensing and multi-agent cooperative control schemes are apt
+solutions.
+\\
 Many of the control laws that previous researches have developed from the dynamic models of flapping-winged flight are computationally complex and require powerful on-board computing [CITATIONS].
-In our work, we develop a simpler, less computationally intensive method of control for negotiating narrow passages.
-
+In our work, we develop a simpler, less computationally intensive method of control for accomplishing short-range tactical navigation.
 \begin{figure}[!tb]
@@
 \subsection{H$^2$Bird Ornithoper MAV}
-% humphrey
-In order to conduct these experiments on cooperation we designed a new
-flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
-RC flyer power train and clap-fling wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
-the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
-servo-controlled tail elevator provide control in the yaw and pitch axes. The 
-on-board ImageProc 2.4 electronics package consists of a 40 MIPS microprocessor, 
-6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by 
-a 90 mAh lithium polymer battery\footnote{Expand! Accredit it to Stan Baek!}. In routine flight, the H$^2$Bird 
-averages 1.2 m/s ground speed and can operate for approximately 10 minutes.
+%TODO(humphrey)
+The flying platform used in this experiment is the H$^2$Bird flapping wing 
+MAV. Built around the Silverlit I-Bird RC flyer power train and clap-fling 
+wings, the H$^2$Bird has a wingspan of [40?] cm and has a flight weight of 
+13 grams. A tail propeller and servo-controlled tail elevator provide 
+control in the yaw and pitch axes. The on-board electronics package consists of 
+a 40 MIPS microprocessor, 6 DOF IMU, IEEE 802.15.4 radio, VGA camera, and motor 
+drivers with a 90 mAh lithium polymer battery. In routine flight, the H$^2$Bird 
+averages [1.2 m/s] and can operate for approximately 10 minutes.
+
 %TODO footnote for Silverlit I-Bird
-%TODO footnote for ImageProc
+%TODO measure wingspan
+%TODO ground speed
 
@@
 %TODO ryan
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \includegraphics[width=\linewidth]{figures/process_flow.pdf}
-\caption{Overview of the cooperative control system.}
+\caption{Process flow diagram for the cooperative control system.}
 \label{fig:process_flow}
@@
 \end{figure}
-%TODO check ibird labels
 
@@
 acceptable tracking performance. We estimate the vehicle pose by naively
-integrating the IMU, and use PID for controlling the vehicle.
-%TODO Cite Stan paper
-One behavior of the H$^2$Bird that separates it from other MAVs is the
-high pitch angles achieved in routine flight. Baek\cite{baek:tracking} 
-demonstrated control on a similar vehicle using PID and an Euler angle
-parameterization of orientation. However, Euler angles suffer from singularities
-around high pitch angles, so we instead represent the vehicle orientation
+integrating the IMU, and use PID for controlling the vehicle. The simplistic 
+nature of our approach arises from the stringent resource constraints 
+of our platform, but nonetheless is not a completely trivial implementation.
+
+One characteristic of the H$^2$Bird that separates it from other MAVs is the
+high pitch angles experienced in routine flight. An Euler angle 
+parameterization of the vehicle orientation would suffer from singularities
+around these high pitch angles, so we instead represent the vehicle orientation
 with quaternions. An additional benefit of the quaternion representation is 
 the ability to represent relative body and world coordinate angular displacements 
-compactly with quaternion multiplication
+compactly with quaternion multiplication.~\cite{bowman:reasoning}
 
-Our implementation of quaternion-based control is similar to Knoebe's\cite{knoebe:quatcontrol}: 
 Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
@@
 \includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
-\caption{Frame sequence from video feed used for tracking. Particles 
-are shown in \textcolor{red}{red}, the current state estimate as a 
-\textcolor{blue}{blue} circle, the window bounding box in
-\textcolor{red}{red}, and the target location as a \textcolor{green}{green} 
-circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
-dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
-Particles converge to its location immediately; 4.~Release H$^2$Bird, which 
-begins cooperative autonomous flight toward the target; 5.~Visual servoing 
-controller overshoots the window on the left, and commands H$^2$Bird with a 
-right-turn maneuver to recover; 6.~H$^2$Bird successfully navigates through 
-the window.}
+\caption{Frame sequence from video feed used for tracking, showing 
+particles and state estimate}
 \label{fig:pf_screencap}
@@
 %TODO ryan
-The task we chose to test our cooperative visual servoing system with is a 
-simple target seeking one: the H$^2$Bird must fly through a specified obstacle,
-in this case, a wooden frame meant to simulate a window. As the H$^2$Bird does
-cannot detect the target on its own, it depends on remote guidance from the
-base station. Ground truth data was collected via a motion capture system in
-the test room. A sketch of the experimental setup is shown below in Figure 
-~\ref{fig:experiment_cartoon}.
-
-% TODO clarify or roll into paragraph
-Trials were run as follows:
-\begin{enumerate}
-\item The window is identified by a user. 
-\item The H$^2$Bird is released in view of the camera by hand.
-\item The H$^2$Bird attempts to fly to the window with remote guidance from
-the base station.
-\end{enumerate}
-
-% TODO Fill in test # per battery
-In total, [50+] good trials were run over a variety of starting positions. 
-Variations in the H$^2$Bird hand launch caused the system to respond very 
-differently for similar initial conditions. In addition, the H$^2$Bird 
-battery was changed approximately every [10] tests, adding more variation to
-the H$^2$Bird performance by affecting the motor power output. The effects of 
-these noisy factors is considered in Section ~\ref{sec:performance}.
 
@@
 \includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
-\caption{Conceptual sketch of experimental environment. Dimensions are in meters.}
+\caption{Conceptual sketch of experimental environment.}
 \label{fig:experiment_cartoon}
 \end{figure}
-%TODO label experiment cartoon
 
@@
 \section{Performance}
-\label{sec:performance}
 %TODO cameron
-To bound the performance of our system described in the previous sections, we
+
+To judge the performance of our system described in the previous sections, we
 determined the feasible set of initial conditions that resulted in successful
@@
 \end{figure}
-%TODO reduce number of sig figs
 
@@
 experimentally using the response of the system to the step input of a
-clockwise 90 degree turn. We depict the results of this experiment in
-Figure~\ref{fig:step_response}. We gathered the data for the experiment by 
-integrating the output of H$^2$Bird's on-board gyroscope data. We used MATLAB to
+clockwise 90 degree turn. The results of this experiment are depicted in
+Figure~\ref{fig:step_response}. The data for the experiment was gathered from
+the integration of the H$^2$Bird's on-board gyroscope data. We used Matlab to
 fit a simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the
@@
 reachable region for navigating through the window in simulation.
-
-The step response rise time is approximately 1.4 seconds. The H$^2$Bird
+\\
+The step response has a rise time of approximately 1.4 seconds. The H$^2$Bird
 flies at an average forward velocity of 1.2 m/s, so the estimated minimum
-turning radius of the robot is 1.07 meters. In experiments, we choose a flight
-speed lower than the maximum speed of the ornithopter, to facilitate more
-robust tracking.
-%TODO footnote for MATLAB
-
+turning radius of the robot is 1.07 meters. This flight speed is slower than
+the maximum speed of the ornithopter to facilitate more reliable tracking.
 %----------------------------------------------------------------------------%
@@
 \begin{figure*}[tb]
-
-\begin{minipage}{\linewidth}
-\centering
-\includegraphics[width=0.45\textwidth]{figures/feasible_set.pdf}
-\caption{Plot of camera field-of-view and backwards reachable set for successful window traversal.}
-\label{fig:feasible_set}
-\end{minipage}
-\vspace{1em}
 \begin{minipage}[b]{0.45\linewidth}
 \centering
-\includegraphics[width=\textwidth]{figures/flight_paths_feasible.pdf}
-\caption{Plot of camera field of view and experimental trials within the feasible region.}
-\label{fig:flight_paths_feasible}
+\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
+\caption{Plot of backwards reachable set for successful window traversal.}
+\label{fig:feasible_set}
 \end{minipage}
@@
 Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
-
+\\
 For initial conditions in the middle of the feasible region and in front of
@@
 California, Berkeley for their advice and support.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2012-10-12 manuscript.tex

--- 2012-10-12 manuscript.tex
@@
 expense of payload weight.
+
 %Intro portion about previous works in ornithopter coorperative control/
 %modeling ( won't necessarily stay in this spot)
+Modeling and control of flapping wing MAVs is a well studied subject. 
+Shigeoka discusses velocity and altitude control in
+~\cite{Shigeoka:ornithopter}. In~\cite{delfly:design} deCroon et al. %TODO sentence fragment
+In contrast, there is little previous work on guidance of flapping wing 
+MAVs. Baek demonstrated altitude control with an external 
+camera~\cite{baek:altitude}. de Croon et al. demonstrated obstacle avoidance 
+with an onboard camera and offboard processing~\cite{delfly:avoid}. 
+%[However, these papers do not study the system blah blah blah.]
 
-Autonomous control of flapping winged MAVs is a complex undertaking, because
+Autonomous control of flapping winged MAVs is a complex undertaking because
 of the non-linear flight dynamics and underactuation of the system due to size
@@
 
-Many of the control laws that previous researches have developed from the dynamic models of flapping-winged flight are computationally complex and require powerful on-board computing [CITATIONS].
-In our work, we develop a simpler, less computationally intensive method of control for negotiating narrow passages.
+Many of the control laws that previous researches have developed from the 
+dynamic models of flapping-winged flight are computationally complex and 
+require powerful on-board computing [CITATIONS]. In our work, we develop a 
+simpler, less computationally intensive method of control for negotiating 
+narrow passages.
 
@@
 flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
-RC flyer power train and clap-fling wings [FOOTNOTE], the H$^2$Bird's wingspan
-is 26.5 cm and it has a flight weight of 13 grams. A tail propeller and 
-servo-controlled tail elevator provide control in the yaw and pitch axes. The 
-on-board electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
+RC flyer power train and clap-fling 
+wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
+the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A 
+tail propeller and servo-controlled tail elevator provide control in the yaw 
+and pitch axes. The on-board ImageProc 2.4\footnote{Expand! Accredit it to Stan Baek!} 
+electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
 IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
-lithium polymer battery [FOOTNOTE]. In routine flight, the H$^2$Bird averages 
-a ground speed of 1.2 m/s and operates for approximately 10 minutes.
-%TODO footnote for Silverlit I-Bird
+lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
+ground speed, and can operate for approximately 10 minutes.
 %TODO footnote for ImageProc
@@
 The high pitch angles H$^2$Bird achieves in routine flight distinguishes it 
-from other MAVs. Euler angle parametrizations of the vehicle orientation 
+from other MAVs. Baek~\cite{baek:tracking} 
+demonstrated control on a similar vehicle using PID and an Euler angle
+parameterization of orientation. However, Euler angle parameterizations of the vehicle orientation 
 suffer from singularities around these high pitch angles, so we instead 
@@
 
+Our implementation of quaternion-based control is similar to Knoebe's~\cite{knoebe:quatcontrol}:
 Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
@@
 \section{Experiments}
-%TODO humphrey
+%TODO ryan
+The task we chose to test our cooperative visual servoing system with is a 
+simple target seeking one: the H$^2$Bird must fly through a specified obstacle,
+in this case, a wooden frame meant to simulate a window. As the H$^2$Bird does
+cannot detect the target on its own, it depends on remote guidance from the
+base station. Ground truth data was collected via a motion capture system in
+the test room. A sketch of the experimental setup is shown below in Fig.~\ref{fig:experiment_cartoon}.
 
-\begin{figure}[!tb]
+% TODO clarify or roll into paragraph
+Trials were run as follows:
+\begin{enumerate}
+\item The window is identified by a user. 
+\item The H$^2$Bird is released in view of the camera by hand.
+\item The H$^2$Bird attempts to fly to the window with remote guidance from
+the base station.
+\end{enumerate}
+
+% TODO Fill in test # per battery
+In total, [50+] good trials were run over a variety of starting positions. 
+Variations in the H$^2$Bird hand launch caused the system to respond very 
+differently for similar initial conditions. In addition, the H$^2$Bird 
+battery was changed approximately every [10] tests, adding more variation to
+the H$^2$Bird performance by affecting the motor power output. The effects of 
+these noisy factors is considered in Section ~\ref{sec:performance}.
+
+\begin{figure}[tb]
 \centering
@@
 \subsection{Visual Servoing}
+\label{sec:visual_servoing}
 %TODO cameron

### 2012-10-12 manuscript.tex
Put H2Bird pic back at the top

@@
 \section{Introduction}
-
+%TODO ryan
 \begin{figure}[!tb]
@@
 \includegraphics[width=\linewidth]{figures/h2bird.png}
-\caption{H$^2$Bird Ornithopter MAV}
+\caption{H$^2$Bird ornithopter MAV}
 \label{fig:h2bird}
 \end{figure}
-
-%TODO all
 The power, size, and weight constraints of micro air vehicle (MAV) 
@@
 expense of payload weight.
-
 %Intro portion about previous works in ornithopter coorperative control/
@@
 %TODO make these span the page, with identical scales/style
-\begin{figure*}[tb!]
-
-\begin{minipage}{\linewidth}
-\centering
-\includegraphics[width=0.45\textwidth]{figures/feasible_set.pdf}
-\caption{Plot of camera field-of-view and backwards reachable set for successful window traversal.}
-\label{fig:feasible_set}
-\end{minipage}
-\vspace{1em}
+\begin{figure*}[tb]
 \begin{minipage}[b]{0.45\linewidth}
 \centering
-\includegraphics[width=\textwidth]{figures/flight_paths_feasible.pdf}
-\caption{Plot of camera field of view and experimental trials within the feasible region.}
-\label{fig:flight_paths_feasible}
+\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
+\caption{Plot of backwards reachable set for successful window traversal.}
+\label{fig:feasible_set}
 \end{minipage}

### 2012-10-12 manuscript.tex
Added LaTeX fonts to the drawings.
added heat map

@@
 simpler, less computationally intensive method of control for negotiating 
-narrow passages.
+narrow passages similar to the method presented in [Velocity and Altitude Control of An Ornithopter Micro Aerial Vehicle].
 
@@
 The high pitch angles H$^2$Bird achieves in routine flight distinguishes it 
-from other MAVs. Baek~\cite{baek:tracking} 
+from other MAVs. Baek 
 demonstrated control on a similar vehicle using PID and an Euler angle
-parameterization of orientation. However, Euler angle parameterizations of the vehicle orientation 
+parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of the vehicle orientation 
 suffer from singularities around these high pitch angles, so we instead 
@@
 %TODO make these span the page, with identical scales/style
+
+\begin{figure*}[tb]
+\begin{minipage}[b]{0.45\linewidth}
+\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
+\caption{Plot of backwards reachable set for successful window traversal.}
+\label{fig:feasible_set}
+\end{minipage}
+\hfill
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\centering
+\includegraphics[width=\textwidth]{figures/heat_map.pdf}
+\caption{CAPTION.}
+\label{fig:heat_map}
+\end{minipage}
+\end{figure*}
+
 \begin{figure*}[tb]
@@
 \centering
-\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
-\caption{Plot of backwards reachable set for successful window traversal.}
-\label{fig:feasible_set}
+\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
+\caption{Plot of camera field of view and experimental trials within the feasible region.}
+\label{fig:flight_paths_feasible}
 \end{minipage}

### 2012-10-12 manuscript.tex
Added Probabalistic Robotics citation
Add caption to Monte Carlo
citations and updated block diagram

@@
 
-Autonomous control of flapping winged MAVs is a complex undertaking because
-of the non-linear flight dynamics and underactuation of the system due to size
-constraints. Many of the methods for controlling these platforms make
-extensive use of GPS in outdoor environments [CITATIONS] and extensive motion
-capture systems in indoor environments. While operating indoors, using GPS to 
-estimate the pose of the MAV is not possible, and motion capture requires expensive
-and stationary modification of the environment. The power and payload limits on MAVs
-make on-board sensing difficult, so multi-agent cooperative control schemes are 
-apt solutions.
+Autonomous control of flapping winged MAVs is a complex undertaking because of
+the non-linear flight dynamics as presented
+in~\cite{humbert:dipteran2,cheng:tandrdamping}. Although the dynamics of
+flapping-winged flight can be averaged with respect to the wingbeat period as
+presented in~\cite{Schenato:flightcontrol,cheng:tandrdamping}, we develop a
+linear, one-dimensional model to predict the behavior of the system, similar
+to~\cite{Shigeoka:ornithopter}. In addition to the nonlinearities, the system
+is underactuated due to size constraints. Many of the methods for controlling
+these platforms make extensive use of GPS in outdoor environments [CITATIONS]
+and expansive motion capture systems in indoor environments. While operating
+indoors, using GPS to estimate the pose of the MAV is not possible, and motion
+capture requires expensive and stationary modification of the environment. The
+power and payload limits on MAVs make on-board sensing difficult, so
+multi-agent cooperative control schemes are apt solutions.
 
 Many of the control laws that previous researches have developed from the
-dynamic models of flapping-winged flight are computationally complex and 
-require powerful on-board computing [CITATIONS]. In our work, we develop a 
-simpler, less computationally intensive method of control for negotiating 
-narrow passages similar to the method presented in [Velocity and Altitude Control of An Ornithopter Micro Aerial Vehicle].
+dynamic models of flapping-winged flight, such as periodic forcing and other
+control methods as describesd in~\cite{doman:dynamics,
+khan:longitudinal_control, leonard:averaging}, are computationally complex and
+require powerful on-board computing. In our work, we develop a simpler, less
+computationally intensive method of control for negotiating narrow passages.
 

### 2012-10-12 manuscript.tex
Start on Pose Estimation

--- 2012-10-12 manuscript.tex
@@
 %TODO ryan
+The ground station performs simple two-dimensional pose estimation on the H$^2$Bird using 
+a particle filter-based motion tracking algorithm. This pose estimate is 
+the feedback input to the cooperative visual servoing feedback loop (Sec. \ref{sec:visual_servoing_concept}).
+
+We limit our pose estimation approach to monocular two-dimensional object 
+tracking in pixel-space. This allows us to perform high-rate video tracking 
+using the modest computational resources available on the ground station, 
+and similarly millirobotic ground vehicles.
+
+~\cite{thrun2005probabilistic}
 
@@
 \subsection{Visual Servoing}
+\label{sec:visual_servoing_concept}
 %TODO ryan
 
-%TODO trim PDF
 \begin{figure}[tb]
@@
 \includegraphics[width=\textwidth]{figures/heat_map.pdf}
-\caption{CAPTION.}
+\caption{Probability of success for all starting locations. This graphic is 
+a result of a Monte Carlo simulation of the reachable set model, in which the 
+initial heading of the H$^2$Bird is chosen uniformly at random from $(-90,90)$.}
 \label{fig:heat_map}

### 2012-10-12 manuscript.tex
Fix bibtex errors

@@
 the non-linear flight dynamics as presented
-in~\cite{humbert:dipteran2,cheng:tandrdamping}. Although the dynamics of
+in~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Although the dynamics of
 flapping-winged flight can be averaged with respect to the wingbeat period as
-presented in~\cite{Schenato:flightcontrol,cheng:tandrdamping}, we develop a
+presented in~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, we develop a
 linear, one-dimensional model to predict the behavior of the system, similar
@@
 dynamic models of flapping-winged flight, such as periodic forcing and other
-control methods as describesd in~\cite{doman:dynamics,
-khan:longitudinal_control, leonard:averaging}, are computationally complex and
+control methods as described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
+in~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, are computationally complex and
 require powerful on-board computing. In our work, we develop a simpler, less
@@
 \subsubsection{Reachability Model}
-\label{sec:visual_servoing}
 %TODO make these span the page, with identical scales/style

### 2012-10-12 manuscript.tex
more citations

--- 2012-10-12 manuscript.tex
@@
 is underactuated due to size constraints. Many of the methods for controlling
-these platforms make extensive use of GPS in outdoor environments [CITATIONS]
-and expansive motion capture systems in indoor environments. While operating
-indoors, using GPS to estimate the pose of the MAV is not possible, and motion
-capture requires expensive and stationary modification of the environment. The
-power and payload limits on MAVs make on-board sensing difficult, so
-multi-agent cooperative control schemes are apt solutions.
+these platforms make extensive use of GPS in outdoor
+environments~\cite{kingston:timeattitude, kanade:3dvision} and expansive
+motion capture systems in indoor environments. While operating indoors, using
+GPS to estimate the pose of the MAV is not possible, and motion capture
+requires expensive and stationary modification of the environment. The power
+and payload limits on MAVs make on-board sensing difficult, so multi-agent
+cooperative control schemes are apt solutions.
 
@@
 %TODO ryan
-The ground station performs simple two-dimensional pose estimation on the H$^2$Bird using 
-a particle filter-based motion tracking algorithm. This pose estimate is 
-the feedback input to the cooperative visual servoing feedback loop (Sec. \ref{sec:visual_servoing_concept}).
-
-We limit our pose estimation approach to monocular two-dimensional object 
-tracking in pixel-space. This allows us to perform high-rate video tracking 
-using the modest computational resources available on the ground station, 
-and similarly millirobotic ground vehicles.
-
-~\cite{thrun2005probabilistic}
 
@@
 \subsection{Visual Servoing}
-\label{sec:visual_servoing_concept}
 %TODO ryan
 
+%TODO trim PDF
 \begin{figure}[tb]
@@
 We gathered the data for the experiment by integrating the output of
-H$^2$Bird's on-board gyroscope data. We then used MATLAB [FOOTNOTE] to fit a 
-simple low-order model, $H(s) = \frac{0.94782}{1+0.61998s}$, of the turning 
+H$^2$Bird's on-board gyroscope data. We then used
+MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
+simple low-order model (Eq.~\ref{eq:transfer_func}) of the turning
 behavior to the step response with the physical heading as the state and the
-heading set point as the input. This model serves as a simplification of 
-both the turning dynamics of the ornithopter and the on-board PID-based 
-attitude controller. We use this model in Section~\ref{sec:visual_servoing} 
-to determine the backwards reachable region for navigating through narrow passages.
+heading set point as the input. This model serves as a simplification of both
+the turning dynamics of the ornithopter and the on-board PID-based attitude
+controller. We use this model in Section~\ref{sec:visual_servoing} to
+determine the backwards reachable region for navigating through narrow
+passages.
+\begin{equation}
+\label{eq:transfer_func}
+H(s) = \frac{0.94782}{1+0.61998s}
+\end{equation}
 %TODO H(s) in a labeled equation
@@
 \includegraphics[width=\textwidth]{figures/heat_map.pdf}
-\caption{Probability of success for all starting locations. This graphic is 
-a result of a Monte Carlo simulation of the reachable set model, in which the 
-initial heading of the H$^2$Bird is chosen uniformly at random from $(-90,90)$.}
+\caption{CAPTION.}
 \label{fig:heat_map}

### 2012-10-12 manuscript.tex
fixed sigfigs

@@
 %----------------------------------------------------------------------------%
-\subsection{H$^2$Bird Ornithoper MAV}
+\subsection{H$^2$Bird Ornithoper}
 % humphrey
@@
 %----------------------------------------------------------------------------%
-\subsection{Pose Estimation}
+\subsection{Ground Station Pose Estimation}
 %TODO ryan
@@
 %----------------------------------------------------------------------------%
-\subsection{Visual Servoing}
+\subsection{Cooperative Visual Servoing}
 \label{sec:visual_servoing_concept}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{System Model}
+
+%----------------------------------------------------------------------------%
+\subsection{Ornithopter Attitude Control}
+
+%----------------------------------------------------------------------------%
+\subsection{Ground Station Pose Estimation}
+
+%----------------------------------------------------------------------------%
+\subsection{Cooperative Visual Servoing}
 
@@
 \section{Experiments}
-%TODO ryan
+
+%----------------------------------------------------------------------------%
+\subsection{System Identification}
+
+%............................................................................%
+\subsubsection{Ornithoper Attitude Control}
+
+%............................................................................%
+\subsubsection{Ground Station Pose Estimation}
+
+%............................................................................%
+\subsubsection{Cooperative Visual Servoing}
+
+%----------------------------------------------------------------------------%
+\subsection{Model Verification}
+
 The task we chose to test our cooperative visual servoing system with is a 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Performance}
+\section{Results and Performance}
 \label{sec:performance}
@@
 %----------------------------------------------------------------------------%
-\subsection{Ornithopter Attitude Control}
+\subsection{System Identification}
+
+%............................................................................%
+\subsubsection{Ornithoper Attitude Control}
 \label{sec:flight_control}
@@
 \label{eq:transfer_func}
-H(s) = \frac{0.94782}{1+0.61998s}
+H(s) = \frac{0.95}{1+0.62s}
 \end{equation}
@@
 
-%----------------------------------------------------------------------------%
-\subsection{Visual Servoing}
+%............................................................................%
+\subsubsection{Ground Station Pose Estimation}
+
+%............................................................................%
+\subsubsection{Cooperative Visual Servoing}
 \label{sec:visual_servoing}
@@
 
-\subsubsection{Reachability Model}
 %TODO make these span the page, with identical scales/style
@@
 
-\subsubsection{Experimental Validation}
 To verify the feasible region experimentally, we conducted trials at initial
@@
 For initial conditions in the middle of the feasible region and in front of
-the camera, we achieved a success rate of 79\%. This case is depicted in
+the camera, we achieved a success rate of 80\%. This case is depicted in
 Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
@@
 negotiation to fail.
+
+%----------------------------------------------------------------------------%
+\subsection{Model Verification}
 

### 2012-10-12 manuscript.tex
moved text around

@@
 \section{System Model}
+\label{sec:system_model}
 
@@
 \subsection{Ornithopter Attitude Control}
+
+We gathered the data for the experiment by integrating the output of
+H$^2$Bird's on-board gyroscope data. We then used
+MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
+simple low-order model (Eq.~\ref{eq:transfer_func}) of the turning
+behavior to the step response with the physical heading as the state and the
+heading set point as the input. This model serves as a simplification of both
+the turning dynamics of the ornithopter and the on-board PID-based attitude
+controller. We use this model in Section~\ref{sec:visual_servoing} to
+determine the backwards reachable region for navigating through narrow
+passages.
 
@@
 
+We determined the turning speed and turning radius of the H$^2$Bird
+experimentally using the response of the system to the step input of a
+clockwise 90 degree turn. To measure the response of the system, we recorded
+the angular position data of the ornithopter calculated from the on-board
+gyroscope data.
+
 %............................................................................%
@@
 
-The task we chose to test our cooperative visual servoing system with is a 
-simple target seeking one: the H$^2$Bird must fly through a specified obstacle,
-in this case, a wooden frame meant to simulate a window. As the H$^2$Bird does
-cannot detect the target on its own, it depends on remote guidance from the
-base station. Ground truth data was collected via a motion capture system in
-the test room. A sketch of the experimental setup is shown below in Fig.~\ref{fig:experiment_cartoon}.
+We tested our cooperative visual servoing system with a simple target seeking
+experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
+a wooden frame meant to simulate a window. As the H$^2$Bird cannot detect the
+target on its own,we used the ground station to provide remote headings for
+navigation to the window. Ground truth data was collected via a motion capture
+system in the test room. A sketch of the experimental setup is shown in
+Fig.~\ref{fig:experiment_cartoon}.
 
 % TODO clarify or roll into paragraph
-Trials were run as follows:
+Each trial included the following steps:
 \begin{enumerate}
-\item The window is identified by a user. 
-\item The H$^2$Bird is released in view of the camera by hand.
+\item The window is identified by the user. 
+\item The H$^2$Bird is released in view of the camera by hand at the desired starting grid point.
 \item The H$^2$Bird attempts to fly to the window with remote guidance from
@@
 
-% TODO Fill in test # per battery
-In total, [50+] good trials were run over a variety of starting positions. 
-Variations in the H$^2$Bird hand launch caused the system to respond very 
+In total, 80 trials were run over a variety of starting positions.
+The goal of our experiments was to verify the model described in~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid along the edges of the camera viewplane in 0.05 meter increments from the window to determine the success rate on the edges of the testing space.
+We conducted 20 additional trials directly in front of the camera to determine the success rate near the center of the testing space.
+The H$^2$Bird battery was changed approximately every 5 tests, adding more variation to the H$^2$Bird performance by affecting the motor power output.
+
+% Don't think this goes here
+Variations in the H$^2$Bird hand launch caused the system to respond 
 differently for similar initial conditions. In addition, the H$^2$Bird 
-battery was changed approximately every [10] tests, adding more variation to
+battery was changed approximately every 5 tests, adding more variation to
 the H$^2$Bird performance by affecting the motor power output. The effects of 
 these noisy factors is considered in Section ~\ref{sec:performance}.
+% Might move
 
@@
 
-We determined the turning speed and turning radius of the H$^2$Bird
-experimentally using the response of the system to the step input of a
-clockwise 90 degree turn. We depict the results of this experiment in
+We depict the results of this experiment in
 Figure~\ref{fig:step_response}.
 
-We gathered the data for the experiment by integrating the output of
-H$^2$Bird's on-board gyroscope data. We then used
-MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
-simple low-order model (Eq.~\ref{eq:transfer_func}) of the turning
-behavior to the step response with the physical heading as the state and the
-heading set point as the input. This model serves as a simplification of both
-the turning dynamics of the ornithopter and the on-board PID-based attitude
-controller. We use this model in Section~\ref{sec:visual_servoing} to
-determine the backwards reachable region for navigating through narrow
-passages.
 \begin{equation}
@@
 robust tracking.
+
 %TODO equation(s) for how you arrive at the minimum turning radius
@@
 
-We determine the backwards reachable set of initial states for successful
-narrow passage traversal both geometrically and in simulation 
-(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform 
-grid of 0.05 meter increments within the camera viewing region, and an 
-initial angular position perpendicular to the narrow passage plane.
-
 The geometrically infeasible region is computed using the minimum
@@
 the motion model, with a PID controller steering the entire system to the
-center of the passage as the input to the tuning model. The infeasible region
+center of the passage as the input to the tuning model.
+
+We determine the backwards reachable set of initial states for successful
+narrow passage traversal both geometrically and in simulation 
+(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform 
+grid of 0.05 meter increments within the camera viewing region, and an 
+initial angular position perpendicular to the narrow passage plane.
+
+The infeasible region
 calculated in simulation is a superset of the geometrically infeasible region.
@@
 
-To verify the feasible region experimentally, we conducted trials at initial
-distances in approximately 0.4 meter increments from a window along the
-edges of the camera image frame. We began each trial with an initial heading
-perpendicular to the window plane. The results of the experiments are shown in
+The results of the experiments are shown in
 Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+
+%----------------------------------------------------------------------------%
+\subsection{Model Verification}
 
@@
 negotiation to fail.
-
-%----------------------------------------------------------------------------%
-\subsection{Model Verification}
 

### 2012-10-12 manuscript.tex
Added Pose Estimation/Motion Model

@@
 \usepackage[pdftex]{hyperref}
+\usepackage{amsmath}
+
+\newlength\eqnindent
+\newlength\eqncolwid
+\newcommand\eqntext[2][0pt]{%
+  \setlength\parindent{0pt}%
+      \setlength\eqnindent{#1}\setlength\eqncolwid{\linewidth}%
+  \setlength\eqncolwid{\dimexpr\linewidth-\eqnindent}%
+  \hspace*{\eqnindent}\parbox{\eqncolwid}{#2}%
+}
 
@@
 %TODO ryan
-The ground station performs simple two-dimensional pose estimation on the H$^2$Bird using 
-a particle filter-based motion tracking algorithm. This pose estimate is 
-the feedback input to the cooperative visual servoing feedback loop (Sec. \ref{sec:visual_servoing_concept}).
+The ground station performs simple two-dimensional pose estimation on 
+H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
+estimate is the feedback input to the cooperative visual servoing feedback 
+loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
+approach to monocular two-dimensional object tracking in pixel space. This 
+allows us to perform real-time video tracking using the modest computational 
+resources available on the ground station, and similarly millirobotic ground 
+vehicles.
 
-We limit our pose estimation approach to monocular two-dimensional object 
-tracking in pixel-space. This allows us to perform high-rate video tracking 
-using the modest computational resources available on the ground station, 
-and similarly millirobotic ground vehicles.
+Our pose estimation algorithm tracks H$^2$Bird motion in the frame using a 
+particle filter~\cite[96-114]{thrun2005probabilistic}. We initialze particles 
+uniformly across the frame. To improve numerical stability, we  normalize the 
+weights of all particles on each iteration, and only resample the particle 
+population when the mean particle weight falls below a pre-determined 
+threshold.
 
-~\cite{thrun2005probabilistic}
+%............................................................................%
+\subsubsection{Motion Model}
+Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
+centered around each particle's current position, augmented with an 
+$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
+$x^{[i]}_t$ is the state of the $i^\text{th}$ particle at time $t$, $\Sigma$ 
+is a chosen covariance matrix, and $b$ is vector of the height and width of 
+the camera frame.
+\begin{equation}
+\label{eq:motion_model}
+x^{[i]}_t \sim \begin{cases}
+\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
+\mathcal{U}(0,b),& \text{with probability } \epsilon \\
+\end{cases}\\
+\end{equation}
+The experiments presented in this paper used a diagonal covariance matrix 
+with variance 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
+transition model is sufficient for tracking motion in which frame-to-frame 
+position changes are relatively small. The $\epsilon$-random uniform 
+transitions increase tracking robustness, by forcing the filter to always 
+sample all of the pixel space. This prevents overconfidence and collapse, 
+and decreases reacquisition delay.
+
+%............................................................................%
+\subsubsection{Emission Model}
 
@@
 \label{sec:visual_servoing_concept}
-%TODO ryan
 

### 2012-10-12 manuscript.tex
more text

--- 2012-10-12 manuscript.tex
@@
 \subsection{Ornithopter Attitude Control}
+\label{sec:model_attitude}
 
-We gathered the data for the experiment by integrating the output of
-H$^2$Bird's on-board gyroscope data. We then used
-MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
-simple low-order model (Eq.~\ref{eq:transfer_func}) of the turning
-behavior to the step response with the physical heading as the state and the
-heading set point as the input. This model serves as a simplification of both
-the turning dynamics of the ornithopter and the on-board PID-based attitude
-controller. We use this model in Section~\ref{sec:visual_servoing} to
-determine the backwards reachable region for navigating through narrow
-passages.
+To model the yaw motion of the ornithopter, we developed a one-dimensional model using the physical heading angle as the output and the heading angle set by the ground station as the input.
+This model accounts for both the yaw dynamics of the ornithopter and the internal PID controller on the actuator inputs for the tail rotor.
+We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experiementation.
+The format of the model we decided up on is shown in Equation~\ref{eq:plain_model}.
+
+\begin{equation}
+\label{eq:plain_model}
+H(s) = \frac{K}{1 + Rs}
+\end{equation}
+
+To model the translational position of the ornithopter, we used a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state.
+So the overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the angular setpoint provided by the ground station as the input.
+
+\begin{equation}
+\label{eq:total_model_plain}
+\begin{aligned}
+& \dot{x} = sin(\theta)\\
+& \dot{y} = cos(\theta)\\
+\end{aligned}
+\end{equation}
 
@@
 the angular position data of the ornithopter calculated from the on-board
-gyroscope data.
+gyroscope data. We then used
+MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
+simple low-order model to the step response as discussed in Section~\ref{sec:model_attitude}.
 
@@
 
-In total, 80 trials were run over a variety of starting positions.
-The goal of our experiments was to verify the model described in~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid along the edges of the camera viewplane in 0.05 meter increments from the window to determine the success rate on the edges of the testing space.
-We conducted 20 additional trials directly in front of the camera to determine the success rate near the center of the testing space.
-The H$^2$Bird battery was changed approximately every 5 tests, adding more variation to the H$^2$Bird performance by affecting the motor power output.
-
-% Don't think this goes here
-Variations in the H$^2$Bird hand launch caused the system to respond 
-differently for similar initial conditions. In addition, the H$^2$Bird 
-battery was changed approximately every 5 tests, adding more variation to
-the H$^2$Bird performance by affecting the motor power output. The effects of 
-these noisy factors is considered in Section ~\ref{sec:performance}.
-% Might move
+In total, 80 trials were run over a variety of starting positions. The goal of
+our experiments was to verify the model described in
+Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
+along the edges of the camera viewplane in 0.05 meter increments from the
+window to determine the success rate on the edges of the testing space. We
+conducted 20 additional trials directly in front of the camera to determine
+the success rate near the center of the testing space. After every 5 trials,
+we replaced the current battery with a fully charged one.
 
@@
 
-We depict the results of this experiment in
-Figure~\ref{fig:step_response}.
+The results of the step response experiment are described in
+Figure~\ref{fig:step_response}. The first-order model that we fit to the
+response is in Equation~\ref{eq:transfer_func}.
 
@@
 
-The step response rise time is approximately 1.4 seconds. The H$^2$Bird
-flies at an average forward velocity of 1.2 m/s, so the estimated minimum
-turning radius of the robot is 1.07 meters. In experiments, we choose a flight
-speed lower than the maximum speed of the ornithopter, to facilitate more
-robust tracking.
+The step response rise time is 1.4 seconds. The H$^2$Bird flies
+at an average forward velocity of 1.2 m/s during our experiments, so the
+estimated minimum turning radius of the robot is 1.07 maters, as described in
+Equation~\ref{eq:min_radius}. In the experiments, we choose a flight speed
+lower than the maximum speed of the ornithopter, to facilitate more robust
+tracking.
+
+\begin{equation}
+\label{eq:min_radius}
+\begin{aligned}
+\text{Minimum radius} & = \frac{360^{\circ}*\text{Rise time}*\text{Flight speed}}{\text{Final angle}*2*\pi}\\
+& = \frac{360^{\circ}*1.4s*1.2\frac{m}{s}}{90^{\circ}*2*\pi} = 1.07 m
+\end{aligned}
+\end{equation}
 
@@
 %TODO make these span the page, with identical scales/style
+
+\begin{figure*}[tb]
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
+\caption{Plot of camera field of view and experimental trials within the feasible region.}
+\label{fig:flight_paths_feasible}
+\end{minipage}
+\hfill
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{figures/flight_paths.pdf}
+\caption{Plot of camera field-of-view and perimeter experimental trials overlayed.}
+\label{fig:flight_paths}
+\end{minipage}
+\end{figure*}
+
+The results of our experimental trials are shown in
+Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
+initial conditions in the middle of the testing space and in front of the
+camera, we achieved a success rate of 80\%. This case is depicted in
+Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
+feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
+to 50\% because of several factors. On the edges of the camera frame, the yaw
+control inputs are large in magnitude, although this is not necessary in all
+cases. Since the control algorithm we are using has no notion of depth, it
+applies the same control input whether the ornithopter is close to the camera,
+where small changes in heading are adequate, or far away from the camera,
+where large changes in heading are necessary. Due to this phenomenon, the
+controller often over- or under-compensates in certain regions, depending upon
+the PID tuning and the distance from the camera. Tracking noise can also cause
+passage negotiation to fail. Variations in the H$^2$Bird hand launch can cause
+the system to respond differently for similar initial conditions. In addition,
+the H$^2$Bird battery was changed approximately every 5 tests, adding more
+variation to the H$^2$Bird performance by affecting the motor power output.
+
+%----------------------------------------------------------------------------%
+\subsection{Model Verification}
 
@@
 
-\begin{figure*}[tb]
-\begin{minipage}[b]{0.45\linewidth}
-\centering
-\includegraphics[width=\linewidth]{figures/flight_paths_feasible.pdf}
-\caption{Plot of camera field of view and experimental trials within the feasible region.}
-\label{fig:flight_paths_feasible}
-\end{minipage}
-\hfill
-\begin{minipage}[b]{0.45\linewidth}
-\centering
-\includegraphics[width=\textwidth]{figures/flight_paths.pdf}
-\caption{Plot of camera field-of-view and perimeter experimental trials overlayed.}
-\label{fig:flight_paths}
-\end{minipage}
-\end{figure*}
-
-The geometrically infeasible region is computed using the minimum
-turning radius and the geometry of the camera's field-of-view. Both the 
-regions in red and blue indicate the infeasible region determined in simulation. 
-We ran a simulation on a motion model similar to Dubin's car in ~\cite{lavalle:planning}, with the
-angular heading as the input rather than the angular velocity. We used the
-turning model, described in Section~\ref{sec:flight_control}, as the input to
-the motion model, with a PID controller steering the entire system to the
-center of the passage as the input to the tuning model.
+We simulated the motion of the ornithopter beginning from certain positional
+and angular initial conditions using several components representing the
+experimental layout of our system described in Figure~\ref{fig:block_diagram}.
+For the "ornithopter dynamics" block, we used the translational model
+described by~\ref{eq:total_model_plain} with the angular position model
+described by~\ref{eq:transfer_func}. We tuned the PID controller for the
+simulation so that paths from known positions looked similar to the behavior
+we observed in the experimental data. We also inlcuded the camera capture and
+pose estimation delays, and the PID controller sampling rate in the
+simulation.
 
@@
 (Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform
-grid of 0.05 meter increments within the camera viewing region, and an 
-initial angular position perpendicular to the narrow passage plane.
+grid of 0.05 meter increments within the camera viewing region, and an initial
+angular position perpendicular to the narrow passage plane. The geometrically
+infeasible region is computed using the minimum turning radius and the
+geometry of the camera's field-of-view and is indicated in blue. Both the
+regions in red and blue indicate the infeasible region determined in
+simulation. The feasible region is highlighted in green. A success consisted
+of a path that intersects the window at some point and does not leave the
+camera viewing triangle. Figure~\ref{fig:feasible_set} is validated by our
+experimenal data, as there are no successes within the geometrically
+infeasible region. There are a few successes near the boundary of the
+infeasible region determined in simulation, but this phenomenon canbe
+explained by the results in Figure~\ref{fig:heat_map}.
 
-The infeasible region
-calculated in simulation is a superset of the geometrically infeasible region.
-The simulation is more realistic as the rotational damping during the turn and
-the on-board PID controller cause the real system to react to inputs at a
-non-constant angular velocity.
+Figure~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
+simulating possible paths of the H$^2$Bird during our experiments. For each
+point within a 0.2 meter grid within the camera viewing triangle, we simulated
+10 trials using an initial heading randomly sampled from a uniform
+distribution between -90$^{\circ}$ and 90$^{\circ}$. As shown in the figure,
+there is a region in the middle with a 100\% probability of success, which is
+a subset of the feasible region presented in Figure~\ref{fig:feasible_set}.
+Along the edges of the camera viewing triangle, however, the probability of
+success decreases.
 
-The results of the experiments are shown in
-Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
-
-%----------------------------------------------------------------------------%
-\subsection{Model Verification}
-
-For initial conditions in the middle of the feasible region and in front of
-the camera, we achieved a success rate of 80\%. This case is depicted in
-Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
-because of several factors. On the edges of the camera frame, the yaw control
-inputs are large in magnitude, although this is not necessary in all cases.
-Since the control algorithm we are using has no notion of depth, it applies
-the same control input whether the ornithopter is close to the camera, where
-small changes in heading are adequate, or far away from the camera, where
-large changes in heading are necessary. Due to this phenomenon, the controller
-often over- or under-compensates in certain regions, depending upon the PID
-tuning and the distance from the camera. Tracking noise can also cause passage
-negotiation to fail.
+The results of our experimental trials verify the results of the Monte Carlo
+simulation of the model that we used. The trials that we conducted in
+Figure~\ref{fig:flight_paths_feasible} correspond to the circled region in
+Figure~\ref{heat_map}. The 80\% success rate that we experienced in the
+experiments corroborates our calculated probability of success of
+approximately 80\%.
+Additionally the diminished success rate of 50\% near the edges of the viewing triangle that we determined through experimentation supports the reduced probability of success calculated by the simulation.
 

### 2012-10-12 manuscript.tex
Update PF screenshots

@@
 are shown in \textcolor{red}{red}, the current state estimate as a 
-\textcolor{blue}{blue} circle, the window bounding box in
-\textcolor{red}{red}, and the target location as a \textcolor{green}{green} 
+\textcolor{yellow}{yellow} circle, the window bounding box in
+\textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
 circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 

### 2012-10-13 manuscript.tex
Clean up bibtex

@@
 gyroscope-based control with a ground station, which has power and weight 
-requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
+requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
 station provides heading estimates to the ornithoper by running a 
@@
 In order to conduct these experiments on cooperation, we designed a new
-flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
+flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit i-Bird 
 RC flyer power train and clap-fling 
-wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
+wings\footnote{Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
+\href{http://www.silverlit-flyingclub.com/wingsmaster/}
+     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
 the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A 
 tail propeller and servo-controlled tail elevator provide control in the yaw 
-and pitch axes. The on-board ImageProc 2.4\footnote{Expand! Accredit it to Stan Baek!} 
+and pitch axes. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
+\href{https://github.com/biomimetics/imageproc\_pcb}
+     {https://github.com/biomimetics/imageproc\_pcb}} 
 electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
@@
 constraints of highly mobile millirobotic ground vehicles, such as 
-OctoRoACH, which can traverse rough terrains with speed and agility [CITATION]. 
-%TODO OctoRoACH citation
+OctoRoACH, which can traverse rough terrains with speed and agility~\cite{Pullin2012Dynamic}.
 
@@
 \subsection{Ground Station Pose Estimation}
-%TODO ryan
 The ground station performs simple two-dimensional pose estimation on 
@@
 Our pose estimation algorithm tracks H$^2$Bird motion in the frame using a 
-particle filter~\cite[96-114]{thrun2005probabilistic}. We initialze particles 
+particle filter~\cite{thrun2005probabilistic}. We initialze particles 
 uniformly across the frame. To improve numerical stability, we  normalize the 
@@
 gyroscope data. We then used
-MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
-simple low-order model to the step response as discussed in Section~\ref{sec:model_attitude}.
+MATLAB\footnote{The MathWorks, Inc. Matlab:
+\href{http://www.mathworks.com/products/matlab/}
+     {http://www.mathworks.com/products/matlab/}} to fit a
+simple low-order model to the step response as discussed in Sec.~\ref{sec:model_attitude}.
 
@@
 The results of the step response experiment are described in
-Figure~\ref{fig:step_response}. The first-order model that we fit to the
+Fig.~\ref{fig:step_response}. The first-order model that we fit to the
 response is in Equation~\ref{eq:transfer_func}.
@@
 \end{equation}
-%TODO H(s) in a labeled equation
-%TODO footnote for MATLAB
 
@@
 
-%TODO equation(s) for how you arrive at the minimum turning radius
-%TODO footnote for MATLAB
-
 %............................................................................%
@@
 \label{sec:visual_servoing}
-%TODO cameron
-
-%TODO make these span the page, with identical scales/style
 
@@
 The results of our experimental trials are shown in
-Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
+Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
 initial conditions in the middle of the testing space and in front of the
 camera, we achieved a success rate of 80\%. This case is depicted in
-Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
+Fig.~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
+feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminishes
 to 50\% because of several factors. On the edges of the camera frame, the yaw
@@
 and angular initial conditions using several components representing the
-experimental layout of our system described in Figure~\ref{fig:block_diagram}.
+experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
 For the "ornithopter dynamics" block, we used the translational model
@@
 of a path that intersects the window at some point and does not leave the
-camera viewing triangle. Figure~\ref{fig:feasible_set} is validated by our
-experimenal data, as there are no successes within the geometrically
+camera viewing triangle. Fig.~\ref{fig:feasible_set} is validated by our
+experimental data, as there are no successes within the geometrically
 infeasible region. There are a few successes near the boundary of the
-infeasible region determined in simulation, but this phenomenon canbe
-explained by the results in Figure~\ref{fig:heat_map}.
+infeasible region determined in simulation, but this phenomenon can be
+explained by the results in Fig.~\ref{fig:heat_map}.
 
-Figure~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
+Fig.~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
 simulating possible paths of the H$^2$Bird during our experiments. For each
@@
 there is a region in the middle with a 100\% probability of success, which is
-a subset of the feasible region presented in Figure~\ref{fig:feasible_set}.
+a subset of the feasible region presented in Fig.~\ref{fig:feasible_set}.
 Along the edges of the camera viewing triangle, however, the probability of
@@
 simulation of the model that we used. The trials that we conducted in
-Figure~\ref{fig:flight_paths_feasible} correspond to the circled region in
-Figure~\ref{heat_map}. The 80\% success rate that we experienced in the
+Fig.~\ref{fig:flight_paths_feasible} correspond to the circled region in
+Fig.~\ref{fig:heat_map}. The 80\% success rate that we experienced in the
 experiments corroborates our calculated probability of success of
@@
 \section{Conclusions and Future Work}
-%TODO ryan
+%TODO 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-10-13 manuscript.tex
Various fixes based on Ron's feedback.

@@
 navigating through narrow openings. Finally, we verify this model through 
-with collected in experiments with narrow passage traversal.
+with experiments with narrow passage traversal.
 \end{abstract}
@@
 particularly severe for flapping wing MAVs, a class of MAVs that have inherently superior 
-safety, efficiency, and noise profile to rotor craft MAVs, but at the 
+safety, and noise profile to rotor craft MAVs, but at the 
 expense of payload weight.
@@
 %modeling ( won't necessarily stay in this spot)
-Modeling and control of flapping wing MAVs is a well studied subject. 
+Modeling and control of flapping wing MAVs is becominga well studied subject. 
 Shigeoka discusses velocity and altitude control in
-~\cite{Shigeoka:ornithopter}. In~\cite{delfly:design} deCroon et al. %TODO sentence fragment
+~\cite{Shigeoka:ornithopter}. Design and control of the sub 15 gram DelFly
+ornithopter is described by deCroon et al.~\cite{delfly:design}.
 In contrast, there is little previous work on guidance of flapping wing 
@@
 camera~\cite{baek:altitude}. de Croon et al. demonstrated obstacle avoidance 
-with an onboard camera and offboard processing~\cite{delfly:avoid}. 
+with an onboard camera and offboard processing~\cite{delfly:avoid}. For sub 20 gram
+ornithopters, Baek et al. \cite{baek:tracking} is one of the few examples of
+completely autonomous target seeking.
 %[However, these papers do not study the system blah blah blah.]
@@
 Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
-maintain high-rate attitude control. We estimate the vehicle pose by naively
-integrating the IMU, and use PID for controlling the vehicle.
+maintain high-sample-rate attitude control. We estimate the vehicle pose by naively
+integrating the IMU, and use a proportional plus integral plus derivative (PID) 
+controller for controlling the vehicle.
 
-The high pitch angles H$^2$Bird achieves in routine flight distinguishes it 
-from other MAVs. Baek 
-demonstrated control on a similar vehicle using PID and an Euler angle
+The high pitch angles flapping wing flyers achieve in routine flight distinguish them 
+from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
 parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of the vehicle orientation 
@@
 represent the vehicle orientation with quaternions. An additional benefit of 
-the quaternion representation its ability to represent relative body and 
+the quaternion representation is its ability to represent relative body and 
 world coordinate angular displacements compactly with quaternion 
@@
 a wooden frame meant to simulate a window. As the H$^2$Bird cannot detect the
-target on its own,we used the ground station to provide remote headings for
+target on its own, we used the ground station to provide remote headings for
 navigation to the window. Ground truth data was collected via a motion capture
@@
 \includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
-\caption{Conceptual sketch of experimental environment.}
+\caption{Conceptual sketch of experimental setup.}
 \label{fig:experiment_cartoon}

### 2012-10-13 manuscript.tex
Added Floreano textbook reference.
Minor additions

--- 2012-10-13 manuscript.tex
@@
 \subsubsection{Emission Model}
+Our emission model (Eq.~\ref{eq:emission_model}) effectively performs \naive
 
@@
 This model accounts for both the yaw dynamics of the ornithopter and the internal PID controller on the actuator inputs for the tail rotor.
-We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experiementation.
+We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experimentation.
 The format of the model we decided up on is shown in Equation~\ref{eq:plain_model}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgements}
+\section{Acknowledgments}
 The authors would like to thank Fernando Garcia Bermudez for his 

### 2012-10-13 manuscript.tex

--- 2012-10-13 manuscript.tex
@@
 %modeling ( won't necessarily stay in this spot)
-Modeling and control of flapping wing MAVs is becominga well studied subject. 
+Modeling and control of flapping wing MAVs is becominga well studied subject\cite{floreano:flying}. 
 Shigeoka discusses velocity and altitude control in

### 2012-10-13 manuscript.tex
Fix oops
adits

--- 2012-10-13 manuscript.tex
@@
 \usepackage[pdftex]{hyperref}
-\usepackage{amsmath}
-
-\newlength\eqnindent
-\newlength\eqncolwid
-\newcommand\eqntext[2][0pt]{%
-  \setlength\parindent{0pt}%
-      \setlength\eqnindent{#1}\setlength\eqncolwid{\linewidth}%
-  \setlength\eqncolwid{\dimexpr\linewidth-\eqnindent}%
-  \hspace*{\eqnindent}\parbox{\eqncolwid}{#2}%
-}
 
@@
 gyroscope-based control with a ground station, which has power and weight 
-requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
+requirements appropriate for deployment on 30 gram a ground vehicle. The ground 
 station provides heading estimates to the ornithoper by running a 
@@
 \end{figure}
-The power, size, and weight constraints of micro air vehicle (MAV) 
-platforms significantly limit on-board sensing and computational 
-resources, and thus the complexity of the control and state estimation 
-algorithms which can be executed on-board. These constraints are 
-particularly severe for flapping wing MAVs, a class of MAVs that have inherently superior 
-safety, efficiency, and noise profile to rotor craft MAVs, but at the 
-expense of payload weight.
+The power, size, and weight constraints of micro air vehicle (MAV) platforms
+significantly limit on-board sensing and computational resources, and thus the
+complexity of the control and state estimation algorithms which can be
+executed on-board. These constraints are particularly severe for flapping wing
+MAVs, a class of MAVs that have inherently superior safety, efficiency, and
+noise profile to rotor craft MAVs, but at the expense of payload weight.
 %Intro portion about previous works in ornithopter coorperative control/
-%modeling ( won't necessarily stay in this spot)
-Modeling and control of flapping wing MAVs is a well studied subject. 
-Shigeoka discusses velocity and altitude control in
-~\cite{Shigeoka:ornithopter}. In~\cite{delfly:design} deCroon et al. %TODO sentence fragment
-In contrast, there is little previous work on guidance of flapping wing 
-MAVs. Baek demonstrated altitude control with an external 
-camera~\cite{baek:altitude}. de Croon et al. demonstrated obstacle avoidance 
-with an onboard camera and offboard processing~\cite{delfly:avoid}. 
-%[However, these papers do not study the system blah blah blah.]
+%modeling ( won't necessarily stay in this spot) Modeling and control of
+flapping wing MAVs is a well studied subject. Shigeoka discusses velocity and
+altitude control in ~\cite{Shigeoka:ornithopter}. In~\cite{delfly:design}
+deCroon et al. %TODO sentence fragment In contrast, there is little previous
+work on guidance of flapping wing MAVs. Baek demonstrated altitude control
+with an external camera~\cite{baek:altitude}. de Croon et al. demonstrated
+obstacle avoidance with an onboard camera and offboard
+processing~\cite{delfly:avoid}. %[However, these papers do not study the system blah blah blah.]
 
 Autonomous control of flapping winged MAVs is a complex undertaking because of
-the non-linear flight dynamics as presented
-in~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Although the dynamics of
-flapping-winged flight can be averaged with respect to the wingbeat period as
-presented in~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, we develop a
-linear, one-dimensional model to predict the behavior of the system, similar
-to~\cite{Shigeoka:ornithopter}. In addition to the nonlinearities, the system
-is underactuated due to size constraints. Many of the methods for controlling
-these platforms make extensive use of GPS in outdoor
-environments~\cite{kingston:timeattitude, kanade:3dvision} and expansive
-motion capture systems in indoor environments. While operating indoors, using
+the non-linear flight
+dynamics~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. One approach to
+simplify these dynamic models is to use averaging over the wingbeat
+period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}. We desire a
+simpler approach, so we develop a linear, one-dimensional model to predict the
+yaw behavior of the system, similar to the model that Shigeoka used
+in~\cite{Shigeoka:ornithopter} for simulating the velocity of a MAV in
+one-dimensionally. In addition to the nonlinearities, the system is
+underactuated (there is no roll control) due to size constraints.
+
+Many of the methods for
+controlling these platforms make extensive use of GPS in outdoor
+environments and expansive
+motion capture systems in indoor environments~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
 GPS to estimate the pose of the MAV is not possible, and motion capture
@@
 In order to conduct these experiments on cooperation, we designed a new
-flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit i-Bird 
+flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit I-Bird 
 RC flyer power train and clap-fling 
-wings\footnote{Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
-\href{http://www.silverlit-flyingclub.com/wingsmaster/}
-     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
+wings\footnote{\url{http://www.silverlit-flyingclub.com/wingsmaster/}}, 
 the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A 
 tail propeller and servo-controlled tail elevator provide control in the yaw 
-and pitch axes. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
-\href{https://github.com/biomimetics/imageproc\_pcb}
-     {https://github.com/biomimetics/imageproc\_pcb}} 
+and pitch axes. The on-board ImageProc 2.4\footnote{Expand! Accredit it to Stan Baek!} 
 electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
@@
 constraints of highly mobile millirobotic ground vehicles, such as 
-OctoRoACH, which can traverse rough terrains with speed and agility~\cite{Pullin2012Dynamic}.
+OctoRoACH, which can traverse rough terrains with speed and agility [CITATION]. 
+%TODO OctoRoACH citation
 
@@
 \subsection{Ground Station Pose Estimation}
-The ground station performs simple two-dimensional pose estimation on 
-H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
-estimate is the feedback input to the cooperative visual servoing feedback 
-loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
-approach to monocular two-dimensional object tracking in pixel space. This 
-allows us to perform real-time video tracking using the modest computational 
-resources available on the ground station, and similarly millirobotic ground 
-vehicles.
+%TODO ryan
+The ground station performs simple two-dimensional pose estimation on the H$^2$Bird using 
+a particle filter-based motion tracking algorithm. This pose estimate is 
+the feedback input to the cooperative visual servoing feedback loop (Sec. \ref{sec:visual_servoing_concept}).
 
-Our pose estimation algorithm tracks H$^2$Bird motion in the frame using a 
-particle filter~\cite{thrun2005probabilistic}. We initialze particles 
-uniformly across the frame. To improve numerical stability, we  normalize the 
-weights of all particles on each iteration, and only resample the particle 
-population when the mean particle weight falls below a pre-determined 
-threshold.
+We limit our pose estimation approach to monocular two-dimensional object 
+tracking in pixel-space. This allows us to perform high-rate video tracking 
+using the modest computational resources available on the ground station, 
+and similarly millirobotic ground vehicles.
 
-%............................................................................%
-\subsubsection{Motion Model}
-Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
-centered around each particle's current position, augmented with an 
-$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
-$x^{[i]}_t$ is the state of the $i^\text{th}$ particle at time $t$, $\Sigma$ 
-is a chosen covariance matrix, and $b$ is vector of the height and width of 
-the camera frame.
-\begin{equation}
-\label{eq:motion_model}
-x^{[i]}_t \sim \begin{cases}
-\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
-\mathcal{U}(0,b),& \text{with probability } \epsilon \\
-\end{cases}\\
-\end{equation}
-The experiments presented in this paper used a diagonal covariance matrix 
-with variance 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
-transition model is sufficient for tracking motion in which frame-to-frame 
-position changes are relatively small. The $\epsilon$-random uniform 
-transitions increase tracking robustness, by forcing the filter to always 
-sample all of the pixel space. This prevents overconfidence and collapse, 
-and decreases reacquisition delay.
-
-%............................................................................%
-\subsubsection{Emission Model}
-Our emission model (Eq.~\ref{eq:emission_model}) effectively performs \naive
+~\cite{thrun2005probabilistic}
 
@@
 \label{sec:visual_servoing_concept}
+%TODO ryan
 
@@
 are shown in \textcolor{red}{red}, the current state estimate as a 
-\textcolor{yellow}{yellow} circle, the window bounding box in
-\textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
+\textcolor{blue}{blue} circle, the window bounding box in
+\textcolor{red}{red}, and the target location as a \textcolor{green}{green} 
 circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
@@
 This model accounts for both the yaw dynamics of the ornithopter and the internal PID controller on the actuator inputs for the tail rotor.
-We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experimentation.
+We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experiementation.
 The format of the model we decided up on is shown in Equation~\ref{eq:plain_model}.
@@
 gyroscope data. We then used
-MATLAB\footnote{The MathWorks, Inc. Matlab:
-\href{http://www.mathworks.com/products/matlab/}
-     {http://www.mathworks.com/products/matlab/}} to fit a
-simple low-order model to the step response as discussed in Sec.~\ref{sec:model_attitude}.
+MATLAB\footnote{\url{http://www.mathworks.com/products/matlab/}} to fit a
+simple low-order model to the step response as discussed in Section~\ref{sec:model_attitude}.
 
@@
 The results of the step response experiment are described in
-Fig.~\ref{fig:step_response}. The first-order model that we fit to the
+Figure~\ref{fig:step_response}. The first-order model that we fit to the
 response is in Equation~\ref{eq:transfer_func}.
@@
 \end{equation}
+%TODO H(s) in a labeled equation
+%TODO footnote for MATLAB
 
@@
 
+%TODO equation(s) for how you arrive at the minimum turning radius
+%TODO footnote for MATLAB
+
 %............................................................................%
@@
 \label{sec:visual_servoing}
+%TODO cameron
+
+%TODO make these span the page, with identical scales/style
 
@@
 The results of our experimental trials are shown in
-Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
+Figures~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
 initial conditions in the middle of the testing space and in front of the
 camera, we achieved a success rate of 80\%. This case is depicted in
-Fig.~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminishes
+Figure~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
+feasible region in Figure~\ref{fig:flight_paths}, the success rate diminishes
 to 50\% because of several factors. On the edges of the camera frame, the yaw
@@
 and angular initial conditions using several components representing the
-experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
+experimental layout of our system described in Figure~\ref{fig:block_diagram}.
 For the "ornithopter dynamics" block, we used the translational model
@@
 of a path that intersects the window at some point and does not leave the
-camera viewing triangle. Fig.~\ref{fig:feasible_set} is validated by our
-experimental data, as there are no successes within the geometrically
+camera viewing triangle. Figure~\ref{fig:feasible_set} is validated by our
+experimenal data, as there are no successes within the geometrically
 infeasible region. There are a few successes near the boundary of the
-infeasible region determined in simulation, but this phenomenon can be
-explained by the results in Fig.~\ref{fig:heat_map}.
+infeasible region determined in simulation, but this phenomenon canbe
+explained by the results in Figure~\ref{fig:heat_map}.
 
-Fig.~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
+Figure~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
 simulating possible paths of the H$^2$Bird during our experiments. For each
@@
 there is a region in the middle with a 100\% probability of success, which is
-a subset of the feasible region presented in Fig.~\ref{fig:feasible_set}.
+a subset of the feasible region presented in Figure~\ref{fig:feasible_set}.
 Along the edges of the camera viewing triangle, however, the probability of
@@
 simulation of the model that we used. The trials that we conducted in
-Fig.~\ref{fig:flight_paths_feasible} correspond to the circled region in
-Fig.~\ref{fig:heat_map}. The 80\% success rate that we experienced in the
+Figure~\ref{fig:flight_paths_feasible} correspond to the circled region in
+Figure~\ref{heat_map}. The 80\% success rate that we experienced in the
 experiments corroborates our calculated probability of success of
-approximately 80\%.
-Additionally the diminished success rate of 50\% near the edges of the viewing triangle that we determined through experimentation supports the reduced probability of success calculated by the simulation.
+approximately 80\%. Additionally, the diminished success rate of 50\% near the
+edges of the viewing triangle that we determined through experimentation
+supports the reduced probability of success calculated by the simulation.
 
@@
 \section{Conclusions and Future Work}
-%TODO 
+%TODO ryan
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
+\section{Acknowledgements}
 The authors would like to thank Fernando Garcia Bermudez for his 

### 2012-10-13 manuscript.tex

--- 2012-10-13 manuscript.tex
@@
 Autonomous control of flapping winged MAVs is a complex undertaking because of
-the non-linear flight dynamics as presented
-in~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Although the dynamics of
-flapping-winged flight can be averaged with respect to the wingbeat period as
-presented in~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, we develop a
-linear, one-dimensional model to predict the behavior of the system, similar
-to~\cite{Shigeoka:ornithopter}. In addition to the nonlinearities, the system
-is underactuated due to size constraints. Many of the methods for controlling
-these platforms make extensive use of GPS in outdoor
-environments~\cite{kingston:timeattitude, kanade:3dvision} and expansive
-motion capture systems in indoor environments. While operating indoors, using
+the non-linear flight
+dynamics~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. One approach to
+simplify these dynamic models is to use averaging over the wingbeat
+period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}. We desire a
+simpler approach, so we develop a linear, one-dimensional model to predict the
+yaw behavior of the system, similar to the model that Shigeoka used
+in~\cite{Shigeoka:ornithopter} for simulating the velocity of a MAV in
+one-dimensionally. In addition to the nonlinearities, the system is
+underactuated (there is no roll control) due to size constraints.
+
+Many of the methods for
+controlling these platforms make extensive use of GPS in outdoor
+environments and expansive
+motion capture systems in indoor environments~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
 GPS to estimate the pose of the MAV is not possible, and motion capture
@@
 \subsubsection{Emission Model}
-Our emission model (Eq.~\ref{eq:emission_model}) effectively performs \naive
+Our emission model (Eq.~\ref{eq:emission_model}) effectively performs.
 
@@
 experiments corroborates our calculated probability of success of
-approximately 80\%.
-Additionally the diminished success rate of 50\% near the edges of the viewing triangle that we determined through experimentation supports the reduced probability of success calculated by the simulation.
+approximately 80\%. Additionally, the diminished success rate of 50\% near the
+edges of the viewing triangle that we determined through experimentation
+supports the reduced probability of success calculated by the simulation.
 

### 2012-10-13 manuscript.tex
Drafted cooperative visual servoing description section.

--- 2012-10-13 manuscript.tex
@@
 \label{sec:visual_servoing_concept}
+Remote guidance of the H$^2$Bird is performed by commanding relative changes in the
+reference orientation. With estimates of the vehicle pose, we calculate an altitude
+error and horizontal error. Incremental changes to the H$^2$Bird pitch and yaw are
+then commanded according to a PID control law, with positive altitude error
+corresponding to negative incremental pitches and positive horizontal error
+corresponding to negative incremental yaw. The controller integrator term is windowed
+to prevent windup when the bird is being launched, and the derivative term was tuned
+by hand to dampen oscillations associated with the system latency [?].
 

### 2012-10-13 manuscript.tex
Updated quaternion section for clarity.

@@
 Our implementation of quaternion-based control is similar to Knoebe's~\cite{knoebe:quatcontrol}:
-Given a reference quaternion $q$, we define the error $q_e$ as the rotation 
-required to reach the reference pose from the measured pose (Eq.~\ref{eq:quat_error}). 
-\begin{equation}
+Given a reference pose represented in quaternion form $q_r$, we define the error as the rotation $q_e$ 
+required to reach the reference pose from the estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
+\begin{align}
 \label{eq:quat_error}
-q_e = q'\otimes q_r
-\end{equation}
-We then convert this error quaternion to an angle axis, and use this 
-representation as input to the PID controller 
-(Eqs.~\ref{eq:quat_angle} and~\ref{eq:quat_linearize}). 
+q_e &= q^{\prime}\cdot q_r \\
+		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
+\end{align}
+We then convert this error quaternion to an angle axis representation and use this 
+as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
+These calculations are performed onboard with trigonometric lookup tables for speed.
+\begin{align}
+\label{eq:quat_linearize}
+Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
+		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
+\end{align}
 \begin{equation}
@@
 \alpha = 2\arccos(q_{e,w})
-\end{equation}
-\begin{equation}
-\label{eq:quat_linearize}
-Q_e = \alpha*q_e/\sin(\alpha /2)
 \end{equation}

### 2012-10-13 manuscript.tex
Added footnotes.

@@
 RC flyer power train and clap-fling 
-wings\footnote{Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
+wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
 \href{http://www.silverlit-flyingclub.com/wingsmaster/}
@@
 
-Our ground station software is based on the Ubuntu Linux distribution, 
+Our ground station software is based on the Ubuntu Linux\footnote{\raggedright Ubuntu: 
+Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
 modified with a custom version of the Linux kernel. We use the Python
-programming language for control, robot communication, and telemetry, and 
-the OpenCV computer vision libraries for image capture and processing.
+\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
+{http://www.python.org/}} programming language for control, robot communication, 
+and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
+\href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
+computer vision libraries for image capture and processing.
 %TODO footnotes: Ubuntu, Linux, Python, OpenCV

### 2012-10-13 manuscript.tex
Finish system concept/pose estimation

--- 2012-10-13 manuscript.tex
@@
 \usepackage{amsmath}
-
-\newlength\eqnindent
-\newlength\eqncolwid
-\newcommand\eqntext[2][0pt]{%
-  \setlength\parindent{0pt}%
-      \setlength\eqnindent{#1}\setlength\eqncolwid{\linewidth}%
-  \setlength\eqncolwid{\dimexpr\linewidth-\eqnindent}%
-  \hspace*{\eqnindent}\parbox{\eqncolwid}{#2}%
-}
+\usepackage{amssymb}
+\providecommand{\norm}[1]{\lVert#1\rVert}
 
@@
 \end{equation}
-The experiments presented in this paper used a diagonal covariance matrix 
+The experiments presented in this work use a diagonal covariance matrix 
 with variance 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
@@
 \subsubsection{Emission Model}
-Our emission model (Eq.~\ref{eq:emission_model}) effectively performs.
+Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
+approach to motion tracking via na\"{i}ve background subtraction[CITATION].
+\begin{equation}
+\label{eq:emission_model}
+w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
+\end{equation} 
+The system provides the emission model with $f_0$, the frame representing 
+the background and $f_t$, the most recent frame captured. The model assigns 
+a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
+distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
+$x^{[i]}_t$. Hence, particles are more likely to be resampled in portions of 
+the image which appear different from the background.
+
+For the experiments presented in this work, we assumed that neither the bird 
+nor any other moving object were visible in the first frame captured on system 
+start. We saved the first frame captured by the system, and provided it as 
+the background frame $f_0$ for every iteration of the particle filter. This 
+resulted in a sampling approach to na\"{i}ve background subtraction.
+
+Na\"{i}ve background subtraction is brittle to changes in the scene over 
+time[CITATION]. However, we found it very efficient, and reliable enough for the 
+short time spans necessary for H$^2$Bird to successfully negotiate narrow 
+openings, which are typically less than five seconds. Substituting $f_0$ 
+with a more sophisticated extraction of the background would implement a 
+more sophisticated approach to background subtraction, with no changes to 
+the underlying motion tracking filter.
 
@@
 \label{sec:visual_servoing_concept}
-Remote guidance of the H$^2$Bird is performed by commanding relative changes in the
-reference orientation. With estimates of the vehicle pose, we calculate an altitude
-error and horizontal error. Incremental changes to the H$^2$Bird pitch and yaw are
-then commanded according to a PID control law, with positive altitude error
-corresponding to negative incremental pitches and positive horizontal error
-corresponding to negative incremental yaw. The controller integrator term is windowed
-to prevent windup when the bird is being launched, and the derivative term was tuned
-by hand to dampen oscillations associated with the system latency [?].
 

### 2012-10-13 manuscript.tex
new heat map and text

--- 2012-10-13 manuscript.tex
@@
 \usepackage{amsmath}
-\usepackage{amssymb}
-\providecommand{\norm}[1]{\lVert#1\rVert}
+
+\newlength\eqnindent
+\newlength\eqncolwid
+\newcommand\eqntext[2][0pt]{%
+  \setlength\parindent{0pt}%
+      \setlength\eqnindent{#1}\setlength\eqncolwid{\linewidth}%
+  \setlength\eqncolwid{\dimexpr\linewidth-\eqnindent}%
+  \hspace*{\eqnindent}\parbox{\eqncolwid}{#2}%
+}
 
@@
 \end{equation}
-The experiments presented in this work use a diagonal covariance matrix 
+The experiments presented in this paper used a diagonal covariance matrix 
 with variance 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
@@
 \subsubsection{Emission Model}
-Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
-approach to motion tracking via na\"{i}ve background subtraction[CITATION].
-\begin{equation}
-\label{eq:emission_model}
-w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
-\end{equation} 
-The system provides the emission model with $f_0$, the frame representing 
-the background and $f_t$, the most recent frame captured. The model assigns 
-a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
-distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
-$x^{[i]}_t$. Hence, particles are more likely to be resampled in portions of 
-the image which appear different from the background.
-
-For the experiments presented in this work, we assumed that neither the bird 
-nor any other moving object were visible in the first frame captured on system 
-start. We saved the first frame captured by the system, and provided it as 
-the background frame $f_0$ for every iteration of the particle filter. This 
-resulted in a sampling approach to na\"{i}ve background subtraction.
-
-Na\"{i}ve background subtraction is brittle to changes in the scene over 
-time[CITATION]. However, we found it very efficient, and reliable enough for the 
-short time spans necessary for H$^2$Bird to successfully negotiate narrow 
-openings, which are typically less than five seconds. Substituting $f_0$ 
-with a more sophisticated extraction of the background would implement a 
-more sophisticated approach to background subtraction, with no changes to 
-the underlying motion tracking filter.
+Our emission model (Eq.~\ref{eq:emission_model}) effectively performs.
 
@@
 
+The model of the ground station combines both the camera model and the
+computational model for pose estimation. The camera model contains both a
+latency, which would occur in between the "camera" and "ornithopter dynamics"
+in Fig.~\ref{fig:block_diagram}, and a model for the viewing angle. The
+viewing angle for the camera is modeled as a pixel map from one edge of a
+63$^{\circ}$ viewing triangle to the other. The camera has an image width of
+640 pixels. The desired heading angle in degrees for a given distance from the
+camera is calculated according to Equation~\ref{eq:desired_heading}, where d
+is the distance of the robot in front of the camera and x is the horizontal
+position of the robot relative to the center of the window in meters. This
+formula models the fact that the camera cannot determine the depth of the
+robot.
+
+\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
+\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
+\end{equation}
+
+The computational model consists of the latency for the operations required to
+run the particle filter and calculate the input to send to the robot.
+
 %----------------------------------------------------------------------------%
@@
 \subsubsection{Ground Station Pose Estimation}
+
+To measure the capture latency of the camera, we alternatively flashed black
+and white color in front of the camera and measured the time lapse before it
+detected the color change. We averaged this result over 50 trials to determine
+the capture latency.
+
+To compute the particle filter latency, we used system timers to determine the
+average time from the start of computation to the end of computation.
 
@@
 \begin{aligned}
-\text{Minimum radius} & = \frac{360^{\circ}*\text{Rise time}*\text{Flight speed}}{\text{Final angle}*2*\pi}\\
+\text{Minimum radius} & = \frac{360^{\circ}*\text{Rise time}*\text{Flight
+speed}}{\text{Final angle}*2*\pi}\\
 & = \frac{360^{\circ}*1.4s*1.2\frac{m}{s}}{90^{\circ}*2*\pi} = 1.07 m
@@
 \subsubsection{Ground Station Pose Estimation}
+
+The latency that we measured for the camera was 25ms and the measured
+partiicle filter computational latency was 12.5ms.
 

### 2012-10-13 manuscript.tex
Added Ryan's citations

--- 2012-10-13 manuscript.tex
@@
 \section{Introduction}
-%TODO ryan
 \begin{figure}[!tb]
@@
 
+\cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
+\cite{Rudol2008Micro}\cite{Luo2011Air}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Robotic and Vision Platforms}
-%TODO ryan
-
 %----------------------------------------------------------------------------%

### 2012-10-13 manuscript.tex
Squashing passive voice, various grammar edits.

@@
 %modeling ( won't necessarily stay in this spot)
-Modeling and control of flapping wing MAVs is becominga well studied subject\cite{floreano:flying}. 
-Shigeoka discusses velocity and altitude control in
-~\cite{Shigeoka:ornithopter}. Design and control of the sub 15 gram DelFly
-ornithopter is described by deCroon et al.~\cite{delfly:design}.
-In contrast, there is little previous work on guidance of flapping wing 
-MAVs. Baek demonstrated altitude control with an external 
+Modeling and control of flapping wing MAVs is becoming a well studied subject\cite{floreano:flying}. 
+Shigeoka discusses velocity and altitude control in ~\cite{Shigeoka:ornithopter}. 
+Design and control of the sub--15 gram DelFly ornithopter is described by 
+deCroon et al.~\cite{delfly:design}. In contrast, there is little previous work 
+on guidance of flapping wing MAVs. Baek demonstrated altitude control with an external 
 camera~\cite{baek:altitude}. de Croon et al. demonstrated obstacle avoidance 
-with an onboard camera and offboard processing~\cite{delfly:avoid}. For sub 20 gram
+with an onboard camera and offboard processing~\cite{delfly:avoid}. For sub--20 gram
 ornithopters, Baek et al. \cite{baek:tracking} is one of the few examples of
@@
 Autonomous control of flapping winged MAVs is a complex undertaking because of
-the non-linear flight
-dynamics~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. One approach to
-simplify these dynamic models is to use averaging over the wingbeat
+the non-linear flight dynamics~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
+One approach to simplify these dynamic models is to use averaging over the wingbeat
 period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}. We desire a
@@
 
-Many of the methods for
-controlling these platforms make extensive use of GPS in outdoor
-environments and expansive
-motion capture systems in indoor environments~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
+Many of the methods for controlling these platforms make extensive use of GPS in outdoor
+environments and expansive motion capture systems in indoor environments
+~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
 GPS to estimate the pose of the MAV is not possible, and motion capture
@@
 BeagleBoard consumes approximately 1 watt while running our computer vision 
-and control algorithms. It uses the same processor as popular sub-10 gram 
+and control algorithms. It uses the same processor as popular sub--10 gram 
 processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
@@
 computer vision libraries for image capture and processing.
-%TODO footnotes: Ubuntu, Linux, Python, OpenCV
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 the quaternion representation is its ability to represent relative body and 
-world coordinate angular displacements compactly with quaternion 
-multiplication.
+world coordinate angular displacements compactly with quaternion multiplication.
 
@@
 The experiments presented in this work use a diagonal covariance matrix 
-with variance 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
+with a variance of 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
 transition model is sufficient for tracking motion in which frame-to-frame 
@@
 reference orientation. With estimates of the vehicle pose, we calculate an altitude
-error and horizontal error. Incremental changes to the H$^2$Bird pitch and yaw are
-then commanded according to a PID control law, with positive altitude error
-corresponding to negative incremental pitches and positive horizontal error
-corresponding to negative incremental yaw. The controller integrator term is windowed
-to prevent windup when the bird is being launched, and the derivative term was tuned
-by hand to dampen oscillations associated with the system latency [?].
+error and horizontal error. The ground station then commands incremental changes to 
+the H$^2$Bird pitch and yaw are according to a PID control law, with positive altitude error
+corresponding to negative pitch increments and positive horizontal error
+corresponding to negative yaw increments. We window the controller integrator term 
+to prevent windup when the bird is being launched, and tuned the derivative term
+to dampen oscillations associated with the system latency [?].
 
@@
 We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experimentation.
-The format of the model we decided up on is shown in Equation~\ref{eq:plain_model}.
+The model format is shown in Equation~\ref{eq:plain_model}.
 
@@
 To model the translational position of the ornithopter, we used a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state.
-So the overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the angular setpoint provided by the ground station as the input.
+The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the angular setpoint provided by the ground station as the input.
 
@@
 latency, which would occur in between the "camera" and "ornithopter dynamics"
-in Fig.~\ref{fig:block_diagram}, and a model for the viewing angle. The
-viewing angle for the camera is modeled as a pixel map from one edge of a
+in Fig.~\ref{fig:block_diagram}, and a model for the viewing angle. We model the 
+viewing angle for the camera as a pixel map from one edge of a
 63$^{\circ}$ viewing triangle to the other. The camera has an image width of
 640 pixels. The desired heading angle in degrees for a given distance from the
-camera is calculated according to Equation~\ref{eq:desired_heading}, where d
-is the distance of the robot in front of the camera and x is the horizontal
+camera is calculated according to Equation~\ref{eq:desired_heading}, where $d$
+is the distance of the robot in front of the camera and $x$ is the horizontal
 position of the robot relative to the center of the window in meters. This
-formula models the fact that the camera cannot determine the depth of the
+relation reflects the fact that the camera cannot determine the depth of the
 robot.
@@
 
-The computational model consists of the latency for the operations required to
+The computational model accounts for the latency added by the operations required to
 run the particle filter and calculate the input to send to the robot.
@@
 
-In total, 80 trials were run over a variety of starting positions. The goal of
+In total, we ran 80 trials over a variety of starting positions. The goal of
 our experiments was to verify the model described in
@@
 
-The latency that we measured for the camera was 25ms and the measured
+The latency that we measured for the camera was 25ms, and the measured
 partiicle filter computational latency was 12.5ms.
@@
 where large changes in heading are necessary. Due to this phenomenon, the
-controller often over- or under-compensates in certain regions, depending upon
+controller often over or under--compensates in certain regions, depending upon
 the PID tuning and the distance from the camera. Tracking noise can also cause

### 2012-10-13 manuscript.tex
Fixed dash mistakes.

@@
 Shigeoka discusses velocity and altitude control in ~\cite{Shigeoka:ornithopter}. 
-Design and control of the sub--15 gram DelFly ornithopter is described by 
+Design and control of the sub-15 gram DelFly ornithopter is described by 
 deCroon et al.~\cite{delfly:design}. In contrast, there is little previous work 
@@
 camera~\cite{baek:altitude}. de Croon et al. demonstrated obstacle avoidance 
-with an onboard camera and offboard processing~\cite{delfly:avoid}. For sub--20 gram
+with an onboard camera and offboard processing~\cite{delfly:avoid}. For sub-20 gram
 ornithopters, Baek et al. \cite{baek:tracking} is one of the few examples of
@@
 BeagleBoard consumes approximately 1 watt while running our computer vision 
-and control algorithms. It uses the same processor as popular sub--10 gram 
-processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
+and control algorithms. It uses the same processor as popular sub-10 gram 
+processor modules---such as the Gumstix Overo and LogicPD Torpedo---which are 
 light enough for deployment on millirobotic ground vehicles. We pair the 
@@
 Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
-maintain high-sample-rate attitude control. We estimate the vehicle pose by naively
+maintain high sample rate attitude control. We estimate the vehicle pose by naively
 integrating the IMU, and use a proportional plus integral plus derivative (PID) 
@@
 where large changes in heading are necessary. Due to this phenomenon, the
-controller often over or under--compensates in certain regions, depending upon
+controller often over or under-compensates in certain regions, depending upon
 the PID tuning and the distance from the camera. Tracking noise can also cause

### 2012-10-13 manuscript.tex
Recompress pf_screencap
Reworking introduction. Draft.

@@
 \end{figure}
-The power, size, and weight constraints of micro air vehicle (MAV) 
-platforms significantly limit on-board sensing and computational 
-resources, and thus the complexity of the control and state estimation 
-algorithms which can be executed on-board. These constraints are 
-particularly severe for flapping wing MAVs, a class of MAVs that have inherently superior 
-safety, and noise profile to rotor craft MAVs, but at the 
-expense of payload weight.
-%Intro portion about previous works in ornithopter coorperative control/
-%modeling ( won't necessarily stay in this spot)
-Modeling and control of flapping wing MAVs is becoming a well studied subject\cite{floreano:flying}. 
-Shigeoka discusses velocity and altitude control in ~\cite{Shigeoka:ornithopter}. 
-Design and control of the sub-15 gram DelFly ornithopter is described by 
-deCroon et al.~\cite{delfly:design}. In contrast, there is little previous work 
-on guidance of flapping wing MAVs. Baek demonstrated altitude control with an external 
-camera~\cite{baek:altitude}. de Croon et al. demonstrated obstacle avoidance 
-with an onboard camera and offboard processing~\cite{delfly:avoid}. For sub-20 gram
-ornithopters, Baek et al. \cite{baek:tracking} is one of the few examples of
-completely autonomous target seeking.
-%[However, these papers do not study the system blah blah blah.]
+Power, size, and weight constraints for micro air vehicle (MAV) 
+platforms are amongst the most restrictive of all robotic platforms. With
+significantly limited sensing and computing capabilities, only the most rudimentary
+control and state estimation algorithms can be executed on-board these platforms. 
+Even more restrictive are the constraints for flapping wing MAVs, 
+a class of MAVs that exhibit demonstrate safety and noise profiles superior to those
+of rotor craft MAVs, but at the expense of reduced payload weight.
 
-Autonomous control of flapping winged MAVs is a complex undertaking because of
-the non-linear flight dynamics~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
-One approach to simplify these dynamic models is to use averaging over the wingbeat
-period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}. We desire a
-simpler approach, so we develop a linear, one-dimensional model to predict the
-yaw behavior of the system, similar to the model that Shigeoka used
-in~\cite{Shigeoka:ornithopter} for simulating the velocity of a MAV in
-one-dimensionally. In addition to the nonlinearities, the system is
-underactuated (there is no roll control) due to size constraints.
-
-Many of the methods for controlling these platforms make extensive use of GPS in outdoor
-environments and expansive motion capture systems in indoor environments
-~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
-GPS to estimate the pose of the MAV is not possible, and motion capture
-requires expensive and stationary modification of the environment. The power
-and payload limits on MAVs make on-board sensing difficult, so multi-agent
-cooperative control schemes are apt solutions.
+Compounding these engineering restrictions are the many non-linear
+flight dynamics exhibited by flapping wing flight~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
+Much past research on flapping wing MAVs has been focused on overcoming the 
+difficulties associated with controlling and modeling in the face of these 
+non-linearities\cite{floreano:flying}. Common approaches to simplify these 
+dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
+\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict the 
+behavior in flight~\cite{Shigeoka:ornithopter}. Advances have lead to the
+development of some high-performance platforms, such as the sub-15 gram DelFly 
+as described by deCroon et al.~\cite{delfly:design}.
 
@@
 
+In contrast, there is considerably less previous work on guidance of flapping wing MAVs. 
+Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
+de Croon et al. demonstrated obstacle avoidance with an onboard camera and offboard 
+processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
+\cite{baek:tracking} is one of the few examples of scompletely autonomous target seeking.
+
+There also exists a body of work on guidance and control for fixed wing and rotary wing 
+MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
+~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
+GPS to estimate the pose of the MAV is not possible, and motion capture
+requires expensive and stationary modification of the environment. The power
+and payload limits on flapping wing MAVs make on-board sensing difficult, so multi-agent
+cooperative control schemes are apt solutions.
+
 \cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
 \cite{Rudol2008Micro}\cite{Luo2011Air}
+
+%TODO Add in final paragraph to lead-in to our work
 

### 2012-10-13 manuscript.tex
More voice-normalization commits.

@@
 % humphrey
-In order to conduct these experiments on cooperation, we designed a new
-flapping-wing MAV, known as the H$^2$Bird. Built around the Silverlit i-Bird 
+To fulfill the need for a lightweight, autonomous flapping wing MAV platform, 
+we designed the H$^2$Bird. Built around the Silverlit i-Bird 
 RC flyer power train and clap-fling 
@@
      {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
-the H$^2$Bird's wingspan is 26.5 cm and it has a flight weight of 13 grams. A 
+the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
 tail propeller and servo-controlled tail elevator provide control in the yaw 
@@
 Cooperative behavior is a promising strategy for autonomous navigation of 
-unexplored and unstructured environments. We intend that this system will be 
-fully mobile in these environments, not tied to a static ground station or 
+unexplored and unstructured environments. However, to operate in these environments,
+a cooperative system must be fully mobile, not tied to a static ground station or 
 motion capture laboratory. To demonstrate the feasibility of this approach, 
@@
 Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
-maintain high sample rate attitude control. We estimate the vehicle pose by naively
+high sample rate attitude control. We estimate the vehicle pose by
 integrating the IMU, and use a proportional plus integral plus derivative (PID) 
@@
 parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of the vehicle orientation 
-suffer from singularities around these high pitch angles, so we instead 
+suffer from singularities around high pitch angles, so we instead 
 represent the vehicle orientation with quaternions. An additional benefit of 
@@
 The ground station performs simple two-dimensional pose estimation on 
-H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
-estimate is the feedback input to the cooperative visual servoing feedback 
+H$^2$Bird using a particle filter-based motion tracking algorithm.
+The produced pose estimate is the feedback input to the cooperative visual servoing feedback 
 loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
@@
 allows us to perform real-time video tracking using the modest computational 
-resources available on the ground station, and similarly millirobotic ground 
+resources available on the ground station, and similarly, millirobotic ground 
 vehicles.
 
-Our pose estimation algorithm tracks H$^2$Bird motion in the frame using a 
-particle filter~\cite{thrun2005probabilistic}. We initialze particles 
-uniformly across the frame. To improve numerical stability, we normalize the 
-weights of all particles on each iteration, and only resample the particle 
-population when the mean particle weight falls below a pre-determined 
-threshold.
+Motion of the H$^2$Bird in the frame is tracked using a particle filter
+with an approach similar to that employed by Thrun~\cite{thrun2005probabilistic}. 
+Particles are distributed uniformly across the frame on initialization. 
+To improve numerical stability, we normalize the weights of all particles on 
+each iteration, and only resample the particle population when the mean particle 
+weight falls below a pre-determined threshold.
 
@@
 
-Na\"{i}ve background subtraction is brittle to changes in the scene over 
-time[CITATION]. However, we found it very efficient, and reliable enough for the 
+Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
+time[CITATION]. However, we found it very efficient and sufficiently reliable for the 
 short time spans necessary for H$^2$Bird to successfully negotiate narrow 
@@
 \label{sec:visual_servoing_concept}
-Remote guidance of the H$^2$Bird is performed by commanding relative changes in the
-reference orientation. With estimates of the vehicle pose, we calculate an altitude
+The H$^2$Bird is remotely guided by commanding relative changes in its
+reference orientation. Using relative changes removes the need to estimate the relative
+orientation between the ground station and vehicle frames of reference. 
+With estimates of the vehicle pose, we calculate an altitude
 error and horizontal error. The ground station then commands incremental changes to 

### 2012-10-13 manuscript.tex
Need more caffeine.

@@
 
-To model the yaw motion of the ornithopter, we developed a one-dimensional model using the physical heading angle as the output and the heading angle set by the ground station as the input.
-This model accounts for both the yaw dynamics of the ornithopter and the internal PID controller on the actuator inputs for the tail rotor.
-We decided upon a one-dimensional model to reduce the complexity of the overall model and enable simple angle inputs and outputs, which is similar to our methods during experimentation.
+Modeling of the vehicle pose is divided into two independent models for simplicity: one for
+yaw motion, and one for translational motion.
+
+We model the yaw motion of the ornithopter with a one-dimensional model to reduce the complexity of the overall model.
+Matching our physical implementation, we choose the reference heading angle as an input and model the dynamics from the vehicle and on-board PID controller to output the physical heading angle.
 The model format is shown in Equation~\ref{eq:plain_model}.
@@
 
-To model the translational position of the ornithopter, we used a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state.
-The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the angular setpoint provided by the ground station as the input.
+To model the translational position of the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state.
+The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.
 
@@
 We determined the turning speed and turning radius of the H$^2$Bird
-experimentally using the response of the system to the step input of a
+experimentally by measuring the response of the system to the step input of a
 clockwise 90 degree turn. To measure the response of the system, we recorded

### 2012-10-13 manuscript.tex
Various small edits.

@@
 target on its own, we used the ground station to provide remote headings for
-navigation to the window. Ground truth data was collected via a motion capture
-system in the test room. A sketch of the experimental setup is shown in
+navigation to the window. A motion capture system in the test room provided
+ground truth data. A sketch of the experimental setup is shown in
 Fig.~\ref{fig:experiment_cartoon}.
@@
 simulation so that paths from known positions looked similar to the behavior
-we observed in the experimental data. We also inlcuded the camera capture and
+we observed in the experimental data. We also account for the camera capture and
 pose estimation delays, and the PID controller sampling rate in the
@@
 The results of our experimental trials verify the results of the Monte Carlo
-simulation of the model that we used. The trials that we conducted in
+simulation of our model. The trials that we conducted in
 Fig.~\ref{fig:flight_paths_feasible} correspond to the circled region in
-Fig.~\ref{fig:heat_map}. The 80\% success rate that we experienced in the
+Fig.~\ref{fig:heat_map}. The 80\% success rate that we measured in the
 experiments corroborates our calculated probability of success of

### 2012-10-13 manuscript.tex
Conclusion draft.

@@
 \section{Conclusions and Future Work}
-%TODO 
+We have demonstrated cooperative guidance of a 13 gram flapping wing MAV with a
+lightweight base station in hardware. In addition, we produced a model of the
+overall system by modeling the individual system capabilities, and composing
+it with known system interactions. The corroboration between our simulated and
+experimental results suggests that our simplified modeling approach may be useful for 
+predicting the performance of collaborative systems.
+
+We intend to further explore the benefits and limitations of collaboration in
+unstructured environments by deploying our base station on millirobotic ground
+platforms. As the base station used in this work was designed with this avenue
+in mind, this is a matter of implementation. The effectiveness of our modeling
+approach when applied to systems with more interactions and agents is also
+a future area of interest.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-10-13 manuscript.tex
text changes

--- 2012-10-13 manuscript.tex
@@
 
-To model the translational position of the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state.
+\begin{equation}
+\label{eq:dubin_car}
+\begin{aligned}
+& \dot{x} = u_{s}sin(\theta)\\
+& \dot{y} = u_{s}cos(\theta)\\
+& \dot{\theta} = u\\
+& u \in U=[-\tan{\phi_{max}},\tan{\phi_{max}}]
+\end{aligned}
+\end{equation}
+
+To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.
+~\ref{eq:dubin_car})~\cite{lavalle:planning}, where $\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed.
+We modify this model to treat the angular position $\theta$ as an input rather than a state.
 The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.
@@
 & \dot{x} = sin(\theta)\\
-& \dot{y} = cos(\theta)\\
+& \dot{y} = cos(\theta)
 \end{aligned}
@@
 relation reflects the fact that the camera cannot determine the depth of the
-robot.
+robot. One example of why this is a problem is that if the robot stays at a
+constant horizontal distance from the window, but moves closer and closer to
+the camera, the directed heading steadily increases even though the robot has
+not actually moved farther from the window. This creates tracking problems in
+areas close to the camera.
 
@@
 
-%............................................................................%
-\subsubsection{Cooperative Visual Servoing}
-
 %----------------------------------------------------------------------------%
 \subsection{Model Verification}
+\label{sec:experiments:verification}
 
 We tested our cooperative visual servoing system with a simple target seeking
-experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
-a wooden frame meant to simulate a window. As the H$^2$Bird cannot detect the
+experiment: H$^2$Bird must fly through a specified obstacle, in this case,
+a wooden frame meant to simulate a window. As H$^2$Bird cannot detect the
 target on its own, we used the ground station to provide remote headings for
-navigation to the window. A motion capture system in the test room provided
+navigation to the window. A VICON~\footnote{Vicon:~\url{http://www.vicon.com/}} motion capture system in the test room provided
 ground truth data. A sketch of the experimental setup is shown in
-Fig.~\ref{fig:experiment_cartoon}.
+Fig.~\ref{fig:experiment_cartoon} and a screen capture from the camera perspective is in Fig.~\ref{fig:pf_screencap}, pane 1.
 
@@
 \begin{enumerate}
-\item The window is identified by the user. 
-\item The H$^2$Bird is released in view of the camera by hand at the desired starting grid point.
+\item The window is identified by the user (Fig.~\ref{fig:pf_screencap}, pane 2). 
+\item The H$^2$Bird is released in view of the camera by hand at the desired starting grid point (Fig.~\ref{fig:pf_screencap}, pane 3).
 \item The H$^2$Bird attempts to fly to the window with remote guidance from
-the base station.
+the base station (Fig.~\ref{fig:pf_screencap}, panes 4,5,6).
 \end{enumerate}
@@
 %TODO cameron
-To bound the performance of the cooperative control system, we determine 
-the feasible set of initial conditions that result in successful narrow 
-passage traversal using both analytical and empirical methods. We conducted 
-experiments to validate this model.
-
-The following sections detail how we calculated the feasible and infeasible
-regions using system identification and present the results of our
-experimental trials.
+The results of our system identification of the cooperative visual servoing model are detailed in the following sections.
+We use the model to empirically and analytically predict the feasibility of the cooperative control system for various initial conditions.
+Furthermore, we validate the model using experimental trials.
 
@@
 
-The results of the step response experiment are described in
-Fig.~\ref{fig:step_response}. The first-order model that we fit to the
-response is in Equation~\ref{eq:transfer_func}.
+The first-order model that we fit to the
+step response of the onrithopter yaw motion is in Equation~\ref{eq:transfer_func}.
+This model represents the tuning dynamics of the ornithopter as well as the PID controller providing inputs to the actuators.
+The results of the step response experiment along with the response of the fitted model are depicted in
+Fig.~\ref{fig:step_response}.
 
@@
 
-The step response rise time is 1.4 seconds. The H$^2$Bird flies
+The step response rise time is 1.4 seconds. H$^2$Bird flies
 at an average forward velocity of 1.2 m/s during our experiments, so the
-estimated minimum turning radius of the robot is 1.07 maters, as described in
-Equation~\ref{eq:min_radius}. In the experiments, we choose a flight speed
-lower than the maximum speed of the ornithopter, to facilitate more robust
+estimated minimum turning radius of the robot is 1.07 meters, as described in
+Equation~\ref{eq:min_radius}. In the experiments, we choose a slower flight speed
+than the maximum speed of the ornithopter, to facilitate more robust
 tracking.
@@
 
-The results of our experimental trials are shown in
-Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
+Conducting the experimental trials as described in Section~\ref{sec:experiments:verification}, we achieved a success rate of 80\% for
 initial conditions in the middle of the testing space and in front of the
-camera, we achieved a success rate of 80\%. This case is depicted in
-Fig.~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminishes
-to 50\% because of several factors. On the edges of the camera frame, the yaw
+camera. As we moved towards the edges of the
+feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminished
+to 50\%. The results of our experimental trials are shown in
+Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+
+The lower success rate on the edges of the camera frame can be explained by several factors. On the edges of the viewing triangle, the yaw
 control inputs are large in magnitude, although this is not necessary in all
@@
 \centering
-\includegraphics[width=\textwidth]{figures/heat_map.pdf}
+\includegraphics[width=\textwidth]{figures/heat_map.png}
 \caption{Probability of success for all starting locations. This graphic is 
@@
 
-We simulated the motion of the ornithopter beginning from certain positional
-and angular initial conditions using several components representing the
-experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
-For the "ornithopter dynamics" block, we used the translational model
-described by~\ref{eq:total_model_plain} with the angular position model
-described by~\ref{eq:transfer_func}. We tuned the PID controller for the
-simulation so that paths from known positions looked similar to the behavior
-we observed in the experimental data. We also account for the camera capture and
-pose estimation delays, and the PID controller sampling rate in the
-simulation.
+We simulated the motion of the ornithopter beginning from various positional
+and angular heading initial conditions.
+We used the model that we determined through system identification (Sec.~\ref{sec:system_model}) using the cooperative control method in Fig.~\ref{fig:block_diagram}.
 
-We determine the backwards reachable set of initial states for successful
+We determined the backwards reachable set of initial states for successful
 narrow passage traversal both geometrically and in simulation
-(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform
-grid of 0.05 meter increments within the camera viewing region, and an initial
-angular position perpendicular to the narrow passage plane. The geometrically
+(Fig.~\ref{fig:feasible_set}). In the figure, the blue lines represent the camera viewing triangle, the black line
+represents the window, and the black dot represents the camera. For the initial conditions, we used a uniform
+grid of 5 cm increments within the camera viewing region, and an initial
+angular heading perpendicular to the window plane. A success, or feasible initial condition, consists
+of a path that intersects the window at some point and does not leave the
+camera viewing triangle, and is indicated in green. The geometrically
 infeasible region is computed using the minimum turning radius and the
-geometry of the camera's field-of-view and is indicated in blue. Both the
-regions in red and blue indicate the infeasible region determined in
-simulation. The feasible region is highlighted in green. A success consisted
-of a path that intersects the window at some point and does not leave the
-camera viewing triangle. Fig.~\ref{fig:feasible_set} is validated by our
-experimental data, as there are no successes within the geometrically
+geometry of the camera's field-of-view and is indicated in blue. The region in red is not geometrically feasible, but is infeasible as a result of the cooperative control simulation dynamics.
+
+Our experimental data validates Fig.~\ref{fig:feasible_set}, as there are no successes within the geometrically
 infeasible region. There are a few successes near the boundary of the
-infeasible region determined in simulation, but this phenomenon can be
-explained by the results in Fig.~\ref{fig:heat_map}.
+infeasible region determined in simulation.
+While there are successes along the boundary of the camera viewing triangle, there are also many failures.
+We further explore this behavior using Monte Carlo simulation.
 
-Fig.~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
-simulating possible paths of the H$^2$Bird during our experiments. For each
-point within a 0.2 meter grid within the camera viewing triangle, we simulated
-10 trials using an initial heading randomly sampled from a uniform
-distribution between -90$^{\circ}$ and 90$^{\circ}$. As shown in the figure,
+We conducted a Monte Carlo simulation to determine the probability of successful window navigation for a given point within the testing area.
+Using this information we can make assumptions about ideal starting positions for coopertive control.
+For each point within a 0.1 meter grid within the camera viewing triangle, we simulated
+40 trials using an initial heading randomly sampled from a uniform
+distribution between -90$^{\circ}$ and 90$^{\circ}$.
+Fig.~\ref{fig:heat_map} depicts the results of the Monte Carlo simulation. As shown in the figure,
 there is a region in the middle with a 100\% probability of success, which is

### 2012-10-13 manuscript.tex
Integrate comments from paper walkthrough

--- 2012-10-13 manuscript.tex
@@
 less mobility than MAVs, but relaxed size constraints and typically more 
-computing power, creating opportunities for cooperation. In this work, we 
+computing power. These specializations present many opportunities for cooperation. In this work, we 
 demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
 and a lightweight ground station using computer vision.
+We develop models for the ornithopter, ground station, and cooperative system dynamics.
+We determine model parameters of the real systems through experimental system identification.
+Finally, we verify those models using experiments on narrow passage traversal, and arrive at a cooperative system model which accurately predicts the backwards-reachable region for successfully negotiating ornithopther flight through narrow openings.
 
@@
 payload of 2.8 grams. We augment the ornithopter's built-in 
-gyroscope-based control with a ground station, which has power and weight 
+gyroscope-based control with a lightweight ground station design, which has power and weight 
 requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
@@
 real-time motion tracking algorithm on a live video stream. This allows 
-the ornithopter to reliably negotiate flight through narrow openings.
-
-We demonstrate the performance of the location estimation and navigation 
-algorithms. We model the backwards-reachable region for successfully 
-navigating through narrow openings. Finally, we verify this model through 
-with experiments with narrow passage traversal.
+the ornithopter to reliably negotiate flight through narrow openings. We demonstrate the performance of the location estimation and navigation  algorithms.
 \end{abstract}
@@
 \end{figure}
-Power, size, and weight constraints for micro air vehicle (MAV) 
-platforms are amongst the most restrictive of all robotic platforms. With
-significantly limited sensing and computing capabilities, only the most rudimentary
-control and state estimation algorithms can be executed on-board these platforms. 
-Even more restrictive are the constraints for flapping wing MAVs, 
-a class of MAVs that exhibit demonstrate safety and noise profiles superior to those
-of rotor craft MAVs, but at the expense of reduced payload weight.
+Power, size, and weight constraints on micro air vehicles (MAV) 
+significantly limit sensing their and computational capabilities. They can execute only the most rudimentary
+control and state estimation algorithms in real-time. 
+The constraints on flapping-wing MAVs are even more restrictive. 
+Though they exhibit demonstrate safety and noise profiles superior to those
+of rotor craft MAVs, this comes at the expense of reduced payload weight.
 
-Compounding these engineering restrictions are the many non-linear
-flight dynamics exhibited by flapping wing flight~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
-Much past research on flapping wing MAVs has been focused on overcoming the 
-difficulties associated with controlling and modeling in the face of these 
-non-linearities\cite{floreano:flying}. Common approaches to simplify these 
+The non-linear
+flight dynamics exhibited of flapping wing flight compounding constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
+Much past research on flapping wing MAVs focused on overcoming the 
+difficulties associated with controlling and modeling flapping-wing vehicles in the face of this
+non-linearity\cite{floreano:flying}. Common approaches to simplify these 
 dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
-\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict the 
-behavior in flight~\cite{Shigeoka:ornithopter}. Advances have lead to the
+\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict flight behavior~\cite{Shigeoka:ornithopter}. These advances lead to the
 development of some high-performance platforms, such as the sub-15 gram DelFly 
-as described by deCroon et al.~\cite{delfly:design}.
+described by deCroon et al.~\cite{delfly:design}.
 
-Many of the control laws that previous researches have developed from the
+Many control laws that previous researchers developed from
 dynamic models of flapping-winged flight, such as periodic forcing and other
-control methods as described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
-in~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, are computationally complex and
-require powerful on-board computing. In our work, we develop a simpler, less
-computationally intensive method of control for negotiating narrow passages.
+control methods described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
+by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
+require powerful on-board computing. In this work, we develop a simpler, less
+computationally intensive method of control for negotiating narrow passages, by exploiting cooperation between specialized robotic agents.
 
-In contrast, there is considerably less previous work on guidance of flapping wing MAVs. 
+In contrast to flight control, there is considerably less previous work on guidance and of flapping-wing MAVs. 
 Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
@@
 processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
-\cite{baek:tracking} is one of the few examples of scompletely autonomous target seeking.
+\cite{baek:tracking} provides one of the few examples of completely autonomous target seeking for vehicles at this scale.
 
-There also exists a body of work on guidance and control for fixed wing and rotary wing 
+There also exists a body of work on the guidance and control of fixed- and rotary-wing 
 MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
-~\cite{kingston:timeattitude}\cite{kanade:3dvision}. While operating indoors, using
-GPS to estimate the pose of the MAV is not possible, and motion capture
-requires expensive and stationary modification of the environment. The power
-and payload limits on flapping wing MAVs make on-board sensing difficult, so multi-agent
+~\cite{kingston:timeattitude}\cite{kanade:3dvision}. Indoors, using
+GPS to estimate the pose of an MAV is impossible, and motion capture
+requires extensive and stationary modification of the environment. The power
+and payload limits on flapping wing MAVs make complex on-board sensing difficult, so multi-agent
 cooperative control schemes are apt solutions.
@@
 % humphrey
-To fulfill the need for a lightweight, autonomous flapping wing MAV platform, 
-we designed the H$^2$Bird. Built around the Silverlit i-Bird 
+To explore these cooperative control concepts, we designed a new flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{h2bird}). Built around the Silverlit i-Bird 
 RC flyer power train and clap-fling 
@@
 tail propeller and servo-controlled tail elevator provide control in the yaw 
-and pitch axes. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
+and pitch axes (Fig.~\ref{h2Bird_axes}. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
 \href{https://github.com/biomimetics/imageproc\_pcb}
      {https://github.com/biomimetics/imageproc\_pcb}} 
-electronics package consists of a 40 MIPS microprocessor, 6 DOF IMU, 
+controller consists of a 40 MIPS microprocessor, 6 DOF IMU, 
 IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
 lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
-ground speed, and can operate for approximately 10 minutes.
-%TODO footnote for ImageProc
+ground speed, and operates for approximately 10 minutes.
 
@@
 BeagleBoard single-board computer as reference computation platform. The
-BeagleBoard consumes approximately 1 watt while running our computer vision 
+BeagleBoard consumes approximately one watt while running our computer vision 
 and control algorithms. It uses the same processor as popular sub-10 gram 
@@
 quality and resolution to the miniature camera modules available at 
-millirobot scale. The BeagleBoard communicates with the H$^2$Bird using a 
+millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a
 USB radio module which implements the IEEE 802.15.4 wireless standard.
@@
 Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
-modified with a custom version of the Linux kernel. We use the Python
-\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
+modified with a custom version of the Linux kernel. We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
 {http://www.python.org/}} programming language for control, robot communication, 
@@
 
-Attitude estimation and control for the H$^2$Bird runs on-board to achieve 
-high sample rate attitude control. We estimate the vehicle pose by
-integrating the IMU, and use a proportional plus integral plus derivative (PID) 
-controller for controlling the vehicle.
+We implement attitude estimation and control of the H$^2$Bird on-board to achieve 
+high sample-rate attitude control. We estimate the vehicle pose by
+integrating the IMU, and use a proportional-integral-derivative (PID) 
+controller to feed these estimates back to the attitude control flight surfaces (Figs.\ref{process_flow} and~\ref{block_diagram}).
 
-The high pitch angles flapping wing flyers achieve in routine flight distinguish them 
+The high pitch angles flapping-wing MAVs regularly operate at distinguish them 
 from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
-parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of the vehicle orientation 
-suffer from singularities around high pitch angles, so we instead 
+parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of vehicle orientation 
+suffer from singularities around high pitch angles. So we instead 
 represent the vehicle orientation with quaternions. An additional benefit of 
@@
 \end{align}
-We then convert this error quaternion to an angle axis representation and use this 
+We then convert this error quaternion to an angle axis representation and use it 
 as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
-These calculations are performed onboard with trigonometric lookup tables for speed.
+We perform these calculations on-board, with trigonometric lookup tables for real-time performance.
 \begin{align}
@@
 
-Motion of the H$^2$Bird in the frame is tracked using a particle filter
-with an approach similar to that employed by Thrun~\cite{thrun2005probabilistic}. 
-Particles are distributed uniformly across the frame on initialization. 
+Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a particle filter ~\cite{thrun2005probabilistic}. 
+We initialize the particles uniformly across the frame before tracking begins. 
 To improve numerical stability, we normalize the weights of all particles on 
@@
 $\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
-$x^{[i]}_t$ is the state of the $i^\text{th}$ particle at time $t$, $\Sigma$ 
-is a chosen covariance matrix, and $b$ is vector of the height and width of 
-the camera frame.
+$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at time $t$, $\Sigma$ 
+is a chosen covariance matrix, and $b$ is vector of the bounds of the pixel space.
 \begin{equation}
@@
 \end{equation}
-The experiments presented in this work use a diagonal covariance matrix 
-with a variance of 256, and an $\epsilon$ of 0.3. We find that the Gaussian 
+In the experiments we present in this work, we used a diagonal covariance matrix 
+with a uniform variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
 transition model is sufficient for tracking motion in which frame-to-frame 
@@
 transitions increase tracking robustness, by forcing the filter to always 
-sample all of the pixel space. This prevents overconfidence and collapse, 
+sample broadly across of the pixel space. This helps prevent overconfidence and collapse, 
 and decreases reacquisition delay.
@@
 Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
-approach to motion tracking via na\"{i}ve background subtraction[CITATION].
+approach to motion tracking via na\"{i}ve background subtraction, e.g. frame differencing[CITATION].
+On each iteration, the system provides the emission model with $f_0$, the frame representing 
+the background and $f_t$, the most recent frame captured. The model assigns 
+a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
+distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
+$x^{[i]}_t$.
 \begin{equation}
@@
 \end{equation} 
-The system provides the emission model with $f_0$, the frame representing 
-the background and $f_t$, the most recent frame captured. The model assigns 
-a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
-distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
-$x^{[i]}_t$. Hence, particles are more likely to be resampled in portions of 
-the image which appear different from the background.
+Hence, particles are more likely to be resampled in regions of 
+the image which are dissimilar in color from the background.
 
-For the experiments presented in this work, we assumed that neither the bird 
-nor any other moving object were visible in the first frame captured on system 
+During experiments presented in this work, we assumed that neither the bird, 
+nor any other moving object, were visible in the frame on system 
 start. We saved the first frame captured by the system, and provided it as 
-the background frame $f_0$ for every iteration of the particle filter. This 
+the background frame $f_0$ for every iteration of the particle filter. This choice of $f_0$ 
 resulted in a sampling approach to na\"{i}ve background subtraction.
@@
 Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
-time[CITATION]. However, we found it very efficient and sufficiently reliable for the 
+time[CITATION]. However, we find it very efficient and sufficiently reliable for the 
 short time spans necessary for H$^2$Bird to successfully negotiate narrow 
 openings, which are typically less than five seconds. Substituting $f_0$ 
-with a more sophisticated extraction of the background would implement a 
-more sophisticated approach to background subtraction, with no changes to 
-the underlying motion tracking filter.
+with a more sophisticated background extraction would retrain the underlying motion tracking features of the filter, while providing it with a more robust way of differentiating subject motion from the background.
 
@@
 \label{sec:visual_servoing_concept}
-The H$^2$Bird is remotely guided by commanding relative changes in its
-reference orientation. Using relative changes removes the need to estimate the relative
-orientation between the ground station and vehicle frames of reference. 
-With estimates of the vehicle pose, we calculate an altitude
-error and horizontal error. The ground station then commands incremental changes to 
-the H$^2$Bird pitch and yaw are according to a PID control law, with positive altitude error
-corresponding to negative pitch increments and positive horizontal error
-corresponding to negative yaw increments. We window the controller integrator term 
-to prevent windup when the bird is being launched, and tuned the derivative term
-to dampen oscillations associated with the system latency [?].
+The ground station guides the H$^2$Bird by commanding changes to its
+reference attitude, as governed by its on-board attitude controllers (Figs.\ref{process_flow} and~\ref{block_diagram}). 
+With the estimate of 2D vehicle pose from computer vision, we calculate an altitude
+error and horizontal error. The ground station then commands a reference attitude to 
+the H$^2$Bird pitch and yaw. according to a PID control law. We apply a fixed-lag window the PID controller integrator term 
+to prevent wind-up while we launch the ornithopter. We tuned the proportional and derivative terms of this outer controller 
+to dampen the overshoot and drift cause by the cooperative system latency.
 
@@
 \end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{System Model}
+\label{sec:system_model}
+
+%----------------------------------------------------------------------------%
+\subsection{Ornithopter Attitude Control}
+\label{sec:model_attitude}
+
+We divide our model of the vehicle pose into two independent models for simplicity: one for
+yaw motion, and one for translational motion.
+
+We approximate the yaw dynamics of the ornithopter with a one-dimensional model to limit complexity and because it matches our physical implementation.
+Likewise, we choose the reference heading angle as input to the model and model the dynamics of vehicle and on-board PID controller as one system, which outputs the physical heading angle.
+We show the resulting model form in Equation~\ref{eq:plain_model}.
+
+\begin{equation}
+\label{eq:plain_model}
+H(s) = \frac{K}{1 + Rs}
+\end{equation}
+
+To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.~\ref{eq:total_model_plain})~\cite{lavalle:planning}. 
+\begin{equation}
+\label{eq:total_model_plain}
+\begin{aligned}
+& \dot{x} = sin(\theta)\\
+& \dot{y} = cos(\theta)\\
+\end{aligned}
+\end{equation}
+Our form uses the angular position, $\theta$ as an input rather than a state.
+We model the overall motion of the system as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (Eq.
+\ref{eq:plain_model}) with the reference heading angle as the input.
+
+%----------------------------------------------------------------------------%
+\subsection{Ground Station Pose Estimation}
+
+Our model of the ground station comprises models of both the camera and the
+pose estimation computation. The camera model contains both a
+latency, which is in series with the path between the ``camera'' and ``ornithopter dynamics''
+ blocks in Fig.~\ref{fig:block_diagram}, and a model of the camera's viewing angle constraints.
+
+ We model the 
+viewing angle of the camera as a pixel map from one edge of a
+63$^{\circ}$ viewing triangle to the other. The camera has an image width of
+640 pixels. We calculate the desired heading angle in degrees for a given distance from the
+camera according to Equation~\ref{eq:desired_heading}, where $d$
+is the distance between the ornithopter and the camera, and $x$ is the horizontal
+position of the robot relative to the center of the window in meters. This
+explicitly fact that the pose estimation algorithm does not estimate the depth of the
+robot.
+\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
+\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
+\end{equation}
+
+We model the pose estimation computation by the by the latency of running 
+the particle filter, calculating a new reference attitude, and sending to the ornithopter via radio.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}
+
+%----------------------------------------------------------------------------%
+\subsection{System Identification}
+
+%............................................................................%
+\subsubsection{Ornithoper Attitude Control}
+
+We determined the turning speed and turning radius of the H$^2$Bird
+experimentally, by measuring the response of the system to a step input of a
+clockwise 90$^{\circ}$ turn. We measured the response of the system by recording
+the attitude estimates calculated by the ornithopter using its on-board
+gyroscope. We then used
+MATLAB\footnote{The MathWorks, Inc. Matlab:
+\href{http://www.mathworks.com/products/matlab/}
+     {http://www.mathworks.com/products/matlab/}} to fit a
+simple low-order model to the step response, as discussed in Sec.~\ref{sec:model_attitude}.
+
+%............................................................................%
+\subsubsection{Ground Station Pose Estimation}
+
+To measure the capture latency of the camera, we wrote a simple test 
+application which changes a large colored rectangle on the ground station's screen, then uses the camera to detect that change.
+By noting the value of the system timer just before we change the screen, and just after the camera detects that change, the application arrives at an estimate of the video capture pipeline latency.
+We averaged these result over several hundred rapidly-executed trials to determine the capture latency we used to model the system..
+
+To compute the particle filter and controller latency, we again used system timers to determine the
+average time between the start and end of computation.
+
+%----------------------------------------------------------------------------%
+\subsection{Model Verification}
+
+We tested our cooperative visual servoing system with a simple target seeking
+experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
+a wooden frame we used to simulate a window. As the H$^2$Bird cannot detect the
+target on its own, we used the cooperative control with the ground station, which to provided remote guidance to H$^2$Bird to allow it to navigate to the window.
+We used a motion capture system to record 
+ground truth data. We present a sketch of this experimental setup in
+Fig.~\ref{fig:experiment_cartoon}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{System Model}
-\label{sec:system_model}
-
-%----------------------------------------------------------------------------%
-\subsection{Ornithopter Attitude Control}
-\label{sec:model_attitude}
-
-Modeling of the vehicle pose is divided into two independent models for simplicity: one for
-yaw motion, and one for translational motion.
-
-We model the yaw motion of the ornithopter with a one-dimensional model to reduce the complexity of the overall model.
-Matching our physical implementation, we choose the reference heading angle as an input and model the dynamics from the vehicle and on-board PID controller to output the physical heading angle.
-The model format is shown in Equation~\ref{eq:plain_model}.
-
-\begin{equation}
-\label{eq:plain_model}
-H(s) = \frac{K}{1 + Rs}
-\end{equation}
-
-To model the translational position of the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning} with the angular position, $\theta$ as an input rather than a state.
-The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.
-
-\begin{equation}
-\label{eq:total_model_plain}
-\begin{aligned}
-& \dot{x} = sin(\theta)\\
-& \dot{y} = cos(\theta)\\
-\end{aligned}
-\end{equation}
-
-%----------------------------------------------------------------------------%
-\subsection{Ground Station Pose Estimation}
-
-The model of the ground station combines both the camera model and the
-computational model for pose estimation. The camera model contains both a
-latency, which would occur in between the "camera" and "ornithopter dynamics"
-in Fig.~\ref{fig:block_diagram}, and a model for the viewing angle. We model the 
-viewing angle for the camera as a pixel map from one edge of a
-63$^{\circ}$ viewing triangle to the other. The camera has an image width of
-640 pixels. The desired heading angle in degrees for a given distance from the
-camera is calculated according to Equation~\ref{eq:desired_heading}, where $d$
-is the distance of the robot in front of the camera and $x$ is the horizontal
-position of the robot relative to the center of the window in meters. This
-relation reflects the fact that the camera cannot determine the depth of the
-robot.
-
-\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
-\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
-\end{equation}
-
-The computational model accounts for the latency added by the operations required to
-run the particle filter and calculate the input to send to the robot.
-
-%----------------------------------------------------------------------------%
-\subsection{Cooperative Visual Servoing}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
-
-%----------------------------------------------------------------------------%
-\subsection{System Identification}
-
-%............................................................................%
-\subsubsection{Ornithoper Attitude Control}
-
-We determined the turning speed and turning radius of the H$^2$Bird
-experimentally by measuring the response of the system to the step input of a
-clockwise 90 degree turn. To measure the response of the system, we recorded
-the angular position data of the ornithopter calculated from the on-board
-gyroscope data. We then used
-MATLAB\footnote{The MathWorks, Inc. Matlab:
-\href{http://www.mathworks.com/products/matlab/}
-     {http://www.mathworks.com/products/matlab/}} to fit a
-simple low-order model to the step response as discussed in Sec.~\ref{sec:model_attitude}.
-
-%............................................................................%
-\subsubsection{Ground Station Pose Estimation}
-
-To measure the capture latency of the camera, we alternatively flashed black
-and white color in front of the camera and measured the time lapse before it
-detected the color change. We averaged this result over 50 trials to determine
-the capture latency.
-
-To compute the particle filter latency, we used system timers to determine the
-average time from the start of computation to the end of computation.
-
-%............................................................................%
-\subsubsection{Cooperative Visual Servoing}
-
-%----------------------------------------------------------------------------%
-\subsection{Model Verification}
-
-We tested our cooperative visual servoing system with a simple target seeking
-experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
-a wooden frame meant to simulate a window. As the H$^2$Bird cannot detect the
-target on its own, we used the ground station to provide remote headings for
-navigation to the window. A motion capture system in the test room provided
-ground truth data. A sketch of the experimental setup is shown in
-Fig.~\ref{fig:experiment_cartoon}.
-
-% TODO clarify or roll into paragraph
-Each trial included the following steps:
+We performed each trial as follows (see Fig.~\ref{fig:pf_screencap}):
 \begin{enumerate}
 \item The window is identified by the user. 
-\item The H$^2$Bird is released in view of the camera by hand at the desired starting grid point.
-\item The H$^2$Bird attempts to fly to the window with remote guidance from
-the base station.
+\item We release H$^2$Bird by hand, in view of the camera at the desired starting grid point.
+\item The H$^2$Bird attempts to fly through the window with, with remote guidance from
+the ground station.
 \end{enumerate}
 
-In total, we ran 80 trials over a variety of starting positions. The goal of
-our experiments was to verify the model described in
+In total, we ran 80 trials over a variety of starting positions. Our goal for the experiments was to verify the model described in
 Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
-along the edges of the camera viewplane in 0.05 meter increments from the
-window to determine the success rate on the edges of the testing space. We
+along the edges of the camera view plane, and in 0.05 meter increments from the
+window, in an effort to determine the success rate on the edges of the testing space. We
 conducted 20 additional trials directly in front of the camera to determine
-the success rate near the center of the testing space. After every 5 trials,
-we replaced the current battery with a fully charged one.
+the success rate near the center of viewing plane. After every 5 trials,
+we replaced the ornithopter's battery with a fully charged one.
 
@@
 \end{figure}
-%TODO label experiment cartoon
 
@@
 \label{sec:performance}
-%TODO cameron
 To bound the performance of the cooperative control system, we determine 
 the feasible set of initial conditions that result in successful narrow 
-passage traversal using both analytical and empirical methods. We conducted 
+passage traversal, using both analytical and empirical methods. We then conducts 
 experiments to validate this model.
@@
 \label{sec:flight_control}
-%TODO cameron
 \begin{figure}[tb]
@@
 The step response rise time is 1.4 seconds. The H$^2$Bird flies
-at an average forward velocity of 1.2 m/s during our experiments, so the
-estimated minimum turning radius of the robot is 1.07 maters, as described in
-Equation~\ref{eq:min_radius}. In the experiments, we choose a flight speed
-lower than the maximum speed of the ornithopter, to facilitate more robust
-tracking.
-
+at an average forward velocity of 1.2 m/s during our experiments, so we
+estimate the minimum turning radius of the robot is 1.07 maters, as described in
+Equation~\ref{eq:min_radius}.
 \begin{equation}
@@
 \end{equation}
+In the experiments presented in this work, we choose a flight speed
+lower than the maximum speed of the ornithopter, to facilitate more robust
+tracking.
 
@@
 
-The results of our experimental trials are shown in
+We present the results of our experimental trials in
 Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
@@
 
-We simulated the motion of the ornithopter beginning from certain positional
+We simulated the motion of the ornithopter, beginning from certain positional
 and angular initial conditions using several components representing the
 experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
-For the "ornithopter dynamics" block, we used the translational model
-described by~\ref{eq:total_model_plain} with the angular position model
+For the ``ornithopter dynamics'' block, we used the translational model
+described by~\ref{eq:total_model_plain}, with the angular position model
 described by~\ref{eq:transfer_func}. We tuned the PID controller for the

### 2012-10-13 manuscript.tex

@@
 \end{figure}
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
+\caption{Frame sequence from video feed used for tracking. Particles 
+are shown in \textcolor{red}{red}, the current state estimate as a 
+\textcolor{yellow}{yellow} circle, the window bounding box in
+\textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
+circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
+dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
+Particles converge on its location; 4.~Release H$^2$Bird, which 
+begins cooperative autonomous flight toward the target; 5.~Visual servoing 
+controller overshoots the window on the left, and commands a right-turn 
+maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
+\label{fig:pf_screencap}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.~\ref{eq:total_model_plain})~\cite{lavalle:planning}. 
 \begin{equation}
-\label{eq:total_model_plain}
+\label{eq:dubin_car}
 \begin{aligned}
-& \dot{x} = sin(\theta)\\
-& \dot{y} = cos(\theta)\\
+& \dot{x} = u_{s}sin(\theta)\\
+& \dot{y} = u_{s}cos(\theta)\\
+& \dot{\theta} = u\\
+& u \in U=[-\tan{\phi_{max}},\tan{\phi_{max}}]
 \end{aligned}
 \end{equation}
+
+To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.
+~\ref{eq:dubin_car})~\cite{lavalle:planning}, where $\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed.
+We modify this model to treat the angular position $\theta$ as an input rather than a state.
+The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.
+
 Our form uses the angular position, $\theta$ as an input rather than a state.
@@
 position of the robot relative to the center of the window in meters. This
-explicitly fact that the pose estimation algorithm does not estimate the depth of the
-robot.
+relation reflects the fact that the camera cannot determine the depth of the
+robot. One example of why this is a problem is that if the robot stays at a
+constant horizontal distance from the window, but moves closer and closer to
+the camera, the directed heading steadily increases even though the robot has
+not actually moved farther from the window. This creates tracking problems in
+areas close to the camera.
+
 \begin{equation} \label{eq:desired_heading} \text{Desired heading} =
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/pf_screencap.pdf}
-\caption{Frame sequence from video feed used for tracking. Particles 
-are shown in \textcolor{red}{red}, the current state estimate as a 
-\textcolor{yellow}{yellow} circle, the window bounding box in
-\textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
-circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
-dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
-Particles converge on its location; 4.~Release H$^2$Bird, which 
-begins cooperative autonomous flight toward the target; 5.~Visual servoing 
-controller overshoots the window on the left, and commands a right-turn 
-maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
-\label{fig:pf_screencap}
-\end{figure}
-
-We performed each trial as follows (see Fig.~\ref{fig:pf_screencap}):
+% TODO clarify or roll into paragraph
+Each trial included the following steps:
 \begin{enumerate}
@@
 \label{sec:performance}
+%TODO cameron
 To bound the performance of the cooperative control system, we determine 
 the feasible set of initial conditions that result in successful narrow 
-passage traversal, using both analytical and empirical methods. We then conducts 
+passage traversal using both analytical and empirical methods. We conducted 
 experiments to validate this model.
@@
 \label{sec:flight_control}
+%TODO cameron
 \begin{figure}[tb]
@@
 
-We present the results of our experimental trials in
-Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}. For
+Conducting the experimental trials as described in Section~\ref{sec:experiments:verification}, we achieved a success rate of 80\% for
 initial conditions in the middle of the testing space and in front of the
-camera, we achieved a success rate of 80\%. This case is depicted in
-Fig.~\ref{fig:flight_paths_feasible}. As we moved towards the edges of the
-feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminishes
-to 50\% because of several factors. On the edges of the camera frame, the yaw
+camera. As we moved towards the edges of the
+feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminished
+to 50\%. The results of our experimental trials are shown in
+Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+
+The lower success rate on the edges of the camera frame can be explained by several factors. On the edges of the viewing triangle, the yaw
 control inputs are large in magnitude, although this is not necessary in all
@@
 
-We simulated the motion of the ornithopter, beginning from certain positional
-and angular initial conditions using several components representing the
-experimental layout of our system described in Fig.~\ref{fig:block_diagram}.
-For the ``ornithopter dynamics'' block, we used the translational model
-described by~\ref{eq:total_model_plain}, with the angular position model
-described by~\ref{eq:transfer_func}. We tuned the PID controller for the
-simulation so that paths from known positions looked similar to the behavior
-we observed in the experimental data. We also account for the camera capture and
-pose estimation delays, and the PID controller sampling rate in the
-simulation.
+We simulated the motion of the ornithopter beginning from various positional
+and angular heading initial conditions.
+We used the model that we determined through system identification (Sec.~\ref{sec:system_model}) using the cooperative control method in Fig.~\ref{fig:block_diagram}.
 
-We determine the backwards reachable set of initial states for successful
+We determined the backwards reachable set of initial states for successful
 narrow passage traversal both geometrically and in simulation
-(Fig.~\ref{fig:feasible_set}). For the initial conditions, we use a uniform
-grid of 0.05 meter increments within the camera viewing region, and an initial
-angular position perpendicular to the narrow passage plane. The geometrically
+(Fig.~\ref{fig:feasible_set}). In the figure, the blue lines represent the camera viewing triangle, the black line
+represents the window, and the black dot represents the camera. For the initial conditions, we used a uniform
+grid of 5 cm increments within the camera viewing region, and an initial
+angular heading perpendicular to the window plane. A success, or feasible initial condition, consists
+of a path that intersects the window at some point and does not leave the
+camera viewing triangle, and is indicated in green. The geometrically
 infeasible region is computed using the minimum turning radius and the
-geometry of the camera's field-of-view and is indicated in blue. Both the
-regions in red and blue indicate the infeasible region determined in
-simulation. The feasible region is highlighted in green. A success consisted
-of a path that intersects the window at some point and does not leave the
-camera viewing triangle. Fig.~\ref{fig:feasible_set} is validated by our
-experimental data, as there are no successes within the geometrically
+geometry of the camera's field-of-view and is indicated in blue. The region in red is not geometrically feasible, but is infeasible as a result of the cooperative control simulation dynamics.
+
+Our experimental data validates Fig.~\ref{fig:feasible_set}, as there are no successes within the geometrically
 infeasible region. There are a few successes near the boundary of the
-infeasible region determined in simulation, but this phenomenon can be
-explained by the results in Fig.~\ref{fig:heat_map}.
+infeasible region determined in simulation.
+While there are successes along the boundary of the camera viewing triangle, there are also many failures.
+We further explore this behavior using Monte Carlo simulation.
 
-Fig.~\ref{fig:heat_map} depicts the results of a Monte Carlo method of
-simulating possible paths of the H$^2$Bird during our experiments. For each
-point within a 0.2 meter grid within the camera viewing triangle, we simulated
-10 trials using an initial heading randomly sampled from a uniform
-distribution between -90$^{\circ}$ and 90$^{\circ}$. As shown in the figure,
+We conducted a Monte Carlo simulation to determine the probability of successful window navigation for a given point within the testing area.
+Using this information we can make assumptions about ideal starting positions for coopertive control.
+For each point within a 0.1 meter grid within the camera viewing triangle, we simulated
+40 trials using an initial heading randomly sampled from a uniform
+distribution between -90$^{\circ}$ and 90$^{\circ}$.
+Fig.~\ref{fig:heat_map} depicts the results of the Monte Carlo simulation. As shown in the figure,
 there is a region in the middle with a 100\% probability of success, which is

### 2012-10-13 manuscript.tex
Edit after merge

--- 2012-10-13 manuscript.tex
@@
 their on-board sensing and computational resources. Ground vehicles have 
-less mobility than MAVs, but relaxed size constraints and typically more 
+less mobility than MAVs, but relaxed size constraints, and typically more 
 computing power. These specializations present many opportunities for cooperation. In this work, we 
@@
 and a lightweight ground station using computer vision.
+We demonstrate the performance of the location estimation and navigation algorithms.
 We develop models for the ornithopter, ground station, and cooperative system dynamics.
@@
 real-time motion tracking algorithm on a live video stream. This allows 
-the ornithopter to reliably negotiate flight through narrow openings. We demonstrate the performance of the location estimation and navigation  algorithms.
+the ornithopter to reliably negotiate flight through narrow openings.
 \end{abstract}
@@
 Power, size, and weight constraints on micro air vehicles (MAV) 
-significantly limit sensing their and computational capabilities. They can execute only the most rudimentary
-control and state estimation algorithms in real-time. 
-The constraints on flapping-wing MAVs are even more restrictive. 
-Though they exhibit demonstrate safety and noise profiles superior to those
-of rotor craft MAVs, this comes at the expense of reduced payload weight.
+significantly limit their sensing and computational capabilities. They can 
+execute only the most rudimentary control and state estimation algorithms in 
+real-time.  The constraints on flapping-wing MAVs are even more restrictive. 
+Though they demonstrate safety and noise profiles superior to those of 
+rotorcraft MAVs, this comes at the expense of reduced payload.
 
-The non-linear
-flight dynamics exhibited of flapping wing flight compounding constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. 
-Much past research on flapping wing MAVs focused on overcoming the 
-difficulties associated with controlling and modeling flapping-wing vehicles in the face of this
+The non-linear dynamics flapping-wing flight compound 
+these constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Much 
+past research on flapping-wing MAVs focuses on overcoming the difficulties 
+of controlling and modeling flapping-wing vehicles in the face of this
 non-linearity\cite{floreano:flying}. Common approaches to simplify these 
-dynamic models include averaging over the wingbeat period~\cite{Schenato:flightcontrol}
-\cite{cheng:tandrdamping}, and using linear, low-dimensional models to predict flight behavior~\cite{Shigeoka:ornithopter}. These advances lead to the
-development of some high-performance platforms, such as the sub-15 gram DelFly 
-described by deCroon et al.~\cite{delfly:design}.
+dynamic models include averaging over the wing-beat 
+period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, and using 
+linear, low-dimensional models to predict flight 
+behavior~\cite{Shigeoka:ornithopter}. These advances lead to the development 
+of high-performance platforms such as the sub-15 gram DelFly, described by 
+deCroon et al.~\cite{delfly:design}.
 
-Many control laws that previous researchers developed from
-dynamic models of flapping-winged flight, such as periodic forcing and other
-control methods described %TODO you need to use real names, it's not kosher to refer to citation numbers directly
-by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
-require powerful on-board computing. In this work, we develop a simpler, less
-computationally intensive method of control for negotiating narrow passages, by exploiting cooperation between specialized robotic agents.
+Many control laws previous researchers developed from dynamic models of 
+flapping-wing flight, such as periodic forcing and other control methods 
+described by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
+require powerful on-board computing. In this work, we develop a simpler, 
+less computationally-intensive method of control for negotiating narrow 
+passages, by exploiting cooperation between specialized robotic agents.
 
-In contrast to flight control, there is considerably less previous work on guidance and of flapping-wing MAVs. 
-Baek demonstrated altitude control with an external camera~\cite{baek:altitude}. 
-de Croon et al. demonstrated obstacle avoidance with an onboard camera and offboard 
-processing~\cite{delfly:avoid}. For sub-20 gram ornithopters, Baek et al. 
-\cite{baek:tracking} provides one of the few examples of completely autonomous target seeking for vehicles at this scale.
+In contrast to flight control, there is considerably less previous work on 
+guidance and navigation of flapping-wing MAVs. Baek demonstrated altitude 
+control with an external camera~\cite{baek:altitude}. de Croon et al. 
+demonstrated obstacle avoidance with an onboard camera and offboard 
+processing~\cite{delfly:avoid}. Baek et al. provide one of the few examples 
+of completely autonomous target seeking for vehicles at the sub-20 gram 
+scale~\cite{baek:tracking}.
 
-There also exists a body of work on the guidance and control of fixed- and rotary-wing 
-MAV platforms, but many of these methods make extensive use of GPS or motion capture systems
-~\cite{kingston:timeattitude}\cite{kanade:3dvision}. Indoors, using
-GPS to estimate the pose of an MAV is impossible, and motion capture
-requires extensive and stationary modification of the environment. The power
-and payload limits on flapping wing MAVs make complex on-board sensing difficult, so multi-agent
-cooperative control schemes are apt solutions.
+There also exists a body of work on the guidance and control of fixed- and 
+rotary-wing MAV platforms, but many of these methods make extensive use of 
+GPS or motion capture systems~\cite{kingston:timeattitude}\cite{kanade:3dvision}. 
+Indoors, using GPS to estimate the pose of an MAV is impossible, and motion 
+capture requires extensive and stationary modification of the environment. 
+The power and payload limits on flapping-wing MAVs make complex on-board 
+sensing difficult, so multiagent cooperative control schemes are apt 
+solutions.
 
+%TODO Ryan's paragraph
 \cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
@@
 % humphrey
-To explore these cooperative control concepts, we designed a new flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{h2bird}). Built around the Silverlit i-Bird 
-RC flyer power train and clap-fling 
-wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
+To explore these cooperative control concepts, we designed a new 
+flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{fig:h2bird}). Built 
+around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
 \href{http://www.silverlit-flyingclub.com/wingsmaster/}
@@
 tail propeller and servo-controlled tail elevator provide control in the yaw 
-and pitch axes (Fig.~\ref{h2Bird_axes}. The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
+and pitch axes (Fig.~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
 \href{https://github.com/biomimetics/imageproc\_pcb}
      {https://github.com/biomimetics/imageproc\_pcb}} 
-controller consists of a 40 MIPS microprocessor, 6 DOF IMU, 
+controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
 IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
@@
 % ryan
-Cooperative behavior is a promising strategy for autonomous navigation of 
-unexplored and unstructured environments. However, to operate in these environments,
-a cooperative system must be fully mobile, not tied to a static ground station or 
-motion capture laboratory. To demonstrate the feasibility of this approach, 
-we designed our ground station to conform to the power and weight 
-constraints of highly mobile millirobotic ground vehicles, such as 
-OctoRoACH, which can traverse rough terrains with speed and agility~\cite{Pullin2012Dynamic}.
+Cooperative behavior is a promising strategy for autonomous navigation in 
+unexplored and unstructured environments. However, to operate in these 
+environments, a cooperative system must be fully mobile, not tied to a 
+static ground station or motion capture laboratory. To demonstrate the 
+feasibility of this approach, we designed our ground station to conform to 
+the power and weight constraints of highly mobile millirobotic ground 
+vehicles, such as OctoRoACH, which can traverse rough terrains with speed and 
+agility~\cite{Pullin2012Dynamic}.
 
-Rather than power-intensive PC processors, we used the ARM-based 
-BeagleBoard single-board computer as reference computation platform. The
-BeagleBoard consumes approximately one watt while running our computer vision 
-and control algorithms. It uses the same processor as popular sub-10 gram 
-processor modules---such as the Gumstix Overo and LogicPD Torpedo---which are 
+Rather than power-intensive PC processors, we used the ARM-based BeagleBoard 
+single-board computer as reference computation platform. The BeagleBoard 
+consumes approximately one watt while running our computer vision and 
+control algorithms. It uses the same processor as popular sub-10 gram 
+processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
 light enough for deployment on millirobotic ground vehicles. We pair the 
@@
 Canonical Ltd., \href{http://www.ubuntu.com/}{http://www.ubuntu.com/}} distribution
-modified with a custom version of the Linux kernel. We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
-{http://www.python.org/}} programming language for control, robot communication, 
-and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
+modified with a custom version of the Linux kernel. 
+We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
+{http://www.python.org/}} programming language for control, robot 
+communication, and telemetry, and the OpenCV\footnote{\raggedright OpenCV: Willow Garage, 
 \href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
@@
 \includegraphics[height=150pt]{figures/h2bird_axes.pdf}
-\caption{H$^2$Bird ornithopter with attitude control axes and labeled control surfaces.}
+\caption{H$^2$Bird ornithopter with attitude control axes and labeled 
+control surfaces.}
 \label{fig:h2Bird_axes}
@@
 
-We implement attitude estimation and control of the H$^2$Bird on-board to achieve 
-high sample-rate attitude control. We estimate the vehicle pose by
+We implement attitude estimation and control of the H$^2$Bird on-board to 
+achieve high sample-rate attitude control. We estimate the vehicle pose by
 integrating the IMU, and use a proportional-integral-derivative (PID) 
-controller to feed these estimates back to the attitude control flight surfaces (Figs.\ref{process_flow} and~\ref{block_diagram}).
+controller to feed these estimates back to the attitude control flight 
+surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).
 
-The high pitch angles flapping-wing MAVs regularly operate at distinguish them 
-from other MAVs. Baek demonstrated control on a similar vehicle using PID and an Euler angle
-parameterization of orientation~\cite{baek:tracking}. However, Euler angle parameterizations of vehicle orientation 
-suffer from singularities around high pitch angles. So we instead 
-represent the vehicle orientation with quaternions. An additional benefit of 
-the quaternion representation is its ability to represent relative body and 
-world coordinate angular displacements compactly with quaternion multiplication.
+The high pitch angles at which flapping-wing MAVs regularly operate 
+distinguish them from other MAVs. Baek demonstrated control on a similar 
+vehicle using PID and an Euler angle parameterization of orientation~\cite{baek:tracking}. 
+However, Euler angle parameterizations of vehicle orientation suffer from 
+singularities around high pitch angles. So we instead represent the vehicle 
+orientation with quaternions. The quaternion representation allows us to 
+represent relative body and world coordinate angular displacements compactly 
+with quaternion multiplication.
 
-Our implementation of quaternion-based control is similar to Knoebe's~\cite{knoebe:quatcontrol}:
-Given a reference pose represented in quaternion form $q_r$, we define the error as the rotation $q_e$ 
-required to reach the reference pose from the estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
+Our implementation of quaternion-based control is similar to 
+Knoebe's~\cite{knoebe:quatcontrol}:
+Given a reference pose represented in quaternion form $q_r$, we define the 
+error as the rotation $q_e$ required to reach the reference pose from the 
+estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
 \begin{align}
@@
 \end{align}
-We then convert this error quaternion to an angle axis representation and use it 
-as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
-We perform these calculations on-board, with trigonometric lookup tables for real-time performance.
+We then convert this error quaternion to an angle axis representation and 
+use it as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
+We perform these calculations on-board, with trigonometric lookup tables for 
+real-time performance.
 \begin{align}
@@
 The ground station performs simple two-dimensional pose estimation on 
-H$^2$Bird using a particle filter-based motion tracking algorithm.
-The produced pose estimate is the feedback input to the cooperative visual servoing feedback 
-loop (Sec. \ref{sec:visual_servoing_concept}). We limit our pose estimation 
-approach to monocular two-dimensional object tracking in pixel space. This 
-allows us to perform real-time video tracking using the modest computational 
-resources available on the ground station, and similarly, millirobotic ground 
-vehicles.
+H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
+estimate is the feedback input to the cooperative visual servoing feedback 
+loop, which is represented as the outermost feedback paths on the block 
+diagrams in Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}. We 
+limit our pose estimation approach to monocular two-dimensional object 
+tracking in pixel space. This allows us to perform real-time video tracking 
+using the modest computational resources available on the ground station, 
+and similarly, millirobotic ground vehicles.
 
-Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a particle filter ~\cite{thrun2005probabilistic}. 
-We initialize the particles uniformly across the frame before tracking begins. 
-To improve numerical stability, we normalize the weights of all particles on 
-each iteration, and only resample the particle population when the mean particle 
-weight falls below a pre-determined threshold.
-
+Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a 
+particle filter~\cite{thrun2005probabilistic}. We initialize the particles 
+uniformly across the frame before tracking begins. To improve numerical 
+stability, we normalize the weights of all particles on each iteration, and 
+only resample the particle population when the mean particle weight falls 
+below a pre-determined threshold. 
 %............................................................................%
@@
 $\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
-$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at time $t$, $\Sigma$ 
-is a chosen covariance matrix, and $b$ is vector of the bounds of the pixel space.
+$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
+time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
+bounds of the pixel space.
 \begin{equation}
@@
 In the experiments we present in this work, we used a diagonal covariance matrix 
-with a uniform variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
+with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
 transition model is sufficient for tracking motion in which frame-to-frame 
@@
 transitions increase tracking robustness, by forcing the filter to always 
-sample broadly across of the pixel space. This helps prevent overconfidence and collapse, 
-and decreases reacquisition delay.
+sample broadly across of the pixel space. This helps prevent overconfidence 
+and collapse, and decreases reacquisition delay.
 
@@
 Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
-approach to motion tracking via na\"{i}ve background subtraction, e.g. frame differencing[CITATION].
-On each iteration, the system provides the emission model with $f_0$, the frame representing 
-the background and $f_t$, the most recent frame captured. The model assigns 
-a weight $w^{[i]}_t$ to each particle $x^{[i]}_t$, proportional to the color 
-distance in RGB space between the pixels of $f_0$ and $f_t$ at the location 
-$x^{[i]}_t$.
+approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
+differencing[CITATION]. On each iteration, the system provides the emission 
+model with $f_0$, the frame representing the background, and $f_t$, the most 
+recent frame captured. The model assigns a weight $w^{[i]}_t$ to each 
+particle $x^{[i]}_t$, proportional to the color distance in RGB space 
+between the pixels of $f_0$ and $f_t$ at the location $x^{[i]}_t$.
 \begin{equation}
@@
 
-During experiments presented in this work, we assumed that neither the bird, 
-nor any other moving object, were visible in the frame on system 
+During the experiments presented in this work, we assumed that neither the 
+bird, nor any other moving object, were visible in the frame on system 
 start. We saved the first frame captured by the system, and provided it as 
-the background frame $f_0$ for every iteration of the particle filter. This choice of $f_0$ 
-resulted in a sampling approach to na\"{i}ve background subtraction.
+the background frame $f_0$ for every iteration of the particle filter. This 
+choice of $f_0$ resulted in a sampling approach to na\"{i}ve background 
+subtraction.
 
-Na\"{i}ve background subtraction is known to be brittle to changes in the scene over 
-time[CITATION]. However, we find it very efficient and sufficiently reliable for the 
-short time spans necessary for H$^2$Bird to successfully negotiate narrow 
-openings, which are typically less than five seconds. Substituting $f_0$ 
-with a more sophisticated background extraction would retrain the underlying motion tracking features of the filter, while providing it with a more robust way of differentiating subject motion from the background.
+Na\"{i}ve background subtraction is brittle to changes in a scene over 
+time [CITATION]. However, we find it very efficient, and sufficiently 
+reliable for the short time spans--less than five seconds--necessary for 
+H$^2$Bird to successfully negotiate narrow openings. Calculating $f_0$ 
+using a more sophisticated background extraction would retrain the 
+underlying motion tracking features of the filter, while providing the 
+algorithm with a more robust way of differentiating subject motion from the 
+background.
 
@@
 The ground station guides the H$^2$Bird by commanding changes to its 
-reference attitude, as governed by its on-board attitude controllers (Figs.\ref{process_flow} and~\ref{block_diagram}). 
-With the estimate of 2D vehicle pose from computer vision, we calculate an altitude
-error and horizontal error. The ground station then commands a reference attitude to 
-the H$^2$Bird pitch and yaw. according to a PID control law. We apply a fixed-lag window the PID controller integrator term 
-to prevent wind-up while we launch the ornithopter. We tuned the proportional and derivative terms of this outer controller 
-to dampen the overshoot and drift cause by the cooperative system latency.
-
+reference attitude, which is in turn governed by its on-board attitude 
+controllers (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
+the estimate of 2D vehicle pose from the pose estimation algorithm, the 
+ground station calculates an altitude error and horizontal error. The ground 
+station then commands a reference attitude to the H$^2$Bird over radio, 
+according to a PID control law. We apply a fixed-lag window the PID 
+controller integral term to prevent wind-up while we launch the ornithopter. 
+We tuned the proportional and derivative terms of this outer controller to 
+dampen the overshoot and drift caused by the cooperative system latency.
 \begin{figure}[tb]
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{System Model}
+\label{sec:system_model}
+
+%----------------------------------------------------------------------------%
+\subsection{Ornithopter Attitude Control}
+\label{sec:model_attitude}
+
+We divide our model of the vehicle pose into two independent models for 
+simplicity: one for yaw motion, and one for translational motion.
+
+We approximate the yaw dynamics of the ornithopter with a one-dimensional 
+model to limit complexity, and because it matches our physical 
+implementation. Likewise, we choose the reference heading angle as input to 
+the model, and model the dynamics of vehicle and on-board PID controller as 
+one system, whose output is the physical heading angle. We show the 
+resulting model form in Equation~\ref{eq:plain_model}.
+\begin{equation}
+\label{eq:plain_model}
+H(s) = \frac{K}{1 + Rs}
+\end{equation}
+To model the translational position of the ornithopter, we use a form of 
+Dubin's car model (Eq.~\ref{eq:dubin_car})~\cite{lavalle:planning}, where 
+$\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed. 
+\begin{equation}
+\label{eq:dubin_car}
+\begin{aligned}
+& \dot{x} = u_{s}sin(\theta)\\
+& \dot{y} = u_{s}cos(\theta)\\
+& \dot{\theta} = u\\
+& u \in U=[-\tan{\phi_{max}},\tan{\phi_{max}}]
+\end{aligned}
+\end{equation}
+We modify this model to treat the angular position $\theta$ as an input 
+rather than a state. The overall motion of the system can be modeled as shown 
+in Equation~\ref{eq:dubin_car}, where $\theta$ is the time-domain output of 
+Equation~\ref{eq:plain_model} with the reference heading angle as input.
+
+%----------------------------------------------------------------------------%
+\subsection{Ground Station Pose Estimation}
+
+Our model of the ground station comprises models of both the camera and the
+pose estimation computation. The camera model contains both a latency, which 
+is in series with the path between the ``Camera'' and 
+``Ornithopter Dynamics'' blocks in Figure~\ref{fig:block_diagram}, and a 
+model of the camera's viewing angle constraints.
+
+We model the viewing angle of the camera as a pixel map from one edge of a 
+63$^{\circ}$ viewing triangle to the other. The camera has an image width of 
+640 pixels. We calculate the desired heading angle in degrees, for a given 
+distance from the camera, according to Equation~\ref{eq:desired_heading}.
+\begin{equation} 
+\label{eq:desired_heading} 
+\text{Desired heading} = \frac{640\,\text{px}}{2d\tan{\frac{63^{\circ}}{2}}}{\times\frac{63^{\circ}}{640\,\text{px}}\times x}
+\end{equation}
+where $d$ is the distance between the ornithopter and the camera, and $x$ is 
+the horizontal position of the robot relative to the center of the window in 
+meters. This relation reflects the fact that the pose estimation algorithm 
+does not determine the depth of the robot, which can be problematic for 
+control algorithms. For example, if the ornithopter remains at a constant 
+horizontal distance from the window, but moves closer and closer to the 
+camera, the ground station's directed heading steadily increases, even 
+though the robot has not actually moved farther from the window. These 
+problems are most severe in areas close to the camera.
+
+We model the pose estimation computation delay, which include the latency of 
+updating the particle filter, calculating a new reference attitude, and 
+sending the reference attitude command the ornithopter via radio.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}
+
+%----------------------------------------------------------------------------%
+\subsection{System Identification}
+
+%............................................................................%
+\subsubsection{Ornithoper Attitude Control}
+We determined the turning speed and turning radius of the H$^2$Bird 
+experimentally, by measuring the response of the system to a step input of a 
+clockwise 90$^{\circ}$ turn. We measured the response of the system by 
+recording the attitude estimates calculated by the ornithopter using its 
+on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
+\href{http://www.mathworks.com/products/matlab/}
+     {http://www.mathworks.com/products/matlab/}} 
+to fit a simple low-order model to the step response, as discussed in 
+Sec.~\ref{sec:model_attitude}.
+
+%............................................................................%
+\subsubsection{Ground Station Pose Estimation}
+To measure the capture latency of the camera, we wrote a simple test 
+application which draws a large colored rectangle on the ground station's 
+screen, then uses the camera to detect that changes in that rectangle's 
+color. By noting the value of the system timer just before it changes the 
+rectangle, and just after the camera detects that change, the application 
+arrives at an estimate of the video capture pipeline latency. We averaged 
+these result over several hundred rapidly-executed trials to determine the 
+capture latency we used to model the system.
+
+To compute the particle filter and controller latency, we again used system 
+timers to determine the average time between the start and end of computation.
+
+%----------------------------------------------------------------------------%
+\subsection{Model Verification}
+\label{sec:experiments_verification}
+
+We tested our cooperative visual servoing system with a simple target seeking
+experiment: the H$^2$Bird must fly through a specified obstacle, in this 
+case, a wooden frame we used to simulate a window.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
+\caption{Conceptual sketch of experimental setup.}
+\label{fig:experiment_cartoon}
+\end{figure}
+
+As the H$^2$Bird cannot detect the target on its own, we used cooperative 
+control with the ground station, which to provided remote guidance to 
+H$^2$Bird to allow it to navigate to the window. We used a motion capture 
+system to record ground truth data. We present a sketch of this experimental 
+setup in Figure~\ref{fig:experiment_cartoon}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{System Model}
-\label{sec:system_model}
-
-%----------------------------------------------------------------------------%
-\subsection{Ornithopter Attitude Control}
-\label{sec:model_attitude}
-
-We divide our model of the vehicle pose into two independent models for simplicity: one for
-yaw motion, and one for translational motion.
-
-We approximate the yaw dynamics of the ornithopter with a one-dimensional model to limit complexity and because it matches our physical implementation.
-Likewise, we choose the reference heading angle as input to the model and model the dynamics of vehicle and on-board PID controller as one system, which outputs the physical heading angle.
-We show the resulting model form in Equation~\ref{eq:plain_model}.
-
-\begin{equation}
-\label{eq:plain_model}
-H(s) = \frac{K}{1 + Rs}
-\end{equation}
-
-\begin{equation}
-\label{eq:dubin_car}
-\begin{aligned}
-& \dot{x} = u_{s}sin(\theta)\\
-& \dot{y} = u_{s}cos(\theta)\\
-& \dot{\theta} = u\\
-& u \in U=[-\tan{\phi_{max}},\tan{\phi_{max}}]
-\end{aligned}
-\end{equation}
-
-To model the translational position of the ornithopter, we use a form of Dubin's car model (Eq.
-~\ref{eq:dubin_car})~\cite{lavalle:planning}, where $\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed.
-We modify this model to treat the angular position $\theta$ as an input rather than a state.
-The overall motion of the system can be modeled as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (\ref{eq:plain_model}) with the reference heading angle as the input.
-
-Our form uses the angular position, $\theta$ as an input rather than a state.
-We model the overall motion of the system as shown in Equation~\ref{eq:total_model_plain}, where $\theta$ is the time domain output of (Eq.
-\ref{eq:plain_model}) with the reference heading angle as the input.
-
-%----------------------------------------------------------------------------%
-\subsection{Ground Station Pose Estimation}
-
-Our model of the ground station comprises models of both the camera and the
-pose estimation computation. The camera model contains both a
-latency, which is in series with the path between the ``camera'' and ``ornithopter dynamics''
- blocks in Fig.~\ref{fig:block_diagram}, and a model of the camera's viewing angle constraints.
-
- We model the 
-viewing angle of the camera as a pixel map from one edge of a
-63$^{\circ}$ viewing triangle to the other. The camera has an image width of
-640 pixels. We calculate the desired heading angle in degrees for a given distance from the
-camera according to Equation~\ref{eq:desired_heading}, where $d$
-is the distance between the ornithopter and the camera, and $x$ is the horizontal
-position of the robot relative to the center of the window in meters. This
-relation reflects the fact that the camera cannot determine the depth of the
-robot. One example of why this is a problem is that if the robot stays at a
-constant horizontal distance from the window, but moves closer and closer to
-the camera, the directed heading steadily increases even though the robot has
-not actually moved farther from the window. This creates tracking problems in
-areas close to the camera.
-
-\begin{equation} \label{eq:desired_heading} \text{Desired heading} =
-\frac{640\text{px}}{2*\text{d}*\tan{\frac{63^{\circ}}{2}}}*\frac{63^{\circ}}{640\text{px}}*\text{x}
-\end{equation}
-
-We model the pose estimation computation by the by the latency of running 
-the particle filter, calculating a new reference attitude, and sending to the ornithopter via radio.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
-
-%----------------------------------------------------------------------------%
-\subsection{System Identification}
-
-%............................................................................%
-\subsubsection{Ornithoper Attitude Control}
-
-We determined the turning speed and turning radius of the H$^2$Bird
-experimentally, by measuring the response of the system to a step input of a
-clockwise 90$^{\circ}$ turn. We measured the response of the system by recording
-the attitude estimates calculated by the ornithopter using its on-board
-gyroscope. We then used
-MATLAB\footnote{The MathWorks, Inc. Matlab:
-\href{http://www.mathworks.com/products/matlab/}
-     {http://www.mathworks.com/products/matlab/}} to fit a
-simple low-order model to the step response, as discussed in Sec.~\ref{sec:model_attitude}.
-
-%............................................................................%
-\subsubsection{Ground Station Pose Estimation}
-
-To measure the capture latency of the camera, we wrote a simple test 
-application which changes a large colored rectangle on the ground station's screen, then uses the camera to detect that change.
-By noting the value of the system timer just before we change the screen, and just after the camera detects that change, the application arrives at an estimate of the video capture pipeline latency.
-We averaged these result over several hundred rapidly-executed trials to determine the capture latency we used to model the system..
-
-To compute the particle filter and controller latency, we again used system timers to determine the
-average time between the start and end of computation.
-
-%----------------------------------------------------------------------------%
-\subsection{Model Verification}
-
-We tested our cooperative visual servoing system with a simple target seeking
-experiment: the H$^2$Bird must fly through a specified obstacle, in this case,
-a wooden frame we used to simulate a window. As the H$^2$Bird cannot detect the
-target on its own, we used the cooperative control with the ground station, which to provided remote guidance to H$^2$Bird to allow it to navigate to the window.
-We used a motion capture system to record 
-ground truth data. We present a sketch of this experimental setup in
-Fig.~\ref{fig:experiment_cartoon}.
-
-% TODO clarify or roll into paragraph
-Each trial included the following steps:
+Each trial included the following steps (see Fig.~\ref{fig:pf_screencap}):
 \begin{enumerate}
-\item The window is identified by the user. 
-\item We release H$^2$Bird by hand, in view of the camera at the desired starting grid point.
-\item The H$^2$Bird attempts to fly through the window with, with remote guidance from
-the ground station.
+\item We identify the window to the ground station by dragging a bounding 
+box over it using the mouse. 
+\item We release H$^2$Bird by hand, in view of the camera at the desired 
+starting grid point.
+\item The H$^2$Bird attempts to fly through the window with, with remote 
+guidance from the ground station.
 \end{enumerate}
 
-In total, we ran 80 trials over a variety of starting positions. Our goal for the experiments was to verify the model described in
-Secttion~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a grid
-along the edges of the camera view plane, and in 0.05 meter increments from the
-window, in an effort to determine the success rate on the edges of the testing space. We
-conducted 20 additional trials directly in front of the camera to determine
-the success rate near the center of viewing plane. After every 5 trials,
-we replaced the ornithopter's battery with a fully charged one.
+After every 5 trials, we replaced the ornithopter's battery with a fully 
+charged one.
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
-\caption{Conceptual sketch of experimental setup.}
-\label{fig:experiment_cartoon}
-\end{figure}
-%TODO label experiment cartoon
+In total, we collected 80 trials over a variety of starting positions. Our 
+goal for these experiments was to verify the model described in 
+Section~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a 
+grid along the edges of the camera view plane, and in 5 cm meter increments 
+from the window, in an effort to determine the success rate on the edges of 
+the testing space. We conducted 20 additional trials directly in front of 
+the camera, to determine the success rate near the center of viewing plane.
 
@@
 the feasible set of initial conditions that result in successful narrow 
-passage traversal using both analytical and empirical methods. We conducted 
-experiments to validate this model.
-
-The following sections detail how we calculated the feasible and infeasible
-regions using system identification and present the results of our
-experimental trials.
+passage traversal, using both analytical and empirical methods. We then 
+conduct experiments to validate this model.
 
@@
 
-The first-order model that we fit to the
-step response of the onrithopter yaw motion is in Equation~\ref{eq:transfer_func}.
-This model represents the tuning dynamics of the ornithopter as well as the PID controller providing inputs to the actuators.
-The results of the step response experiment along with the response of the fitted model are depicted in
-Fig.~\ref{fig:step_response}.
-
+The first-order model that we fit to the step response of the onrithopter 
+yaw motion is in Equation~\ref{eq:transfer_func}. This model represents the 
+tuning dynamics of the ornithopter as well as the PID controller providing 
+inputs to the actuators. We present the results of the the step response 
+experiment, along with the response of the fitted model, in 
+Figure~\ref{fig:step_response}.
 \begin{equation}
@@
 \end{equation}
-
-The step response rise time is 1.4 seconds. The H$^2$Bird flies
-at an average forward velocity of 1.2 m/s during our experiments, so we
-estimate the minimum turning radius of the robot is 1.07 maters, as described in
+The step response rise time is 1.4 seconds. The H$^2$Bird flies at an 
+average forward velocity of 1.2 m/s during our experiments, so we estimate 
+the minimum turning radius of the robot is 1.07 meters, as described in 
 Equation~\ref{eq:min_radius}.
@@
 \begin{aligned}
-\text{Minimum radius} & = \frac{360^{\circ}*\text{Rise time}*\text{Flight
-speed}}{\text{Final angle}*2*\pi}\\
-& = \frac{360^{\circ}*1.4s*1.2\frac{m}{s}}{90^{\circ}*2*\pi} = 1.07 m
+\text{Minimum radius} & = \frac{360^{\circ}\times\text{Rise time}\times\text{Flight
+speed}}{\text{Final angle}\times 2\pi}\\
+& = \frac{360^{\circ}\times 1.4s\times 1.2\frac{m}{s}}{90^{\circ}\times2*\pi} = 1.07 m
 \end{aligned}
 \end{equation}
-In the experiments presented in this work, we choose a flight speed
-lower than the maximum speed of the ornithopter, to facilitate more robust
+In the experiments we present in this work, we choose a flight speed lower 
+than the maximum speed of the ornithopter, to facilitate more robust 
 tracking.
@@
 
-The latency that we measured for the camera was 25ms, and the measured
-partiicle filter computational latency was 12.5ms.
+We measured a camera capture latency 40 ms, pose estimation computation 
+latency of 12.5 ms, and radio latency of 25 ms.
 
@@
 
-Conducting the experimental trials as described in Section~\ref{sec:experiments:verification}, we achieved a success rate of 80\% for
-initial conditions in the middle of the testing space and in front of the
-camera. As we moved towards the edges of the
-feasible region in Fig.~\ref{fig:flight_paths}, the success rate diminished
-to 50\%. The results of our experimental trials are shown in
-Figs.~\ref{fig:flight_paths} and~\ref{fig:flight_paths_feasible}.
+Conducting the experimental trials as described in 
+Section~\ref{sec:experiments_verification}, we achieved a success rate of 
+80\% for initial conditions in the middle of the testing space and in front 
+of the camera. As we moved towards the edges of the feasible region in 
+Figure~\ref{fig:flight_paths}, the success rate diminished to 50\%. We 
+present the results of our experimental trials in Figures~\ref{fig:flight_paths} 
+and~\ref{fig:flight_paths_feasible}.
 
-The lower success rate on the edges of the camera frame can be explained by several factors. On the edges of the viewing triangle, the yaw
-control inputs are large in magnitude, although this is not necessary in all
-cases. Since the control algorithm we are using has no notion of depth, it
-applies the same control input whether the ornithopter is close to the camera,
-where small changes in heading are adequate, or far away from the camera,
-where large changes in heading are necessary. Due to this phenomenon, the
-controller often over or under-compensates in certain regions, depending upon
-the PID tuning and the distance from the camera. Tracking noise can also cause
-passage negotiation to fail. Variations in the H$^2$Bird hand launch can cause
-the system to respond differently for similar initial conditions. In addition,
-the H$^2$Bird battery was changed approximately every 5 tests, adding more
-variation to the H$^2$Bird performance by affecting the motor power output.
+The lower success rate on the edges of the camera frame is explained by 
+several factors. On the edges of the viewing triangle, the yaw control 
+inputs are large in magnitude, although this is not necessary in all cases. 
+Since the control algorithm we are using has no notion of depth, it applies 
+the same control input whether the ornithopter is close to the camera, where 
+small changes in heading are adequate, or far away from the camera, where 
+large changes in heading are necessary. Due to this phenomenon, the 
+controller often over or under-compensates in certain regions, depending 
+upon the PID tuning and the distance from the camera. Tracking noise can 
+also cause passage negotiation to fail. Variations in the H$^2$Bird hand 
+launch can cause the system to respond differently for similar initial 
+conditions. In addition, we changed H$^2$Bird battery approximately every 5 
+tests, adding more variation to the H$^2$Bird performance by affecting the 
+motor power output.
 
@@
 a result of a Monte Carlo simulation of the reachable set model, in which the 
-initial heading of the H$^2$Bird is chosen uniformly at random from $(-90,90)$.}
+initial heading of the H$^2$Bird is chosen uniformly at random from $(-90^{\circ},90^{\circ})$.}
 \label{fig:heat_map}
@@
 
-We simulated the motion of the ornithopter beginning from various positional
-and angular heading initial conditions.
-We used the model that we determined through system identification (Sec.~\ref{sec:system_model}) using the cooperative control method in Fig.~\ref{fig:block_diagram}.
+We simulate the motion of the ornithopter beginning from various positional 
+and angular heading initial conditions. We use this model, as determined 
+through system identification (Sec.~\ref{sec:system_model}) using the 
+cooperative control method in Figure~\ref{fig:block_diagram}.
 
-We determined the backwards reachable set of initial states for successful
+We determine the backwards reachable set of initial states for successful 
 narrow passage traversal both geometrically and in simulation 
-(Fig.~\ref{fig:feasible_set}). In the figure, the blue lines represent the camera viewing triangle, the black line
-represents the window, and the black dot represents the camera. For the initial conditions, we used a uniform
+(Fig.~\ref{fig:feasible_set}). In the figure, the blue lines represent the 
+camera viewing triangle, the black line represents the window, and the black 
+dot represents the camera. For the initial conditions, we used a uniform
 grid of 5 cm increments within the camera viewing region, and an initial
-angular heading perpendicular to the window plane. A success, or feasible initial condition, consists
-of a path that intersects the window at some point and does not leave the
-camera viewing triangle, and is indicated in green. The geometrically
-infeasible region is computed using the minimum turning radius and the
-geometry of the camera's field-of-view and is indicated in blue. The region in red is not geometrically feasible, but is infeasible as a result of the cooperative control simulation dynamics.
+angular heading perpendicular to the window plane. A success, or feasible 
+initial condition, consists of a path that intersects the window at some 
+point and does not leave the camera viewing triangle, and is indicated in 
+green. The geometrically infeasible region is computed using the minimum 
+turning radius and the geometry of the camera's field-of-view and is 
+indicated in blue. The region in red is not geometrically feasible, but is 
+infeasible as a result of the cooperative control simulation dynamics.
 
-Our experimental data validates Fig.~\ref{fig:feasible_set}, as there are no successes within the geometrically
-infeasible region. There are a few successes near the boundary of the
-infeasible region determined in simulation.
-While there are successes along the boundary of the camera viewing triangle, there are also many failures.
-We further explore this behavior using Monte Carlo simulation.
+Our experimental data validates Figure~\ref{fig:feasible_set}, as there are 
+no successes within the geometrically infeasible region. There are a few 
+successes near the boundary of the infeasible region determined in 
+simulation. While there are successes along the boundary of the camera 
+viewing triangle, there are also many failures. We further explore this 
+behavior using Monte Carlo simulation.
 
-We conducted a Monte Carlo simulation to determine the probability of successful window navigation for a given point within the testing area.
-Using this information we can make assumptions about ideal starting positions for coopertive control.
-For each point within a 0.1 meter grid within the camera viewing triangle, we simulated
-40 trials using an initial heading randomly sampled from a uniform
-distribution between -90$^{\circ}$ and 90$^{\circ}$.
-Fig.~\ref{fig:heat_map} depicts the results of the Monte Carlo simulation. As shown in the figure,
-there is a region in the middle with a 100\% probability of success, which is
-a subset of the feasible region presented in Fig.~\ref{fig:feasible_set}.
-Along the edges of the camera viewing triangle, however, the probability of
-success decreases.
+We conducted a Monte Carlo simulation to determine the probability of 
+successful window navigation for a given point within the testing area. 
+Using this information we can make assumptions about ideal starting 
+positions for coopertive control. For each point within a 10 cm grid within 
+the camera viewing triangle, we simulated 40 trials using an initial heading 
+randomly sampled from a uniform distribution between -90$^{\circ}$ and 
+90$^{\circ}$. Figure~\ref{fig:heat_map} depicts the results of the Monte 
+Carlo simulation.
+
+As shown in the figure, there is a region in the middle with a 100\% 
+probability of success, which is a subset of the feasible region presented 
+in Figure~\ref{fig:feasible_set}. Along the edges of the camera viewing 
+triangle, however, the probability of success decreases.
 
@@
 simulation of our model. The trials that we conducted in
-Fig.~\ref{fig:flight_paths_feasible} correspond to the circled region in
-Fig.~\ref{fig:heat_map}. The 80\% success rate that we measured in the
+Figure~\ref{fig:flight_paths_feasible} correspond to the circled region in
+Figure~\ref{fig:heat_map}. The 80\% success rate that we measured in the
 experiments corroborates our calculated probability of success of
@@
 \section{Conclusions and Future Work}
-%TODO 
+We have demonstrated cooperative guidance of a 13 gram flapping wing MAV with a
+lightweight base station in hardware. In addition, we developed a model of the
+cooperative system by modeling the individual system capabilities, and composing
+it with known system interactions. The corroboration between our simulated and
+experimental results shows that our simplified modeling approach is useful for 
+predicting the performance of cooperative systems.
+
+We intend to further explore the benefits and limitations of collaboration in
+unstructured environments by deploying our base station on millirobotic ground
+vehicles. As the ground station used in this work was designed with this goal
+in mind, this is a matter of implementation. The effectiveness of our modeling
+approach, when applied to more complex systems with more interactions and 
+agents is also an future area of interest.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-10-13 manuscript.tex
Trim abstract

@@
 % METADATA
-\def \papertitle{Cooperative Control for Narrow Passage Traversal with an 
-Ornithopter Micro Air Vehicle and Lightweight Ground Station}
+\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with an 
+Ornithopter MAV and Lightweight Ground Station}
 \def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
@@
 \def \paperterms{Algorithms, Performance, Design, Experimentation}
-\def \paperauthors{Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing}
+%\def \paperauthors{Ryan C. Julian, Cameron J. Rose, Humphrey Hu, and Ronald S. Fearing}
+\def \paperauthors{Paper 580}
 
@@
 %% Paper number instead of authors
-%\numberofauthors{1}
-%\alignauthor Paper  580
+\numberofauthors{1}
+\author{
+\alignauthor Paper  580
+}
 
-\numberofauthors{4}
-\author{
-\alignauthor Ryan C. Julian\\
-       \affaddr{Department of EECS}\\
-       \affaddr{Univ. of California, Berkeley}\\
-       \affaddr{Berkeley, CA 94720}\\
-       \email{ryanjulian@berkeley.edu}
-\alignauthor Cameron J. Rose\\
-       \affaddr{Department of EECS}\\
-       \affaddr{Univ. of California, Berkeley}\\
-       \affaddr{Berkeley, CA 94720}\\
-       \email{c\_rose@eecs.berkeley.edu}
-\alignauthor Humphrey Hu\\
-       \affaddr{Robotics Institute}\\
-       \affaddr{Carnegie Mellon University}\\
-       \affaddr{Pittsburgh, PA 15213}\\
-       \email{humhu@cmu.edu}
-\and
-\alignauthor Ronald S. Fearing\\
-       \affaddr{Department of EECS}\\
-       \affaddr{Univ. of California, Berkeley}\\
-       \affaddr{Berkeley, CA 94720}\\
-       \email{ronf@eecs.berkeley.edu}
-}
+%\numberofauthors{4}
+%\author{
+%\alignauthor Ryan C. Julian\\
+%       \affaddr{Department of EECS}\\
+%       \affaddr{Univ. of California, Berkeley}\\
+%       \affaddr{Berkeley, CA 94720}\\
+%       \email{ryanjulian@berkeley.edu}
+%\alignauthor Cameron J. Rose\\
+%       \affaddr{Department of EECS}\\
+%       \affaddr{Univ. of California, Berkeley}\\
+%       \affaddr{Berkeley, CA 94720}\\
+%       \email{c\_rose@eecs.berkeley.edu}
+%\alignauthor Humphrey Hu\\
+%       \affaddr{Robotics Institute}\\
+%       \affaddr{Carnegie Mellon University}\\
+%       \affaddr{Pittsburgh, PA 15213}\\
+%       \email{humhu@cmu.edu}
+%\and
+%\alignauthor Ronald S. Fearing\\
+%       \affaddr{Department of EECS}\\
+%       \affaddr{Univ. of California, Berkeley}\\
+%       \affaddr{Berkeley, CA 94720}\\
+%       \email{ronf@eecs.berkeley.edu}
+%}
 
@@
 less mobility than MAVs, but relaxed size constraints, and typically more 
-computing power. These specializations present many opportunities for cooperation. In this work, we 
-demonstrate cooperative target-seeking between a 13 gram ornithopter MAV 
-and a lightweight ground station using computer vision.
-We demonstrate the performance of the location estimation and navigation algorithms.
-We develop models for the ornithopter, ground station, and cooperative system dynamics.
-We determine model parameters of the real systems through experimental system identification.
-Finally, we verify those models using experiments on narrow passage traversal, and arrive at a cooperative system model which accurately predicts the backwards-reachable region for successfully negotiating ornithopther flight through narrow openings.
+computing power. These specializations present many opportunities for cooperation. 
+In this work, we demonstrate cooperative target-seeking between a 13 gram 
+ornithopter MAV and a lightweight ground station using computer vision. We 
+develop models for the ornithopter, ground station, and cooperative system 
+dynamics. We determine model parameters of the real systems through experimental 
+system identification. Finally, we verify those models using experiments on 
+narrow passage traversal, and arrive at a cooperative system model which 
+accurately predicts the backwards-reachable region for successfully 
+negotiating ornithopther flight through narrow openings.
 
-We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
-features clap and fling wings, improves upon a previous design by 
+We introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
+features clap and fling wings, improves upon previous designs by 
 utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
-payload of 2.8 grams. We augment the ornithopter's built-in 
-gyroscope-based control with a lightweight ground station design, which has power and weight 
+2.8 gram payload. We augment the ornithopter's built-in gyroscope-based 
+control with a lightweight ground station design, which has power and weight 
 requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
-station provides heading estimates to the ornithoper by running a 
-real-time motion tracking algorithm on a live video stream. This allows 
-the ornithopter to reliably negotiate flight through narrow openings.
+station provides heading estimates to the ornithoper by running a real-time 
+motion tracking algorithm on a live video stream.
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-The authors would like to thank Fernando Garcia Bermudez for his 
-assistance with the Vicon motion capture system, Andrew Pullin for his 
-help with robot photography, and the members of the Biomimetic 
-Millisystems Laboratory and the EECS community at the University of 
-California, Berkeley for their advice and support.
+%\section{Acknowledgments}
+%The authors would like to thank Fernando Garcia Bermudez for his 
+%assistance with the Vicon motion capture system, Andrew Pullin for his 
+%help with robot photography, and the members of the Biomimetic 
+%Millisystems Laboratory and the EECS community at the University of 
+%California, Berkeley for their advice and support.
 

### 2012-10-13 manuscript.tex
some additions to conclusion

@@
 \section{Conclusions and Future Work}
-We have demonstrated cooperative guidance of a 13 gram flapping wing MAV with a
+We have demonstrated cooperative guidance of a 13 gram flapping-wing MAV with a
 lightweight base station in hardware. In addition, we developed a model of the
-cooperative system by modeling the individual system capabilities, and composing
+cooperative system using system identification methods on the individual system capabilities, and composing
 it with known system interactions. The corroboration between our simulated and
 experimental results shows that our simplified modeling approach is useful for 
-predicting the performance of cooperative systems.
+predicting the performance of cooperative systems at low computational cost.
+Our model could also be used to determine when to begin control with high probability of success, even though we lack information about the complete pose of the MAV.
 
@@
 approach, when applied to more complex systems with more interactions and 
-agents is also an future area of interest.
+agents is also a future area of interest.
 

### 2012-10-13 manuscript.tex
fixed some typos

@@
 
-The non-linear dynamics flapping-wing flight compound 
+The non-linear dynamics of flapping-wing flight compound 
 these constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Much 
@@
 Rather than power-intensive PC processors, we used the ARM-based BeagleBoard 
-single-board computer as reference computation platform. The BeagleBoard 
+single-board computer as a reference computation platform. The BeagleBoard 
 consumes approximately one watt while running our computer vision and 

### 2012-10-13 manuscript.tex
Fixup intro

--- 2012-10-13 manuscript.tex
@@
 
-%TODO Ryan's paragraph
-\cite{Jung1998Range}\cite{Hyams1999Cooperative}\cite{Mehta2006Adaptive}
-\cite{Rudol2008Micro}\cite{Luo2011Air}
+The literature on cooperative control of MAVs is similarly limited to larger
+vehicles and purely analytical explorations. Nebot, Hyams, and Jung all 
+demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Nebot2003Agents}\cite{Hyams1999Cooperative}\cite{Jung1998Range}. 
+Luo and Mehta analyze  UAV-UGV cooperation from a modeling perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. 
+Rudol demonstrated a cooperative visual servoing system similar with 
+UGVs and MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of 
+magnitude larger than the millirobotic systems we target, and so were able 
+to take advantage of high-performance processors and specialized cameras.
 
-%TODO Add in final paragraph to lead-in to our work
+The small size of millirobotic air and ground vehicles necessitates 
+specialization, but it also pays dividends agility. A group of specialized 
+heterogeneous millirobots can navigate more challenging terrains and 
+interact with their environments in ways that larger, monolithic vehicles 
+cannot. Despite this advantage, in order to collectively equal the sensing and 
+computational capabilities of larger robots, they will also have to interact 
+so that multiple robots can benefit from the specialized capabilities of 
+cooperating vehicles. Millirobotic systems will benefit from cooperation 
+even more than their larger counterparts.
 
@@
 approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
-differencing[CITATION]. On each iteration, the system provides the emission 
+differencing~\cite{Ahad2011Computer}. On each iteration, the system provides the emission 
 model with $f_0$, the frame representing the background, and $f_t$, the most 
@@
 Na\"{i}ve background subtraction is brittle to changes in a scene over 
-time [CITATION]. However, we find it very efficient, and sufficiently 
+time~\cite{Ahad2011Computer}. However, we find it very efficient, and sufficiently 
 reliable for the short time spans--less than five seconds--necessary for 

### 2013-01-16 manuscript.tex
Add authors and acknowledgements

@@
 %% Paper number instead of authors
-\numberofauthors{1}
+%\numberofauthors{1}
+%\author{
+%\alignauthor Paper  580
+%}
+
+\numberofauthors{4}
 \author{
-\alignauthor Paper  580
+\alignauthor Ryan C. Julian\\
+       \affaddr{Department of EECS}\\
+       \affaddr{Univ. of California, Berkeley}\\
+       \affaddr{Berkeley, CA 94720}\\
+       \email{ryanjulian@berkeley.edu}
+\alignauthor Cameron J. Rose\\
+       \affaddr{Department of EECS}\\
+       \affaddr{Univ. of California, Berkeley}\\
+       \affaddr{Berkeley, CA 94720}\\
+       \email{c\_rose@eecs.berkeley.edu}
+\alignauthor Humphrey Hu\\
+       \affaddr{Robotics Institute}\\
+       \affaddr{Carnegie Mellon University}\\
+       \affaddr{Pittsburgh, PA 15213}\\
+       \email{humhu@cmu.edu}
+\and
+\alignauthor Ronald S. Fearing\\
+       \affaddr{Department of EECS}\\
+       \affaddr{Univ. of California, Berkeley}\\
+       \affaddr{Berkeley, CA 94720}\\
+       \email{ronf@eecs.berkeley.edu}
 }
-
-%\numberofauthors{4}
-%\author{
-%\alignauthor Ryan C. Julian\\
-%       \affaddr{Department of EECS}\\
-%       \affaddr{Univ. of California, Berkeley}\\
-%       \affaddr{Berkeley, CA 94720}\\
-%       \email{ryanjulian@berkeley.edu}
-%\alignauthor Cameron J. Rose\\
-%       \affaddr{Department of EECS}\\
-%       \affaddr{Univ. of California, Berkeley}\\
-%       \affaddr{Berkeley, CA 94720}\\
-%       \email{c\_rose@eecs.berkeley.edu}
-%\alignauthor Humphrey Hu\\
-%       \affaddr{Robotics Institute}\\
-%       \affaddr{Carnegie Mellon University}\\
-%       \affaddr{Pittsburgh, PA 15213}\\
-%       \email{humhu@cmu.edu}
-%\and
-%\alignauthor Ronald S. Fearing\\
-%       \affaddr{Department of EECS}\\
-%       \affaddr{Univ. of California, Berkeley}\\
-%       \affaddr{Berkeley, CA 94720}\\
-%       \email{ronf@eecs.berkeley.edu}
-%}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Acknowledgments}
-%The authors would like to thank Fernando Garcia Bermudez for his 
-%assistance with the Vicon motion capture system, Andrew Pullin for his 
-%help with robot photography, and the members of the Biomimetic 
-%Millisystems Laboratory and the EECS community at the University of 
-%California, Berkeley for their advice and support.
+\section{Acknowledgments}
+The authors would like to thank Fernando Garcia Bermudez for his 
+assistance with the Vicon motion capture system, Andrew Pullin for his 
+help with robot photography, and the members of the Biomimetic 
+Millisystems Laboratory and the EECS community at the University of 
+California, Berkeley for their advice and support.
 

### 2013-01-16 manuscript.tex
Merge "Experiments" with "Results and Performance"

@@
 % METADATA
-\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with an 
-Ornithopter MAV and Lightweight Ground Station}
+\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with Millirobots}
 \def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
+\section{Experiments, Results, and Performance}
+\label{sec:performance}
+%TODO cameron
+To bound the performance of the cooperative control system, we determine 
+the feasible set of initial conditions that result in successful narrow 
+passage traversal, using both analytical and empirical methods. We then 
+conduct experiments to validate this model.
 
@@
 \subsubsection{Ornithoper Attitude Control}
+\label{sec:flight_control}
 We determined the turning speed and turning radius of the H$^2$Bird 
@@
 Sec.~\ref{sec:model_attitude}.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
+\caption{H$^2$Bird step response}
+\label{fig:step_response}
+\end{figure}
+%TODO reduce number of sig figs
+
+The first-order model that we fit to the step response of the onrithopter 
+yaw motion is in Equation~\ref{eq:transfer_func}. This model represents the 
+tuning dynamics of the ornithopter as well as the PID controller providing 
+inputs to the actuators. We present the results of the the step response 
+experiment, along with the response of the fitted model, in 
+Figure~\ref{fig:step_response}.
+\begin{equation}
+\label{eq:transfer_func}
+H(s) = \frac{0.95}{1+0.62s}
+\end{equation}
+The step response rise time is 1.4 seconds. The H$^2$Bird flies at an 
+average forward velocity of 1.2 m/s during our experiments, so we estimate 
+the minimum turning radius of the robot is 1.07 meters, as described in 
+Equation~\ref{eq:min_radius}.
+\begin{equation}
+\label{eq:min_radius}
+\begin{aligned}
+\text{Minimum radius} & = \frac{360^{\circ}\times\text{Rise time}\times\text{Flight
+speed}}{\text{Final angle}\times 2\pi}\\
+& = \frac{360^{\circ}\times 1.4s\times 1.2\frac{m}{s}}{90^{\circ}\times2*\pi} = 1.07 m
+\end{aligned}
+\end{equation}
+In the experiments we present in this work, we choose a flight speed lower 
+than the maximum speed of the ornithopter, to facilitate more robust 
+tracking.
 
@@
 
+We measured a camera capture latency 40 ms, pose estimation computation 
+latency of 12.5 ms, and radio latency of 25 ms.
 %----------------------------------------------------------------------------%
 \subsection{Model Verification}
+
+%............................................................................%
+\subsubsection{Simulation}
+
+\begin{figure*}[tb]
+\begin{minipage}[b]{0.45\linewidth}
+\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
+\caption{Plot of backwards reachable set for successful window traversal.}
+\label{fig:feasible_set}
+\end{minipage}
+\hfill
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\centering
+\includegraphics[width=\textwidth]{figures/heat_map.png}
+\caption{Probability of success for all starting locations. This graphic is 
+a result of a Monte Carlo simulation of the reachable set model, in which the 
+initial heading of the H$^2$Bird is chosen uniformly at random from $(-90^{\circ},90^{\circ})$.}
+\label{fig:heat_map}
+\end{minipage}
+\end{figure*}
+
+We determine the backwards reachable set of initial states for successful 
+narrow passage traversal both geometrically and using Monte Carlo simulation. 
+(Figs.~\ref{fig:feasible_set} and~\ref{fig:heat_map}) In Figure ~\ref{fig:feasible_set}, 
+the blue lines represent the camera viewing triangle, the black line represents the window, and the black 
+dot represents the camera. For the initial conditions, we used a uniform
+grid of 5 cm increments within the camera viewing region, and an initial
+angular heading perpendicular to the window plane. A success, or feasible 
+initial condition, consists of a path that intersects the window at some 
+point and does not leave the camera viewing triangle, and is indicated in 
+green. The geometrically infeasible region is computed using the minimum 
+turning radius and the geometry of the camera's field-of-view and is 
+indicated in blue. The region in red is not geometrically feasible, but is 
+infeasible as a result of the cooperative control simulation dynamics.
+
+We simulate the motion of the ornithopter beginning from various positional 
+and angular heading initial conditions. We use this model, as determined 
+through system identification (Sec.~\ref{sec:system_model}) using the 
+cooperative control method in Figure~\ref{fig:block_diagram}.
+
+We conducted a Monte Carlo simulation to determine the probability of 
+successful window navigation for a given point within the testing area. 
+Using this information we can make assumptions about ideal starting 
+positions for cooperative control. For each point within a 10 cm grid within 
+the camera viewing triangle, we simulated 40 trials using an initial heading 
+randomly sampled from a uniform distribution between -90$^{\circ}$ and 
+90$^{\circ}$. Figure~\ref{fig:heat_map} depicts the results of the Monte 
+Carlo simulation.
+
+As shown in the figure, there is a region in the middle with a 100\% 
+probability of success, which is a subset of the feasible region presented 
+in Figure~\ref{fig:feasible_set}. Along the edges of the camera viewing 
+triangle, however, the probability of success decreases.
+
+%............................................................................%
+\subsubsection{Experiments}
 \label{sec:experiments_verification}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and Performance}
-\label{sec:performance}
-%TODO cameron
-To bound the performance of the cooperative control system, we determine 
-the feasible set of initial conditions that result in successful narrow 
-passage traversal, using both analytical and empirical methods. We then 
-conduct experiments to validate this model.
-
-%----------------------------------------------------------------------------%
-\subsection{System Identification}
-
-%............................................................................%
-\subsubsection{Ornithoper Attitude Control}
-\label{sec:flight_control}
-%TODO cameron
-\begin{figure}[tb]
-\centering
-\includegraphics[width=\linewidth]{figures/step_response_total.pdf}
-\caption{H$^2$Bird step response}
-\label{fig:step_response}
-\end{figure}
-%TODO reduce number of sig figs
-
-The first-order model that we fit to the step response of the onrithopter 
-yaw motion is in Equation~\ref{eq:transfer_func}. This model represents the 
-tuning dynamics of the ornithopter as well as the PID controller providing 
-inputs to the actuators. We present the results of the the step response 
-experiment, along with the response of the fitted model, in 
-Figure~\ref{fig:step_response}.
-\begin{equation}
-\label{eq:transfer_func}
-H(s) = \frac{0.95}{1+0.62s}
-\end{equation}
-The step response rise time is 1.4 seconds. The H$^2$Bird flies at an 
-average forward velocity of 1.2 m/s during our experiments, so we estimate 
-the minimum turning radius of the robot is 1.07 meters, as described in 
-Equation~\ref{eq:min_radius}.
-\begin{equation}
-\label{eq:min_radius}
-\begin{aligned}
-\text{Minimum radius} & = \frac{360^{\circ}\times\text{Rise time}\times\text{Flight
-speed}}{\text{Final angle}\times 2\pi}\\
-& = \frac{360^{\circ}\times 1.4s\times 1.2\frac{m}{s}}{90^{\circ}\times2*\pi} = 1.07 m
-\end{aligned}
-\end{equation}
-In the experiments we present in this work, we choose a flight speed lower 
-than the maximum speed of the ornithopter, to facilitate more robust 
-tracking.
-
-%............................................................................%
-\subsubsection{Ground Station Pose Estimation}
-
-We measured a camera capture latency 40 ms, pose estimation computation 
-latency of 12.5 ms, and radio latency of 25 ms.
-
-%............................................................................%
-\subsubsection{Cooperative Visual Servoing}
-\label{sec:visual_servoing}
-
 \begin{figure*}[tb]
@@
 
-%----------------------------------------------------------------------------%
-\subsection{Model Verification}
-
-\begin{figure*}[tb]
-\begin{minipage}[b]{0.45\linewidth}
-\includegraphics[width=\textwidth]{figures/feasible_set.pdf}
-\caption{Plot of backwards reachable set for successful window traversal.}
-\label{fig:feasible_set}
-\end{minipage}
-\hfill
-\begin{minipage}[b]{0.45\linewidth}
-\centering
-\centering
-\includegraphics[width=\textwidth]{figures/heat_map.png}
-\caption{Probability of success for all starting locations. This graphic is 
-a result of a Monte Carlo simulation of the reachable set model, in which the 
-initial heading of the H$^2$Bird is chosen uniformly at random from $(-90^{\circ},90^{\circ})$.}
-\label{fig:heat_map}
-\end{minipage}
-\end{figure*}
-
-We simulate the motion of the ornithopter beginning from various positional 
-and angular heading initial conditions. We use this model, as determined 
-through system identification (Sec.~\ref{sec:system_model}) using the 
-cooperative control method in Figure~\ref{fig:block_diagram}.
-
-We determine the backwards reachable set of initial states for successful 
-narrow passage traversal both geometrically and in simulation 
-(Fig.~\ref{fig:feasible_set}). In the figure, the blue lines represent the 
-camera viewing triangle, the black line represents the window, and the black 
-dot represents the camera. For the initial conditions, we used a uniform
-grid of 5 cm increments within the camera viewing region, and an initial
-angular heading perpendicular to the window plane. A success, or feasible 
-initial condition, consists of a path that intersects the window at some 
-point and does not leave the camera viewing triangle, and is indicated in 
-green. The geometrically infeasible region is computed using the minimum 
-turning radius and the geometry of the camera's field-of-view and is 
-indicated in blue. The region in red is not geometrically feasible, but is 
-infeasible as a result of the cooperative control simulation dynamics.
-
-Our experimental data validates Figure~\ref{fig:feasible_set}, as there are 
-no successes within the geometrically infeasible region. There are a few 
-successes near the boundary of the infeasible region determined in 
-simulation. While there are successes along the boundary of the camera 
-viewing triangle, there are also many failures. We further explore this 
-behavior using Monte Carlo simulation.
-
-We conducted a Monte Carlo simulation to determine the probability of 
-successful window navigation for a given point within the testing area. 
-Using this information we can make assumptions about ideal starting 
-positions for coopertive control. For each point within a 10 cm grid within 
-the camera viewing triangle, we simulated 40 trials using an initial heading 
-randomly sampled from a uniform distribution between -90$^{\circ}$ and 
-90$^{\circ}$. Figure~\ref{fig:heat_map} depicts the results of the Monte 
-Carlo simulation.
-
-As shown in the figure, there is a region in the middle with a 100\% 
-probability of success, which is a subset of the feasible region presented 
-in Figure~\ref{fig:feasible_set}. Along the edges of the camera viewing 
-triangle, however, the probability of success decreases.
-
 The results of our experimental trials verify the results of the Monte Carlo
@@
 supports the reduced probability of success calculated by the simulation.
+
+Our experimental data validates Figure~\ref{fig:feasible_set}, as there are 
+no successes within the geometrically infeasible region. There are a few 
+successes near the boundary of the infeasible region determined in 
+simulation. While there are successes along the boundary of the camera 
+viewing triangle, there are also many failures. We further explore this 
+behavior using Monte Carlo simulation.
 

### 2013-01-17 manuscript.tex
some edits

@@
 
-We divide our model of the vehicle pose into two independent models for 
+We divide our model of the vehicle dynamics into two independent models for 
 simplicity: one for yaw motion, and one for translational motion.
@@
 model to limit complexity, and because it matches our physical 
-implementation. Likewise, we choose the reference heading angle as input to 
+implementation. Likewise, we choose the reference heading angle as the input to 
 the model, and model the dynamics of vehicle and on-board PID controller as 
-one system, whose output is the physical heading angle. We show the 
-resulting model form in Equation~\ref{eq:plain_model}.
+one system, whose output is the physical heading angle:
 \begin{equation}
 \label{eq:plain_model}
-H(s) = \frac{K}{1 + Rs}
+\begin{aligned}
+& H(s) = \frac{K}{1 + Rs}
+& \text{where }H(s) = \frac{\theta(s)}{u(s)}
+\end{aligned}
 \end{equation}
 To model the translational position of the ornithopter, we use a form of 
-Dubin's car model (Eq.~\ref{eq:dubin_car})~\cite{lavalle:planning}, where 
-$\phi_{max}$ is the maximum steering angle and $u_{s}$ is the forward speed. 
+Dubin's car model~\cite{lavalle:planning}: 
 \begin{equation}
@@
 \end{equation} 
-We modify this model to treat the angular position $\theta$ as an input 
-rather than a state. The overall motion of the system can be modeled as shown 
-in Equation~\ref{eq:dubin_car}, where $\theta$ is the time-domain output of 
-Equation~\ref{eq:plain_model} with the reference heading angle as input.
+In the car model, $\phi_{max}$ is the maximum steering angle and $u_{s}$ is
+the forward speed. We modify this model to treat the angular position $\theta$
+as an input rather than a state. The overall motion of the system can be
+modeled as shown in Equation~\ref{eq:dubin_car}, where $\theta$ is the
+time-domain output of Equation~\ref{eq:plain_model}, $\theta(t)$, with the
+reference heading angle, $u(t)$, as input.
 
@@
 
-We model the pose estimation computation delay, which include the latency of 
-updating the particle filter, calculating a new reference attitude, and 
-sending the reference attitude command the ornithopter via radio.
+We experienced latency during experiementation in estimating the pose of the
+ornithopter on the ground station. Referring to
+Figure~\ref{fig:block_diagram}, there were delays in the pose estimation block
+while updating the particle filter, in the PID controller block when
+calculating the new reference attitude, and in the transmission of the desired
+heading from the external controller to the controller on the robot via radio.
 
@@
 The first-order model that we fit to the step response of the onrithopter 
-yaw motion is in Equation~\ref{eq:transfer_func}. This model represents the 
-tuning dynamics of the ornithopter as well as the PID controller providing 
-inputs to the actuators. We present the results of the the step response 
-experiment, along with the response of the fitted model, in 
-Figure~\ref{fig:step_response}.
+yaw motion is as folows:
 \begin{equation}
@@
 \end{equation}
+This model represents the tuning dynamics of the ornithopter as well as the
+PID controller providing inputs to the actuators. We present the results of
+the the step response experiment, along with the response of the fitted model,
+in Figure~\ref{fig:step_response}.
+
 The step response rise time is 1.4 seconds. The H$^2$Bird flies at an 
 average forward velocity of 1.2 m/s during our experiments, so we estimate 
-the minimum turning radius of the robot is 1.07 meters, as described in 
-Equation~\ref{eq:min_radius}.
+the minimum turning radius of the robot is 1.07 meters:
 \begin{equation}
@@
 
-We measured a camera capture latency 40 ms, pose estimation computation 
+We measured a camera capture latency of 40 ms, pose estimation computation 
 latency of 12.5 ms, and radio latency of 25 ms.
@@
 \begin{figure*}[tb]
-\begin{minipage}[b]{0.45\linewidth}
+\begin{minipage}[t]{0.45\linewidth}
 \includegraphics[width=\textwidth]{figures/feasible_set.pdf}
@@
 \hfill
-\begin{minipage}[b]{0.45\linewidth}
+\begin{minipage}[t]{0.47\linewidth}
 \centering
 \centering
-\includegraphics[width=\textwidth]{figures/heat_map.png}
+\includegraphics[width=\textwidth]{figures/heat_map.pdf}
 \caption{Probability of success for all starting locations. This graphic is 
@@
 
-As shown in the figure, there is a region in the middle with a 100\% 
-probability of success, which is a subset of the feasible region presented 
-in Figure~\ref{fig:feasible_set}. Along the edges of the camera viewing 
-triangle, however, the probability of success decreases.
+As shown in the figure, there is a region in the middle where we had success
+in all 40 of the trials, which is a subset of the feasible region presented in
+Figure~\ref{fig:feasible_set}. Along the edges of the camera viewing triangle,
+however, the probability of success decreases.
 

### 2013-01-17 manuscript.tex
Added MAST and NSF funding acknowledgements
Correct spelling of "ornithopter"

@@
 requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
-station provides heading estimates to the ornithoper by running a real-time 
+station provides heading estimates to the ornithopter by running a real-time 
 motion tracking algorithm on a live video stream.
@@
 %----------------------------------------------------------------------------%
-\subsection{H$^2$Bird Ornithoper}
+\subsection{H$^2$Bird Ornithopter}
 % humphrey
@@
 %----------------------------------------------------------------------------%
-\subsection{Ornithoper Attitude Control}
+\subsection{Ornithopter Attitude Control}
 
@@
 %............................................................................%
-\subsubsection{Ornithoper Attitude Control}
+\subsubsection{Ornithopter Attitude Control}
 \label{sec:flight_control}
@@
 Millisystems Laboratory and the EECS community at the University of 
-California, Berkeley for their advice and support.
-
+California, Berkeley for their advice and support. This material is based upon 
+work supported by the U.S. Army Research Laboratory under the Micro Autonomous 
+Systems and Technology Collaborative Technology Alliance, and the National 
+Science Foundation under Grant No. IIS-0931463.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-01-18 manuscript.tex
Remove millirobots from title
Swap H2Bird diagram with block diagram

@@
 % METADATA
-\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with Millirobots}
+\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal}
 \def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
@@
 \subsection{H$^2$Bird Ornithopter}
-% humphrey
+
+\begin{figure}[!tb]
+\centering
+\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
+\caption{H$^2$Bird ornithopter with attitude control axes and labeled 
+control surfaces.}
+\label{fig:h2Bird_axes}
+\end{figure}
+
 To explore these cooperative control concepts, we designed a new 
@@
 \subsection{Ground Station Computer Vision Platform}
-% ryan
 Cooperative behavior is a promising strategy for autonomous navigation in 
@@
 computer vision libraries for image capture and processing.
+%TODO mention code is open source, available on GitHub
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{System Concept and Algorithms}
-%TODO ryan
+
+%----------------------------------------------------------------------------%
+\subsection{Ornithopter Attitude Control}
+
+We implement attitude estimation and control of the H$^2$Bird on-board to 
+achieve high sample-rate attitude control. We estimate the vehicle pose by
+integrating the IMU, and use a proportional-integral-derivative (PID) 
+controller to feed these estimates back to the attitude control flight 
+surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).
 
@@
 \end{figure}
-%TODO h2bird label in diagram should have superscript
-
-%----------------------------------------------------------------------------%
-\subsection{Ornithopter Attitude Control}
-
-\begin{figure}[!tb]
-\centering
-\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
-\caption{H$^2$Bird ornithopter with attitude control axes and labeled 
-control surfaces.}
-\label{fig:h2Bird_axes}
-\end{figure}
-%TODO label h2bird drawing
-
-We implement attitude estimation and control of the H$^2$Bird on-board to 
-achieve high sample-rate attitude control. We estimate the vehicle pose by
-integrating the IMU, and use a proportional-integral-derivative (PID) 
-controller to feed these estimates back to the attitude control flight 
-surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).
 
@@
 \label{sec:performance}
-%TODO cameron
 To bound the performance of the cooperative control system, we determine 

### 2013-01-19 manuscript.tex
fixed the strange figure text sizes
edits

--- 2013-01-17 manuscript.tex
@@
 
+In this section, we describe the models that we developed to simulate the
+behaviors of the ornithopter and the ground station. We use these models to
+determine the set of initial conditions that are most likely to result in
+successful window traversal.
+
 %----------------------------------------------------------------------------%
@@
 model to limit complexity, and because it matches our physical 
-implementation. Likewise, we choose the reference heading angle as the input to 
-the model, and model the dynamics of vehicle and on-board PID controller as 
-one system, whose output is the physical heading angle:
+implementation:
 \begin{equation}
@@
 \end{equation}
-To model the translational position of the ornithopter, we use a form of 
-Dubin's car model~\cite{lavalle:planning}: 
+We choose the reference heading angle as the input to the model, $u(s)$, and
+model the dynamics of vehicle and on-board PID controller as one system, whose
+output is the physical heading angle, $\theta(s)$. To model the translational position of
+the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning}:
 \begin{equation}
@@
 modeled as shown in Equation~\ref{eq:dubin_car}, where $\theta$ is the
-time-domain output of Equation~\ref{eq:plain_model}, $\theta(t)$, with the
-reference heading angle, $u(t)$, as input.
+time-domain output of Equation~\ref{eq:plain_model}, $\theta(t)$.
 
@@
 640 pixels. We calculate the desired heading angle in degrees, for a given 
-distance from the camera, according to Equation~\ref{eq:desired_heading}.
+distance from the camera:
 \begin{equation} 

### 2013-01-19 manuscript.tex
more edits to sections 4 and 5

@@
 
-Our model of the ground station comprises models of both the camera and the
+Our model of the ground station includes models for both the camera and the
 pose estimation computation. The camera model contains both a latency, which 
@@
 average forward velocity of 1.2 m/s during our experiments, so we estimate 
-the minimum turning radius of the robot is 1.07 meters:
+the minimum turning radius of the robot to be 1.07 meters:
 \begin{equation}
@@
 application which draws a large colored rectangle on the ground station's
-screen, then uses the camera to detect that changes in that rectangle's 
-color. By noting the value of the system timer just before it changes the 
-rectangle, and just after the camera detects that change, the application 
-arrives at an estimate of the video capture pipeline latency. We averaged 
-these result over several hundred rapidly-executed trials to determine the 
-capture latency we used to model the system.
+screen, then uses the camera to detect the changes in the rectangle's color.
+By noting the value of the system timer just before the rectangle changes
+color and just after the camera detects the change, the application records
+the estimate of the video capture pipeline latency. We averaged the result
+over several hundred rapidly-executed trials to determine the capture latency
+we used to model the system.
 
@@
 narrow passage traversal both geometrically and using Monte Carlo simulation.
-(Figs.~\ref{fig:feasible_set} and~\ref{fig:heat_map}) In Figure ~\ref{fig:feasible_set}, 
+We simulate the motion of the ornithopter beginning from various positional
+and angular heading initial conditions. We use our model, as determined
+through system identification as described in Section~\ref{sec:system_model},
+using the cooperative control method in Figure~\ref{fig:block_diagram}. The
+results are presented in Figures~\ref{fig:feasible_set}
+and~\ref{fig:heat_map}.
+
+In Figure ~\ref{fig:feasible_set}, 
 the blue lines represent the camera viewing triangle, the black line represents the window, and the black 
@@
 
-We simulate the motion of the ornithopter beginning from various positional 
-and angular heading initial conditions. We use this model, as determined 
-through system identification (Sec.~\ref{sec:system_model}) using the 
-cooperative control method in Figure~\ref{fig:block_diagram}.
-
 We conducted a Monte Carlo simulation to determine the probability of
-successful window navigation for a given point within the testing area. 
-Using this information we can make assumptions about ideal starting 
-positions for cooperative control. For each point within a 10 cm grid within 
-the camera viewing triangle, we simulated 40 trials using an initial heading 
-randomly sampled from a uniform distribution between -90$^{\circ}$ and 
-90$^{\circ}$. Figure~\ref{fig:heat_map} depicts the results of the Monte 
-Carlo simulation.
+successful window navigation for a given point within the testing area. Using
+this information we can make assumptions about ideal starting positions for
+cooperative control. For each point within a 10 cm grid within the camera
+viewing triangle, we simulated 40 trials using an initial heading randomly
+sampled from a uniform distribution between -90$^{\circ}$ and 90$^{\circ}$,
+with the 0$^{\circ}$ heading perpendicular to the window.
+Figure~\ref{fig:heat_map} depicts the results of the Monte Carlo simulation.
 
@@
 As the H$^2$Bird cannot detect the target on its own, we used cooperative 
-control with the ground station, which to provided remote guidance to 
-H$^2$Bird to allow it to navigate to the window. We used a motion capture 
+control with the ground station, which provided remote guidance to 
+H$^2$Bird to allow it to navigate through the window. We used a motion capture 
 system to record ground truth data. We present a sketch of this experimental 
@@
 \item We identify the window to the ground station by dragging a bounding 
-box over it using the mouse. 
+box over it using a mouse. 
 \item We release H$^2$Bird by hand, in view of the camera at the desired 
 starting grid point.
-\item The H$^2$Bird attempts to fly through the window with, with remote 
+\item The H$^2$Bird attempts to fly through the window, with remote 
 guidance from the ground station.
@@
 
-After every 5 trials, we replaced the ornithopter's battery with a fully 
+Every 5 trials, we replaced the ornithopter's battery with a fully 
 charged one.
 
-In total, we collected 80 trials over a variety of starting positions. Our 
+We collected 80 trials over a variety of starting positions. Our 
 goal for these experiments was to verify the model described in 
 Section~\ref{sec:system_model}, so we conducted 60 of the 80 trials in a 
-grid along the edges of the camera view plane, and in 5 cm meter increments 
-from the window, in an effort to determine the success rate on the edges of 
+grid along the edges of the camera view plane, and in 5 cm increments 
+from the window plane, in an effort to determine the success rate on the edges of 
 the testing space. We conducted 20 additional trials directly in front of 
@@
 of the camera. As we moved towards the edges of the feasible region in 
-Figure~\ref{fig:flight_paths}, the success rate diminished to 50\%. We 
-present the results of our experimental trials in Figures~\ref{fig:flight_paths} 
-and~\ref{fig:flight_paths_feasible}.
+Figure~\ref{fig:flight_paths}, the success rate diminished to 50\%.
 
-The lower success rate on the edges of the camera frame is explained by 
+The lower success rate on the edges of the camera frame can be explained by 
 several factors. On the edges of the viewing triangle, the yaw control 

### 2013-01-24 manuscript.tex

@@
 %----------------------------------------------------------------------------%
-\subsection{H$^2$Bird Ornithoper}
-% humphrey
+\subsection{H$^2$Bird Ornithopter}
+
+\begin{figure}[!tb]
+\centering
+\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
+\caption{H$^2$Bird ornithopter with attitude control axes and labeled 
+control surfaces.}
+\label{fig:h2Bird_axes}
+\end{figure}
+
 To explore these cooperative control concepts, we designed a new 
@@
 \end{figure}
-%TODO h2bird label in diagram should have superscript
-
-%----------------------------------------------------------------------------%
-\subsection{Ornithoper Attitude Control}
-
-\begin{figure}[!tb]
-\centering
-\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
-\caption{H$^2$Bird ornithopter with attitude control axes and labeled 
-control surfaces.}
-\label{fig:h2Bird_axes}
-\end{figure}
-%TODO label h2bird drawing
-
-We implement attitude estimation and control of the H$^2$Bird on-board to 
-achieve high sample-rate attitude control. We estimate the vehicle pose by
-integrating the IMU, and use a proportional-integral-derivative (PID) 
-controller to feed these estimates back to the attitude control flight 
-surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).
 

### 2013-01-24 manuscript.tex
Integrate Ethan's and some of Fernando's comments

@@
 less mobility than MAVs, but relaxed size constraints, and typically more 
-computing power. These specializations present many opportunities for cooperation. 
+computing power. These specializations present many opportunities for 
+robot-robot cooperation. 
 In this work, we demonstrate cooperative target-seeking between a 13 gram 
@@
 accurately predicts the backwards-reachable region for successfully 
-negotiating ornithopther flight through narrow openings.
+negotiating ornithopther flight through narrow passages.
 
-We introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
+We also introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
 features clap and fling wings, improves upon previous designs by 
@@
 2.8 gram payload. We augment the ornithopter's built-in gyroscope-based 
-control with a lightweight ground station design, which has power and weight 
-requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
-station provides heading estimates to the ornithopter by running a real-time 
-motion tracking algorithm on a live video stream.
+control with a lightweight ground station, which has power and weight 
+requirements appropriate for deployment on a ground vehicles with 10 gram 
+payloads. The ground station provides heading estimates to the ornithopter by 
+running a real-time motion tracking algorithm over a live video stream.
 \end{abstract}
@@
 Power, size, and weight constraints on micro air vehicles (MAV) 
-significantly limit their sensing and computational capabilities. They can 
+significantly limit their sensing and computational capabilities. MAVs can 
 execute only the most rudimentary control and state estimation algorithms in 
@@
 The non-linear dynamics of flapping-wing flight compound 
-these constraints~\cite{humbert:dipteran2}\cite{cheng:tandrdamping}. Much 
+these constraints~\cite{cheng:tandrdamping}\cite{humbert:dipteran2}. Much 
 past research on flapping-wing MAVs focuses on overcoming the difficulties 
@@
 dynamic models include averaging over the wing-beat 
-period~\cite{Schenato:flightcontrol}\cite{cheng:tandrdamping}, and using 
+period~\cite{cheng:tandrdamping}\cite{Schenato:flightcontrol}, and using 
 linear, low-dimensional models to predict flight 
-behavior~\cite{Shigeoka:ornithopter}. These advances lead to the development 
+behavior~\cite{Shigeoka:ornithopter}. These advances led to the development 
 of high-performance platforms such as the sub-15 gram DelFly, described by 
@@
 
-Many control laws previous researchers developed from dynamic models of 
+Many control laws researchers have developed from dynamic models of 
 flapping-wing flight, such as periodic forcing and other control methods 
@@
 passages, by exploiting cooperation between specialized robotic agents.
+%
 
@@
 rotary-wing MAV platforms, but many of these methods make extensive use of 
-GPS or motion capture systems~\cite{kingston:timeattitude}\cite{kanade:3dvision}. 
-Indoors, using GPS to estimate the pose of an MAV is impossible, and motion 
+GPS or motion capture systems~\cite{kanade:3dvision}\cite{kingston:timeattitude}. 
+Indoors, using GPS to estimate the pose of an MAV is unreliable, and motion 
 capture requires extensive and stationary modification of the environment. 
@@
 The literature on cooperative control of MAVs is similarly limited to larger
-vehicles and purely analytical explorations. Nebot, Hyams, and Jung all 
-demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Nebot2003Agents}\cite{Hyams1999Cooperative}\cite{Jung1998Range}. 
-Luo and Mehta analyze  UAV-UGV cooperation from a modeling perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. 
-Rudol demonstrated a cooperative visual servoing system similar with 
+vehicles and purely analytical explorations. Hyams et al., Jung et al. and Nebot et al. all 
+demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}. 
+Luo, et al. and Mehta, et al. analyze  UAV-UGV cooperation from a modeling perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. 
+Rudol et al. demonstrated a similar cooperative visual servoing system with 
 UGVs and MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of 
-magnitude larger than the millirobotic systems we target, and so were able 
+magnitude larger than the lightweight systems we target, and so were able 
 to take advantage of high-performance processors and specialized cameras.
 
-The small size of millirobotic air and ground vehicles necessitates 
-specialization, but it also pays dividends agility. A group of specialized 
-heterogeneous millirobots can navigate more challenging terrains and 
+The small size of lightweight air and ground vehicles necessitates 
+specialization, but it also pays dividends in agility. A group of specialized 
+heterogeneous, lightweight robots can navigate more challenging terrains and 
 interact with their environments in ways that larger, monolithic vehicles 
 cannot. Despite this advantage, in order to collectively equal the sensing and 
-computational capabilities of larger robots, they will also have to interact 
+computational capabilities of larger robots, small robots will also have to interact 
 so that multiple robots can benefit from the specialized capabilities of 
-cooperating vehicles. Millirobotic systems will benefit from cooperation 
-even more than their larger counterparts.
+cooperating vehicles.
 
@@
 processor modules--such as the Gumstix Overo and LogicPD Torpedo--which are 
-light enough for deployment on millirobotic ground vehicles. We pair the 
+light enough for deployment on lightweight ground vehicles. We pair the 
 BeagleBoard with an off-the-shelf consumer USB web camera, for similar image 
 quality and resolution to the miniature camera modules available at 
-millirobot scale. The BeagleBoard communicates with the H$^2$Bird with a 
+millirobot scale. The BeagleBoard communicates with the H$^2$Bird via a 
 USB radio module which implements the IEEE 802.15.4 wireless standard.
@@
 modified with a custom version of the Linux kernel. 
-We use the Python\footnote{\raggedright Python: Python Software Federation, \href{http://www.python.org/}
+We use the Python\footnote{\raggedright Python: Python Software Foundation, \href{http://www.python.org/}
 {http://www.python.org/}} programming language for control, robot 
@@
 transitions increase tracking robustness, by forcing the filter to always 
-sample broadly across of the pixel space. This helps prevent overconfidence 
+sample broadly across the pixel space. This helps prevent overconfidence 
 and collapse, and decreases reacquisition delay.
@@
 station then commands a reference attitude to the H$^2$Bird over radio, 
-according to a PID control law. We apply a fixed-lag window the PID 
+according to a PID control law. We apply a fixed-lag window to the PID 
 controller integral term to prevent wind-up while we launch the ornithopter. 
@@
 narrow passage traversal both geometrically and using Monte Carlo simulation.
-We simulate the motion of the ornithopter beginning from various positional
+We also simulate the motion of the ornithopter beginning from various position
 and angular heading initial conditions. We use our model, as determined
@@
 Our experimental data validates Figure~\ref{fig:feasible_set}, as there are 
-no successes within the geometrically infeasible region. There are a few 
+no successes within the geometrically infeasible region. There are few 
 successes near the boundary of the infeasible region determined in 
@@
 viewing triangle, there are also many failures. We further explore this 
-behavior using Monte Carlo simulation.
+behavior using a Monte Carlo simulation.
 

### 2013-01-25 manuscript.tex
Integrate Ron's brief edits

@@
 % METADATA
-\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal}
+\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with Millirobots}
 \def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
@@
 \end{figure}
-Power, size, and weight constraints on micro air vehicles (MAV) 
+Power, size, and weight constraints on 10 gram scale micro air vehicles (MAV) 
 significantly limit their sensing and computational capabilities. MAVs can 
-execute only the most rudimentary control and state estimation algorithms in 
+execute only low-complexity control and state estimation algorithms in 
 real-time.  The constraints on flapping-wing MAVs are even more restrictive. 
@@
 The literature on cooperative control of MAVs is similarly limited to larger
-vehicles and purely analytical explorations. Hyams et al., Jung et al. and Nebot et al. all 
+vehicles or analytical explorations. Hyams et al., Jung et al. and Nebot et al. all 
 demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}. 
@@
 so that multiple robots can benefit from the specialized capabilities of 
-cooperating vehicles.
+cooperating vehicles. In this paper we consider both experimentally, and in 
+simulation, the performance of the externally-directed ornithopter to 
+demonstrate how millirobotic system can benefit from cooperation.
 
@@
 The ground station performs simple two-dimensional pose estimation on 
-H$^2$Bird using a particle filter-based motion tracking algorithm. This pose 
+H$^2$Bird using a particle filter-based motion tracking algorithm~\cite{thrun2005probabilistic}. This pose 
 estimate is the feedback input to the cooperative visual servoing feedback 
@@
 
-Our pose estimation algorithm tracks H$^2$Bird's motion in the frame using a 
-particle filter~\cite{thrun2005probabilistic}. We initialize the particles 
+We initialize the particles 
 uniformly across the frame before tracking begins. To improve numerical 
@@
 \label{sec:visual_servoing_concept}
-The ground station guides the H$^2$Bird by commanding changes to its 
-reference attitude, which is in turn governed by its on-board attitude 
-controllers (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
+The ground station guides the H$^2$Bird by setting the
+reference attitude of its on-board attitude 
+controller (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
 the estimate of 2D vehicle pose from the pose estimation algorithm, the 
 ground station calculates an altitude error and horizontal error. The ground 
-station then commands a reference attitude to the H$^2$Bird over radio, 
-according to a PID control law. We apply a fixed-lag window to the PID 
+station then uses a PID control law to calculate a new reference attitude, 
+which it sends to the H$^2$Bird over radio. 
+We apply a fixed-lag window to the PID 
 controller integral term to prevent wind-up while we launch the ornithopter. 

### 2013-01-28 manuscript.tex
more changes

@@
 represent relative body and world coordinate angular displacements compactly 
-with quaternion multiplication.
+with quaternion multiplication~\cite{hamilton1866elements}.
 
@@
 error as the rotation $q_e$ required to reach the reference pose from the 
-estimated pose $q$ (Eq.~\ref{eq:quat_error}). 
+estimated pose $q$:
 \begin{align}
@@
 We divide our model of the vehicle dynamics into two independent models for 
-simplicity: one for yaw motion, and one for translational motion.
+simplicity. We develop one model for yaw motion, and another for translational motion.
 
-We approximate the yaw dynamics of the ornithopter with a one-dimensional 
-model to limit complexity, and because it matches our physical 
-implementation:
+We approximate the yaw dynamics of the ornithopter with a one-dimensional model to 
+limit complexity, and because it matches our physical implementation. The H$^2$Bird 
+uses an on-board PID controller to follow a particular heading sent by the ground 
+station, and the yaw dynamics are largely decoupled from the rest of the fight dynamics, 
+so our one-dimensional model is a good representation of the yaw motion:
 \begin{equation}
@@
 model the dynamics of vehicle and on-board PID controller as one system, whose
-output is the physical heading angle, $\theta(s)$. To model the translational position of
+output is the physical heading angle of the H$^2$Bird, $\theta(s)$. To model the translational position of
 the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning}:
@@
 \subsection{System Identification}
+
+In this section, we present the numerical results of the experiments we used to 
+determine the values of the parameters in our system model.
 
@@
 \subsection{Model Verification}
+
+We validate our system model experimentally by systematically attempting to
+guide the H$^2$Bird through the window using the directed headings from the
+ground station at various starting initial conditions. We compare the results of the
+experiments to the computed backwards reachable set of feasible initial
+conditions and a Monte Carlo simulation to determine the accuracy of our system 
+model.
 
@@
 launch can cause the system to respond differently for similar initial 
-conditions. In addition, we changed H$^2$Bird battery approximately every 5 
+conditions. In addition, we changed the H$^2$Bird battery approximately every 5 
 tests, adding more variation to the H$^2$Bird performance by affecting the 

### 2013-01-29 manuscript.tex
new item class for list with custom bullets

@@
 }
+
+% descriptor list for storyboard figure
+\newenvironment{spaced_description}[1]
+  {\begin{list}{}%
+   {\renewcommand\makelabel[1]{##1:\hfill}%
+   \settowidth\labelwidth{\makelabel{#1}}%
+   \setlength\leftmargin{\labelwidth}
+   \addtolength\leftmargin{\labelsep}}}
+  {\end{list}}
 
@@
 
-We validate our system model experimentally by systematically attempting to
-guide the H$^2$Bird through the window using the directed headings from the
-ground station at various starting initial conditions. We compare the results of the
-experiments to the computed backwards reachable set of feasible initial
-conditions and a Monte Carlo simulation to determine the accuracy of our system 
-model.
+We validate our system model experimentally by systematically guiding the
+H$^2$Bird through the window at various starting initial conditions. We
+compare the results of the experiments to the computed backwards reachable set
+of feasible initial conditions and a Monte Carlo simulation to determine the
+accuracy of our system model.
 
@@
 Each trial included the following steps (see Fig.~\ref{fig:pf_screencap}):
-\begin{enumerate}
-\item We identify the window to the ground station by dragging a bounding 
+\begin{spaced_description}{\textbf{(4-6)}}
+\item[\textbf{(1,2)}] We identify the window to the ground station by dragging a bounding 
 box over it using a mouse. 
-\item We release H$^2$Bird by hand, in view of the camera at the desired 
+\item[\textbf{(3)}] We release H$^2$Bird by hand, in view of the camera at the desired 
 starting grid point.
-\item The H$^2$Bird attempts to fly through the window, with remote 
+\item[\textbf{(4-6)}] The H$^2$Bird attempts to fly through the window, with remote 
 guidance from the ground station.
-\end{enumerate}
+\end{spaced_description}
 

### 2013-01-30 manuscript.tex
final commit before going home

@@
 \section{Robotic and Vision Platforms}
+
+We develop a new flapping-winged MAV and
+
 %----------------------------------------------------------------------------%
@@
 ground station calculates an altitude error and horizontal error. The ground 
-station then uses a PID control law to calculate a new reference attitude, 
+station then uses a PID control law to calculate a new reference pitch and yaw, 
 which it sends to the H$^2$Bird over radio. 
@@
 dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
-Particles converge on its location; 4.~Release H$^2$Bird, which 
+Particles converge on its location; 4.~Release H$^2$Bird, which h
 begins cooperative autonomous flight toward the target; 5.~Visual servoing 

### 2013-01-31 manuscript.tex
added Stirling citation to bib and intro

@@
 of controlling and modeling flapping-wing vehicles in the face of this
-non-linearity\cite{floreano:flying}. Common approaches to simplify these 
+non-linearity~\cite{floreano:flying}. Common approaches to simplify these 
 dynamic models include averaging over the wing-beat 
@@
 The literature on cooperative control of MAVs is similarly limited to larger
-vehicles or analytical explorations. Hyams et al., Jung et al. and Nebot et al. all 
-demonstrated cooperative visual servoing for heterogeneous UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}. 
-Luo, et al. and Mehta, et al. analyze  UAV-UGV cooperation from a modeling perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. 
-Rudol et al. demonstrated a similar cooperative visual servoing system with 
-UGVs and MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of 
-magnitude larger than the lightweight systems we target, and so were able 
-to take advantage of high-performance processors and specialized cameras.
+vehicles or analytical explorations. Hyams et al., Jung et al. and Nebot et
+al. all demonstrated cooperative visual servoing for heterogeneous
+UGVs~\cite{Hyams1999Cooperative}\cite{Jung1998Range}\cite{Nebot2003Agents}.
+Luo, et al. and Mehta, et al. analyze UAV-UGV cooperation from a modeling
+perspective~\cite{Luo2011Air}\cite{Mehta2006Adaptive}. Rudol et al.
+demonstrated a similar cooperative visual servoing system with UGVs and
+MAVs~\cite{Rudol2008Micro}, but the vehicles used were orders of magnitude
+larger than the lightweight systems we target, and so were able to take
+advantage of high-performance processors and specialized cameras. Stirling et
+al. describe a method for guiding quadrotors, designated as explorers, though
+a network of stationary quadrotors, designated as beacons, through an indoor
+environment~\cite{SwarmStirling}. The beacons provide external sensing for
+navigating the explorers through the environment, and the network can be
+extended by using explorers as beacons in undiscovered areas. Their
+implementation relies solely on local sensing, although our size range
+requires that the actual environment sensing be external to the MAV.
 

### 2013-02-01 manuscript.tex
updates

@@
 
-We develop a new flapping-winged MAV and
+We develop a new flapping-winged MAV, the H$^2$Bird, and ground station. We
+use these sytems to cooperatively navigate through a narrow passage, by
+sensing the position of the MAV and the goal using the ground station.
 
@@
 
-The step response rise time is 1.4 seconds. The H$^2$Bird flies at an 
+We calculated the step response rise time to be 1.4 seconds. The H$^2$Bird flies at an 
 average forward velocity of 1.2 m/s during our experiments, so we estimate 

### 2013-02-01 manuscript.tex
fixed figure references

@@
 To explore these cooperative control concepts, we designed a new 
-flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{fig:h2bird}). Built 
+flapping-wing MAV, known as the H$^2$Bird (Figure~\ref{fig:h2bird}). Built 
 around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
@@
 tail propeller and servo-controlled tail elevator provide control in the yaw 
-and pitch axes (Fig.~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
+and pitch axes (Figure~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
 \href{https://github.com/biomimetics/imageproc\_pcb}
@@
 controller to feed these estimates back to the attitude control flight 
-surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).
+surfaces (Figures~\ref{fig:process_flow} and~\ref{fig:block_diagram}).
 
@@
 We then convert this error quaternion to an angle axis representation and 
-use it as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
+use it as input to the PID controller (Equations~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
 We perform these calculations on-board, with trigonometric lookup tables for 
@@
 loop, which is represented as the outermost feedback paths on the block 
-diagrams in Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}. We 
+diagrams in Figures~\ref{fig:process_flow} and~\ref{fig:block_diagram}. We 
 limit our pose estimation approach to monocular two-dimensional object 
@@
 \subsubsection{Motion Model}
-Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
+Our motion model (Equation~\ref{eq:motion_model}) is a simple Gaussian transition 
 centered around each particle's current position, augmented with an 
-$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
+$\epsilon$-random uniform sampling strategy. In Equation~\ref{eq:motion_model}, 
 $x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
@@
 \subsubsection{Emission Model}
-Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
+Our emission model (Equation~\ref{eq:emission_model}) implements a sampling
 approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
@@
 reference attitude of its on-board attitude 
-controller (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
+controller (Figures~\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
 the estimate of 2D vehicle pose from the pose estimation algorithm, the 
@@
 to fit a simple low-order model to the step response, as discussed in 
-Sec.~\ref{sec:model_attitude}.
+Section~\ref{sec:model_attitude}.
 
@@
 
-Each trial included the following steps (see Fig.~\ref{fig:pf_screencap}):
+Each trial included the following steps (see Figure~\ref{fig:pf_screencap}):
 \begin{spaced_description}{\textbf{(4-6)}}
@@
 \section{Conclusions and Future Work}
-We have demonstrated cooperative guidance of a 13 gram flapping-wing MAV with a
-lightweight base station in hardware. In addition, we developed a model of the
-cooperative system using system identification methods on the individual system capabilities, and composing
-it with known system interactions. The corroboration between our simulated and
-experimental results shows that our simplified modeling approach is useful for 
-predicting the performance of cooperative systems at low computational cost.
-Our model could also be used to determine when to begin control with high probability of success, even though we lack information about the complete pose of the MAV.
+We have demonstrated cooperative guidance of a 13 gram flapping-wing MAV with
+a lightweight base station in hardware. In addition, we developed a model of
+the cooperative system using system identification methods on the individual
+system capabilities, and composing it with known system interactions. The
+corroboration between our simulated and experimental results shows that our
+simplified modeling approach is useful for predicting the performance of
+cooperative systems at low computational cost. Our model could also be used to
+determine when to begin control with high probability of success, even though
+we lack information about the complete pose of the MAV.
 

### 2013-02-05 manuscript.tex
Figures out of references

@@
 \pdfpageheight=11truein
+
+\renewcommand{\topfraction}{0.95}
+\renewcommand{\textfraction}{0.05}
+\renewcommand{\floatpagefraction}{0.90}
 
@@
 \includegraphics[width=\textwidth]{figures/heat_map.png}
-\caption{Probability of success for all starting locations. This graphic is 
-a result of a Monte Carlo simulation of the reachable set model, in which the 
-initial heading of the H$^2$Bird is chosen uniformly at random from $(-90^{\circ},90^{\circ})$.}
+\caption{Probability of success, resulting from a Monte Carlo simulation of
+the reachable set model, for all starting locations.}
+%\caption{Probability of success for all starting locations. This graphic is 
+%a result of a Monte Carlo simulation of the reachable set model, in which the 
+%initial heading of the H$^2$Bird is chosen uniformly at random from $(-90^{\circ},90^{\circ})$.}
 \label{fig:heat_map}
@@
 \textcolor{blue}{blue}, and the target location as a \textcolor{green}{green} 
-circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
-dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
-Particles converge on its location; 4.~Release H$^2$Bird, which h
-begins cooperative autonomous flight toward the target; 5.~Visual servoing 
-controller overshoots the window on the left, and commands a right-turn 
-maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
+circle.}
+%circle. 1.~Particles initialized uniformly across the frame; 2.~Human operator 
+%dragging bounding box across the window; 3.~H$^2$Bird enters the frame. 
+%Particles converge on its location; 4.~Release H$^2$Bird, which h
+%begins cooperative autonomous flight toward the target; 5.~Visual servoing 
+%controller overshoots the window on the left, and commands a right-turn 
+%maneuver to recover; 6.~H$^2$Bird successfully navigates through the window.}
 \label{fig:pf_screencap}

### 2013-02-05 manuscript.tex
more edits

@@
 \section{System Concept and Algorithms}
+
+Our system combines both the sensing of the H$^2$Bird and the execution of the
+desired headings to steer the robot towards the goal. The sensing is performed
+on the ground station using frame differencing and a particle filter. The
+H$^2$Bird steers to the heading directed by the ground station using a
+quaternion-based PID controller.
 

### 2013-02-05 manuscript.tex
some section 3 edits

@@
 requires that the actual environment sensing be external to the MAV.
+Additionally, they do not consider time constraints in their planning, whereas our MAV has a minimum forward speed and turning radius.
 
@@
 centered around each particle's current position, augmented with an 
-$\epsilon$-random uniform sampling strategy. In Equation~\ref{eq:motion_model}, 
-$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
-time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
-bounds of the pixel space.
-\begin{equation}
+$\epsilon$-random uniform sampling strategy. The sampling law is as follows:\begin{equation}
 \label{eq:motion_model}
@@
 \end{equation}
+In Equation~\ref{eq:motion_model}, $x^{[i]}_t$ is the position of the $i^\text{th}$ 
+particle in pixel space at time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ 
+is a vector of the bounds of the pixel space. The sampling method requires that, with 
+probability $\epsilon$, the particle filter will choose a pixel from a uniform 
+distribution of all of the pixels in the pixel space for the location of the $i^\text{th}$ 
+particle. The $\epsilon$-random uniform transitions increase tracking robustness, by 
+forcing the filter to always sample broadly across the pixel space. This helps prevent 
+overconfidence and collapse, and decreases reacquisition delay.
+
 In the experiments we present in this work, we used a diagonal covariance matrix 
@@
 transition model is sufficient for tracking motion in which frame-to-frame 
-position changes are relatively small. The $\epsilon$-random uniform 
-transitions increase tracking robustness, by forcing the filter to always 
-sample broadly across the pixel space. This helps prevent overconfidence 
-and collapse, and decreases reacquisition delay.
+position changes are relatively small.
 
@@
 \subsubsection{Emission Model}
-Our emission model (Equation~\ref{eq:emission_model}) implements a sampling
+The emission model weights the particles' likelihood of containing a pixel location that is part of the H$^2$Bird.
+The particles with higher weights have a greater probability of being resampled.
+Our emission model implements a sampling
 approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
-differencing~\cite{Ahad2011Computer}. On each iteration, the system provides the emission 
+differencing~\cite{Ahad2011Computer}: 
+\begin{equation}
+\label{eq:emission_model}
+w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
+\end{equation} 
+On each iteration, the system provides the emission 
 model with $f_0$, the frame representing the background, and $f_t$, the most 
@@
 between the pixels of $f_0$ and $f_t$ at the location $x^{[i]}_t$.
-\begin{equation}
-\label{eq:emission_model}
-w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
-\end{equation} 
 Hence, particles are more likely to be resampled in regions of 

### 2013-02-05 manuscript.tex
more edits in section 3

@@
 In Equation~\ref{eq:motion_model}, $x^{[i]}_t$ is the position of the $i^\text{th}$ 
-particle in pixel space at time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ 
+particle in 640x480 pixel space at time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ 
 is a vector of the bounds of the pixel space. The sampling method requires that, with 

### 2013-02-05 manuscript.tex
fixed bibliography, other text things

@@
 H$^2$Bird steers to the heading directed by the ground station using a
-quaternion-based PID controller.
+quaternion-based proportional-integral-derivative controller.
 
@@
 
-In this section, we describe the models that we developed to simulate the
-behaviors of the ornithopter and the ground station. We use these models to
+We developed models using system identification concepts to simulate the
+behaviors of the H$^2$Bird and the ground station. We use these models to
 determine the set of initial conditions that are most likely to result in
@@
 \end{equation}
-We choose the reference heading angle as the input to the model, $u(s)$, and
+We choose the reference heading angle as the input, $u(s)$, to the model, and
 model the dynamics of vehicle and on-board PID controller as one system, whose
-output is the physical heading angle of the H$^2$Bird, $\theta(s)$. To model the translational position of
+output, $\theta(s)$, is the physical heading angle of the H$^2$Bird. To model the translational position of
 the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning}:
@@
 Our model of the ground station includes models for both the camera and the
-pose estimation computation. The camera model contains both a latency, which 
-is in series with the path between the ``Camera'' and 
-``Ornithopter Dynamics'' blocks in Figure~\ref{fig:block_diagram}, and a 
-model of the camera's viewing angle constraints.
+pose estimation computation. Both models contain latencies, which we
+identified in several locations in the block diagram in
+Figure~\ref{fig:block_diagram}. The camera model experiences a delay in series
+with the path between the ``Camera'' and ``Ornithopter Dynamics'' blocks. The
+pose estimation algorithm experiences latency in the pose estimation block
+when updating the particle filter, and during the transmission of the desired
+heading from the external PID controller to the internal PID controller on
+the robot, via radio.
 
-We model the viewing angle of the camera as a pixel map from one edge of a 
-63$^{\circ}$ viewing triangle to the other. The camera has an image width of 
-640 pixels. We calculate the desired heading angle in degrees, for a given 
-distance from the camera:
+In addition to the latencies in the system, we model the viewing angle of the
+camera as a pixel map from one edge of a 63$^{\circ}$ viewing triangle to the
+other. The camera has an image width of 640 pixels. We calculate the desired
+heading angle in degrees, for a given distance from the camera:
 \begin{equation} 
@@
 
-We experienced latency during experiementation in estimating the pose of the
-ornithopter on the ground station. Referring to
-Figure~\ref{fig:block_diagram}, there were delays in the pose estimation block
-while updating the particle filter, in the PID controller block when
-calculating the new reference attitude, and in the transmission of the desired
-heading from the external controller to the controller on the robot via radio.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments, Results, and Performance}
+\section{Experiments and Results}
 \label{sec:performance}
@@
 We measured a camera capture latency of 40 ms, pose estimation computation 
-latency of 12.5 ms, and radio latency of 25 ms.
+latency of 12.5 ms, and radio transmission latency of 25 ms.
 %----------------------------------------------------------------------------%
@@
 green. The geometrically infeasible region is computed using the minimum 
-turning radius and the geometry of the camera's field-of-view and is 
-indicated in blue. The region in red is not geometrically feasible, but is 
+turning radius and the geometry of the camera's field-of-view, and is 
+indicated in blue. The region in red is geometrically feasible, but is 
 infeasible as a result of the cooperative control simulation dynamics.
@@
 this information we can make assumptions about ideal starting positions for
-cooperative control. For each point within a 10 cm grid within the camera
+cooperative control. For each point in a 10 cm grid within the camera
 viewing triangle, we simulated 40 trials using an initial heading randomly
@@
 box over it using a mouse. 
-\item[\textbf{(3)}] We release H$^2$Bird by hand, in view of the camera at the desired 
+\item[\textbf{(3)}] We release the H$^2$Bird by hand, in view of the camera at the desired 
 starting grid point.
@@
 
-Conducting the experimental trials as described in 
-Section~\ref{sec:experiments_verification}, we achieved a success rate of 
+As a result of the experimental trials, we achieved a success rate of 
 80\% for initial conditions in the middle of the testing space and in front 
@@
 The lower success rate on the edges of the camera frame can be explained by
-several factors. On the edges of the viewing triangle, the yaw control 
-inputs are large in magnitude, although this is not necessary in all cases. 
-Since the control algorithm we are using has no notion of depth, it applies 
-the same control input whether the ornithopter is close to the camera, where 
-small changes in heading are adequate, or far away from the camera, where 
-large changes in heading are necessary. Due to this phenomenon, the 
-controller often over or under-compensates in certain regions, depending 
-upon the PID tuning and the distance from the camera. Tracking noise can 
-also cause passage negotiation to fail. Variations in the H$^2$Bird hand 
-launch can cause the system to respond differently for similar initial 
-conditions. In addition, we changed the H$^2$Bird battery approximately every 5 
-tests, adding more variation to the H$^2$Bird performance by affecting the 
-motor power output.
+several factors. On the edges of the viewing triangle, the yaw control inputs
+are large in magnitude, although this is not necessary in all cases. Since the
+control algorithm we are using has no notion of depth, it applies the same
+control input whether the ornithopter is close to the camera, where small
+changes in heading are adequate, or far away from the camera, where large
+changes in heading are necessary. Due to this phenomenon, the controller often
+over or under-compensates in certain regions, depending upon the PID tuning
+and the distance from the camera. Tracking noise and variations in the
+H$^2$Bird hand launch can cause the system to respond differently for similar
+initial conditions, or to fail completely. In addition, we changed the
+H$^2$Bird battery approximately every 5 tests, adding more variation to the
+H$^2$Bird performance by affecting the motor power output.
 
@@
 experiments corroborates our calculated probability of success of
-approximately 80\%. Additionally, the diminished success rate of 50\% near the
-edges of the viewing triangle that we determined through experimentation
-supports the reduced probability of success calculated by the simulation.
+approximately 80\%, demonstrated by the Monte Carlo simulation. Additionally,
+the diminished success rate of 50\% near the edges of the viewing triangle
+that we determined through experimentation supports the reduced probability of
+success calculated by the simulation.
 
@@
 simulation. While there are successes along the boundary of the camera 
-viewing triangle, there are also many failures. We further explore this 
-behavior using a Monte Carlo simulation.
+viewing triangle, there are also many failures.
 
@@
 We have demonstrated cooperative guidance of a 13 gram flapping-wing MAV with
-a lightweight base station in hardware. In addition, we developed a model of
+a lightweight base station. In addition, we developed a model of
 the cooperative system using system identification methods on the individual

### 2013-02-05 manuscript.tex
more references

@@
 \centering
-\includegraphics[width=\linewidth]{figures/h2bird.png}
+\includegraphics[width=0.95\linewidth]{figures/h2bird.png}
 \caption{H$^2$Bird ornithopter MAV}
@@
 passages, by exploiting cooperation between specialized robotic agents.
-%
 
@@
 capture requires extensive and stationary modification of the environment.
-The power and payload limits on flapping-wing MAVs make complex on-board 
-sensing difficult, so multiagent cooperative control schemes are apt 
-solutions.
 
@@
 advantage of high-performance processors and specialized cameras. Stirling et
-al. describe a method for guiding quadrotors, designated as explorers, though
-a network of stationary quadrotors, designated as beacons, through an indoor
-environment~\cite{SwarmStirling}. The beacons provide external sensing for
-navigating the explorers through the environment, and the network can be
-extended by using explorers as beacons in undiscovered areas. Their
-implementation relies solely on local sensing, although our size range
-requires that the actual environment sensing be external to the MAV.
-Additionally, they do not consider time constraints in their planning, whereas our MAV has a minimum forward speed and turning radius.
+al. describe a method for cooperation amonst quadrotor MAVs to navigate indoor
+environment~\cite{SwarmStirling}. Stationary MAVs provide external sensing for
+navigation through the environment, and the network can be extended by using
+the exploring MAVs for sensing in undiscovered areas. Their implementation
+relies solely on local sensing, although our size range requires that the
+actual environment sensing be external to the MAV. Additionally, they do not
+consider time constraints in their planning, whereas our MAV has a minimum
+forward speed and turning radius.
 
@@
 centered around each particle's current position, augmented with an 
-$\epsilon$-random uniform sampling strategy. The sampling law is as follows:\begin{equation}
+$\epsilon$-random uniform sampling strategy~\cite{Doucet2001}. The sampling law is as follows:
+\begin{equation}
 \label{eq:motion_model}
@@
 \subsubsection{Emission Model}
-The emission model weights the particles' likelihood of containing a pixel location that is part of the H$^2$Bird.
-The particles with higher weights have a greater probability of being resampled.
-Our emission model implements a sampling
-approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
-differencing~\cite{Ahad2011Computer}: 
+The emission model weights the particles' likelihood of containing a pixel
+location that is part of the H$^2$Bird~\cite{Doucet2001}. The particles with
+higher weights have a greater probability of being resampled. Our emission
+model implements a sampling approach to motion tracking via na\"{i}ve
+background subtraction, e.g. frame differencing~\cite{Ahad2011Computer}:
 \begin{equation}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/experiment_cartoon.pdf}
+\includegraphics[width=0.9\linewidth]{figures/experiment_cartoon.pdf}
 \caption{Conceptual sketch of experimental setup.}

### 2013-02-05 manuscript.tex
edits

@@
 
-In this section, we present the numerical results of the experiments we used to 
-determine the values of the parameters in our system model.
+We present the numerical results of the experiments we used to determine the
+values of the parameters in our system model for the H$^2$Bird and the ground
+station.
 

### 2013-02-07 manuscript.tex
final edits

--- 2013-02-05 manuscript.tex
@@
 % METADATA
-\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with Millirobots}
+\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with an 
+Ornithopter MAV and Lightweight Ground Station}
 \def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
@@
 \renewcommand{\floatpagefraction}{0.90}
+
+\clubpenalty=10000 
+\widowpenalty = 10000
 
@@
 accurately predicts the backwards-reachable region for successfully 
-negotiating ornithopther flight through narrow passages.
+negotiating ornithopter flight through narrow passages.
 
@@
 control with a lightweight ground station, which has power and weight 
-requirements appropriate for deployment on a ground vehicles with 10 gram 
+requirements appropriate for deployment on ground vehicles with 10 gram 
 payloads. The ground station provides heading estimates to the ornithopter by 
@@
 \centering
-\includegraphics[width=0.95\linewidth]{figures/h2bird.png}
+\includegraphics[width=\linewidth]{figures/h2bird.png}
 \caption{H$^2$Bird ornithopter MAV}
@@
 \end{figure}
-Power, size, and weight constraints on 10 gram scale micro air vehicles (MAV) 
+Power, size, and weight constraints on 10 gram scale micro air vehicles (MAVs) 
 significantly limit their sensing and computational capabilities. MAVs can 
@@
 real-time.  The constraints on flapping-wing MAVs are even more restrictive. 
-Though they demonstrate safety and noise profiles superior to those of 
+Although they demonstrate safety and noise profiles superior to those of 
 rotorcraft MAVs, this comes at the expense of reduced payload.
@@
 
-Many control laws researchers have developed from dynamic models of 
-flapping-wing flight, such as periodic forcing and other control methods 
+Many control laws, such as periodic forcing and other control methods 
 described by~\cite{doman:dynamics}\cite{khan:longitudinal_control}\cite{leonard:averaging}, 
@@
 advantage of high-performance processors and specialized cameras. Stirling et
-al. describe a method for cooperation amonst quadrotor MAVs to navigate indoor
+al. describe a method for cooperation amongst quadrotor MAVs to navigate an indoor
 environment~\cite{SwarmStirling}. Stationary MAVs provide external sensing for
@@
 cannot. Despite this advantage, in order to collectively equal the sensing and 
-computational capabilities of larger robots, small robots will also have to interact 
+computational capabilities of larger robots, small robots will have to interact 
 so that multiple robots can benefit from the specialized capabilities of 
@@
 \centering
-\includegraphics[height=150pt]{figures/h2bird_axes.pdf}
+\includegraphics[height=170pt]{figures/h2bird_axes.pdf}
 \caption{H$^2$Bird ornithopter with attitude control axes and labeled 
@@
      {https://github.com/biomimetics/imageproc\_pcb}} 
-controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
+controller includes a 40 MIPS microprocessor, 6 DOF IMU, 
 IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
@@
 Cooperative behavior is a promising strategy for autonomous navigation in 
-unexplored and unstructured environments. However, to operate in these 
+unexplored and unstructured environments. To operate in these 
 environments, a cooperative system must be fully mobile, not tied to a 
@@
 
-In the experiments we present in this work, we used a diagonal covariance matrix 
+In the experiments, we used a diagonal covariance matrix 
 with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
@@
 
-During the experiments presented in this work, we assumed that neither the 
+We assumed that neither the 
 bird, nor any other moving object, were visible in the frame on system 
-start. We saved the first frame captured by the system, and provided it as 
-the background frame $f_0$ for every iteration of the particle filter. This 
-choice of $f_0$ resulted in a sampling approach to na\"{i}ve background 
-subtraction.
+start. The first frame was captured by the system, and used as 
+the background frame $f_0$ for every iteration of the particle filter.
 
 Na\"{i}ve background subtraction is brittle to changes in a scene over 
-time~\cite{Ahad2011Computer}. However, we find it very efficient, and sufficiently 
+time~\cite{Ahad2011Computer}. However, we find it very efficient and sufficiently 
 reliable for the short time spans--less than five seconds--necessary for 
 H$^2$Bird to successfully negotiate narrow openings. Calculating $f_0$ 
-using a more sophisticated background extraction would retrain the 
+using a more sophisticated background extraction would retain the 
 underlying motion tracking features of the filter, while providing the 
@@
 the estimate of 2D vehicle pose from the pose estimation algorithm, the 
-ground station calculates an altitude error and horizontal error. The ground 
+ground station calculates an altitude error and horizontal error between the H$^2$Bird and the goal. The ground 
 station then uses a PID control law to calculate a new reference pitch and yaw, 
@@
 %----------------------------------------------------------------------------%
-\subsection{Ornithopter Attitude Control}
+\vspace{0.28in}\subsection{Ornithopter Attitude Control}
 \label{sec:model_attitude}
@@
 limit complexity, and because it matches our physical implementation. The H$^2$Bird 
-uses an on-board PID controller to follow a particular heading sent by the ground 
-station, and the yaw dynamics are largely decoupled from the rest of the fight dynamics, 
-so our one-dimensional model is a good representation of the yaw motion:
+uses an on-board PID controller to follow the heading sent by the ground 
+station, and the yaw dynamics are largely decoupled from the rest of the fight dynamics. 
+Thus, the yaw response $\theta$ is modeled as:
 \begin{equation}
 \label{eq:plain_model}
-\begin{aligned}
-& H(s) = \frac{K}{1 + Rs}
-& \text{where }H(s) = \frac{\theta(s)}{u(s)}
-\end{aligned}
+\theta(s) = \frac{K}{1 + Rs}u(s)
 \end{equation}
-We choose the reference heading angle as the input, $u(s)$, to the model, and
-model the dynamics of vehicle and on-board PID controller as one system, whose
-output, $\theta(s)$, is the physical heading angle of the H$^2$Bird. To model the translational position of
+with constants K and R from system identification, reference heading angle $u(s)$, and
+output $\theta(s)$, the physical heading angle of the H$^2$Bird. To model the translational position of
 the ornithopter, we use a form of Dubin's car model~\cite{lavalle:planning}:
@@
 In the car model, $\phi_{max}$ is the maximum steering angle and $u_{s}$ is
-the forward speed. We modify this model to treat the angular position $\theta$
-as an input rather than a state. The overall motion of the system can be
+the forward speed. Here, the angular position $\theta$
+is an input rather than a state. The overall motion of the system can be
 modeled as shown in Equation~\ref{eq:dubin_car}, where $\theta$ is the
@@
 Our model of the ground station includes models for both the camera and the
-pose estimation computation. Both models contain latencies, which we
-identified in several locations in the block diagram in
-Figure~\ref{fig:block_diagram}. The camera model experiences a delay in series
+pose estimation computation. Referring to
+Figure~\ref{fig:block_diagram}, the camera model experiences a delay in series
 with the path between the ``Camera'' and ``Ornithopter Dynamics'' blocks. The
@@
 the horizontal position of the robot relative to the center of the window in 
-meters. This relation reflects the fact that the pose estimation algorithm 
+meters. Note that the pose estimation algorithm 
 does not determine the depth of the robot, which can be problematic for 
 control algorithms. For example, if the ornithopter remains at a constant 
-horizontal distance from the window, but moves closer and closer to the 
+horizontal distance from the window, but moves closer to the 
 camera, the ground station's directed heading steadily increases, even 
@@
 \subsection{System Identification}
+\label{sec:system_id}
 
-In this section, we present the numerical results of the experiments we used to 
-determine the values of the parameters in our system model.
-
-%............................................................................%
-\subsubsection{Ornithopter Attitude Control}
-\label{sec:flight_control}
 We determined the turning speed and turning radius of the H$^2$Bird 
@@
 recording the attitude estimates calculated by the ornithopter using its 
-on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
+on-board gyroscope. We then used MATLAB\footnote{\raggedright The MathWorks, Inc. Matlab:
 \href{http://www.mathworks.com/products/matlab/}
      {http://www.mathworks.com/products/matlab/}} 
-to fit a simple low-order model to the step response, as discussed in 
-Section~\ref{sec:model_attitude}.
+to fit a simple low-order model to the yaw step response, as discussed in 
+Section~\ref{sec:model_attitude}:
 
@@
 
-The first-order model that we fit to the step response of the onrithopter 
-yaw motion is as folows:
 \begin{equation}
 \label{eq:transfer_func}
-H(s) = \frac{0.95}{1+0.62s}
+\theta(s) = \frac{0.95}{1+0.62s}u(s)
 \end{equation}
-This model represents the tuning dynamics of the ornithopter as well as the
-PID controller providing inputs to the actuators. We present the results of
-the the step response experiment, along with the response of the fitted model,
-in Figure~\ref{fig:step_response}.
 
-We calculated the step response rise time to be 1.4 seconds. The H$^2$Bird flies at an 
-average forward velocity of 1.2 m/s during our experiments, so we estimate 
+We calculated the step response rise time to be 1.4 seconds (Figure~\ref{fig:step_response}). The H$^2$Bird flies at an 
+average forward velocity of 1.2 m/s, thus we estimate 
 the minimum turning radius of the robot to be 1.07 meters:
@@
 \end{equation}
-In the experiments we present in this work, we choose a flight speed lower 
+In the experiments, we choose a flight speed lower 
 than the maximum speed of the ornithopter, to facilitate more robust 
@@
 %............................................................................%
-\subsubsection{Ground Station Pose Estimation}
 To measure the capture latency of the camera, we wrote a simple test
@@
 timers to determine the average time between the start and end of computation.
-
 We measured a camera capture latency of 40 ms, pose estimation computation 
@@
 %----------------------------------------------------------------------------%
-\subsection{Model Verification}
-
-We validate our system model experimentally by systematically guiding the
-H$^2$Bird through the window at various starting initial conditions. We
-compare the results of the experiments to the computed backwards reachable set
-of feasible initial conditions and a Monte Carlo simulation to determine the
-accuracy of our system model.
+\subsection{Model Verification by Simulation}
 
 %............................................................................%
-\subsubsection{Simulation}
 
@@
 and angular heading initial conditions. We use our model, as determined
-through system identification as described in Section~\ref{sec:system_model},
+through system identification and described in Section~\ref{sec:system_id},
 using the cooperative control method in Figure~\ref{fig:block_diagram}. The
@@
 
-In Figure ~\ref{fig:feasible_set}, 
+In Figure~\ref{fig:feasible_set}, 
 the blue lines represent the camera viewing triangle, the black line represents the window, and the black 
@@
 %............................................................................%
-\subsubsection{Experiments}
+\subsection{Experiments}
 \label{sec:experiments_verification}
@@
 experiment: the H$^2$Bird must fly through a specified obstacle, in this 
-case, a wooden frame we used to simulate a window.
+case, a wooden frame we used to simulate a window. The ground station provides remote guidance to 
+H$^2$Bird to allow it to navigate through the window. A motion capture 
+system records ground truth data. Our setup is shown 
+in Figure~\ref{fig:experiment_cartoon}.
 
@@
 \end{figure}
-
-As the H$^2$Bird cannot detect the target on its own, we used cooperative 
-control with the ground station, which provided remote guidance to 
-H$^2$Bird to allow it to navigate through the window. We used a motion capture 
-system to record ground truth data. We present a sketch of this experimental 
-setup in Figure~\ref{fig:experiment_cartoon}.
 
@@
 
-Each trial included the following steps (see Figure~\ref{fig:pf_screencap}):
+Each trial included the following steps (Figure~\ref{fig:pf_screencap}):
 \begin{spaced_description}{\textbf{(4-6)}}
@@
 the testing space. We conducted 20 additional trials directly in front of 
-the camera, to determine the success rate near the center of viewing plane.
+the camera to determine the success rate near the center of viewing plane.
 
@@
 \caption{Plot of camera field of view and experimental trials within the feasible region.}
-\label{fig:flight_paths_feasible}
+\label{fig:flight_paths_feasible}\
 \end{minipage}
@@
 
-As a result of the experimental trials, we achieved a success rate of 
+The H$^2$Bird achieved a success rate of 
 80\% for initial conditions in the middle of the testing space and in front 
-of the camera. As we moved towards the edges of the feasible region in 
+of the camera. As the H$^2$Bird moved towards the edges of the feasible region in 
 Figure~\ref{fig:flight_paths}, the success rate diminished to 50\%.
@@
 several factors. On the edges of the viewing triangle, the yaw control inputs
-are large in magnitude, although this is not necessary in all cases. Since the
-control algorithm we are using has no notion of depth, it applies the same
+can be large in magnitude. Since the
+control algorithm has no notion of depth, it applies the same
 control input whether the ornithopter is close to the camera, where small
@@
 H$^2$Bird hand launch can cause the system to respond differently for similar
-initial conditions, or to fail completely. In addition, we changed the
-H$^2$Bird battery approximately every 5 tests, adding more variation to the
-H$^2$Bird performance by affecting the motor power output.
+initial conditions, or to fail completely. In addition, the
+H$^2$Bird performance is affected by battery power variation.
 
@@
 unstructured environments by deploying our base station on millirobotic ground
-vehicles. As the ground station used in this work was designed with this goal
-in mind, this is a matter of implementation. The effectiveness of our modeling
+vehicles. The effectiveness of our modeling
 approach, when applied to more complex systems with more interactions and 
@@
 \section{Acknowledgments}
-The authors would like to thank Fernando Garcia Bermudez for his 
+The authors thank Fernando Garcia Bermudez for his 
 assistance with the Vicon motion capture system, Andrew Pullin for his 

### 2013-02-09 manuscript.tex
moving headings
shifted some text

@@
 % METADATA
-\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with an 
-Ornithopter MAV and Lightweight Ground Station}
+\def \papertitle{Cooperative Control and Modeling for Narrow Passage Traversal with an Ornithopter MAV and Lightweight Ground Station}
 \def \paperkeywords{AAMAS proceedings, multiagent systems, cooperative control, 
@@
 forward speed and turning radius.
-
+\looseness -1
 The small size of lightweight air and ground vehicles necessitates 
@@
 with quaternion multiplication~\cite{hamilton1866elements}.
-
+\looseness -1
 Our implementation of quaternion-based control is similar to 
@@
 %----------------------------------------------------------------------------%
-\vspace{0.28in}\subsection{Ornithopter Attitude Control}
+\subsection{Ornithopter Attitude Control}
 \label{sec:model_attitude}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusions and Future Work}
@@
 the cooperative system using system identification methods on the individual
-system capabilities, and composing it with known system interactions. The
-corroboration between our simulated and experimental results shows that our
+system capabilities, and composing it with known system interactions.
+
+\noindent The corroboration between our simulated and experimental results shows that our
 simplified modeling approach is useful for predicting the performance of

### 2013-02-10 manuscript.tex
Add footnote for Biomimetics GitHub page

@@
 We develop a new flapping-winged MAV, the H$^2$Bird, and ground station. We
-use these sytems to cooperatively navigate through a narrow passage, by
+use these systems to cooperatively navigate through a narrow passage, by
 sensing the position of the MAV and the goal using the ground station.
@@
 USB radio module which implements the IEEE 802.15.4 wireless standard.
-%TODO footnotes: Gumstix, LogidPD
 
@@
 \href{http://opencv.willowgarage.com/wiki/}{http://opencv.willowgarage.com/wiki/}} 
-computer vision libraries for image capture and processing.
-%TODO mention code is open source, available on GitHub
+computer vision libraries for image capture and processing. All software and 
+hardware we developed for this system is open source and freely available under 
+a BSD license\footnote{\raggedright Biomimetic Millisystems Laboratory on GitHub, \href{https://github.com/biomimetics}{https://github.com/biomimetics}}.
 
